---
title: "Astragalus microcymbus mark-recapture"
author: "Michelle DePrenger-Levin"
date: "2023-05-25"
output:
  word_document: default
  html_document: default
---

MacKenzie et al 2009 Modeling species occurrence dynamics with multiple states and imperfect detection       
        1. Define the region (population) for occurrence, state classification (of the whole population, not individual) - Population level dynamics. close, not quite    
        2. However, if could combine the idea of uncertainty only going one direction (if observed, could be vegetative, or reproductive) to individuals, then this would be great.   

Detection probability matrix   
    For any true state, survey must result in one of the observed states; probability of observing state l during survey j when true state is m $p^(l,m)_j$   

_____________________________________________________________________________________________________
Occupied in state m: $\phi^[m]$    
The state vector must sum to 1   
All individuals will be at least vegetative and possibly reproductive at some point during the growing season
   $\phi^1$: Vegetative    
   $\phi^2$: Reproductive
   1-$\phi^1$-$\phi^2$-$\phi^3$: unoccupied (not alive)     
        
$\phi_0$ = [1-$\phi^1$-$\phi^2$      $\phi^1$       $\phi^2$]
Ambiguity of observations only extends in one direction -->    

Observed state: 0=unobserved, 1=Vegetative and no reproduction, 2=flower or fruit    
True state: 0=unoccupied- not alive, 1=vegetative, 2=reproductive   
$$
    \begin{bmatrix}
    1                           & 0             & 0\\
    1- p^(1,1)_j                & p^(1,1)_j     & 0\\
    1- p^(1,1)_j - p^(2,2)_j    & p^(1,2)_j     & p^(2,2)_j\\
    \end{bmatrix}
$$
    
    
Observing 1 means 0 isn't true, observing 2 means 1 and 0 aren't true state   
$h_i$ = 120, then the state dependent probability vector is:    
 

$$ 
    p_120 =  
    \begin{bmatrix}
    0 \\
    0 \\
    p^(1,2)_1  p^(2,2)_2  (1-p^(1,2)_1 - p^(2,2)_2 
    \end{bmatrix}
$$


Reparameterized as sets of conditional binomial probabilities; so that $p^[m]$ is the probability of correctly observing the true state $m$, and $\beta^(l,m)$ is the the probability of observing the state $l$ when the true state is $m$ and $l < m$       
      1. These binomials allow for covariates    
$$
    \begin{bmatrix}
    1                                   & 0             & 0\\
    1 - p^[1]_j                         & p^[1]_j         & 0\\
    (1 - \beta^(1,1)_j) (1 - p^[2]_j)    & \beta^(1,2)_j (1-p^[2]_j)  & p^[2]_j\\
    \end{bmatrix}
$$
  
# Multi-years  
Transition matrix    
     * represented as a probability row vector given the current location (row) of movement among k states:

$$p = (p_1,...,p_k) $$
$$p_i$$ is the probability that individual in state i    
     
$$p^(j)$$ is state at step , then state at step j+1 is the matrix product    
$$p^(j+1) = p^(j)P$$   
After m additional transitions, the state of individual p is $$p^(j+m)$$

     

In a season $t$, the true state of individual, $i$ is a random variable $(z_{i,t})$ from one trial of a multinomial distribution (or could be from any reparameterization, correct?)    
$z_{i,1} \sim mult(\phi_0,1)$    
$z_{i,t}|z_{i,t-1} \sim mult(\phi_{t-1}(z_{i,t-1},...),1)$     
Then given the imputed true state of an individual in season t, the observed state in survey j of individual i is a random variable $h_{i,t,j}$
$h_{i,t,j} \sim mult(p_{t,j}(z_{i,t},...),1)$

Can calculate the number of individuals in each state   
Covariates and missing observations through link functions, logit, multinomial-logit

```{r}
rm(list=ls())
library(rjags)
library(dplyr)
library(tidyr)
library(ggplot2)
library(R2jags) ## jags
library(R2WinBUGS)

```



Gimenez et al 2007 State-space modelling of data on marked individuals  


conver BUGS to jags   
```{r}

SSM_CJS <- "
model
{
  # Likelihood
  ## System process
  for (i in 1:n) {
    PrX[i, e[i]] <- 1                 # if first time marked, e, Pr(alive | first capture) = 1
    X[i, e[i]+1] ~ dbern(PrX[i,e[i]]) # alive then 1 is generated with certainty
    PrO[i, e[i]] <- X[i, e[i]+1]      # detection at initial marking is 100%
    
      for(j in (e[i]+1):K) {
        PrX[i,j] <- phi[j] * X[i,j]   # Add covariates here??
        X[i,j+1] ~ dbern(PrX[i,j])    # survival from one year to next
        PrO[i,j] <- p[j] * X[i,j+1]   # probability of observing given it survived??
      }
      ## every other cell gets a zero
      for (j in 1:(e[i]-1)) {
        PrX[i,j] <- 0
        X[i,j] <- 1
        PrO[i,j] <- 0
      }
  
      ## Observation process
      for (h in 1:nx) {
        data[h,3] ~ dbern(PrO[data[h,1], data[h,2]])
      }
  }
    
  # Priors
  p[1] <- 1    # detectability
  phi[1] <- 1  # Survival
  phi[1] ~ dbeta(1,1)
  for (j in 2:K) {  ## Annual survival and detectability change
    phi[j] ~ dbeta(1,1)
    p[j] ~ dbeta(1,1)
  }
}"

```

K = number of encounter occasions
n = individuals
nx is K * n
e = vector of first encounters (length n)
data = matrix with nx rows and (1) current individual (1...n), (2) current encounter occasion (1,...,K), and (3) observation (1 or 0)

Make data from AsMi - Use CJS_SeedlingSurvival.Rmd  
```{r}
## Only 2013 data
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/model_data.Rdata")

```




```{r}
# # Set up the data
# model_data <- list(
#   N = N, y = y, X = X,
#   zeros = rep(0, N),
#   p = p,
#   D_inv = D_inv,
#   D = D,
#   W = W,
#   I = I
# )
# 
# # Choose the parameters to watch
# model_parameters <- c("alpha", "tau", "beta")
# 
# # Run the model
# model_run <- jags(
#   data = model_data,
#   parameters.to.save = model_parameters,
#   model.file = textConnection(model_code)
# )

```



















following <https://www.montana.edu/rotella/documents/502/lab07RMark.html> 

```{r}
library(RMark)


```



Prepare data for for RMark 
```{r}

```

