---
title: "Astragalus microcymbus mark-recapture"
author: "Michelle DePrenger-Levin"
date: "2023-05-25"
output:
  word_document: default
  html_document: default
---

MacKenzie et al 2009 Modeling species occurrence dynamics with multiple states and imperfect detection       
        1. Define the region (population) for occurrence, state classification (of the whole population, not individual) - Population level dynamics. close, not quite    
        2. However, if could combine the idea of uncertainty only going one direction (if observed, could be vegetative, or reproductive) to individuals, then this would be great.   
      
Treat each site as an individual?     
Occupied in state m: $\phi^[m]$    
The state vector must sum to 1   
   $\phi^1$: seedling in year t    
   $\phi^2$: seedling after last census (t-1) and resprout  
   $\phi^3$: vegetative (age 2+)
   $\phi^4$: reproductive (age 2+)    
   1-$\phi^1$-$\phi^2$-$\phi^3$-$\phi^3$: unoccupied    
   
$\phi_0$ = [1-$\phi^1$-$\phi^2$-$\phi^3$-$\phi^3$      $\phi^1$       $\phi^2$       $\phi^3$        $\phi^4$]
Ambiguity of observations only extends in one direction -->    
   
   
Detection probability matrix   
    For any true state, survey must result in one of the observed states; probability of observing state l during survey j when true state is m $p^(l,m)_j$   
Observed state: 0=unobserved, 1=cotyledons, 2=small and no cotyledons, 3=thicker stem and no reproduction, 4=flower or fruit    
True state: 0=unoccupied- not alive, 1=seedling year t, 2=seedling year t-1, 3=vegetative, 4=reproductive   
If you observe cotyledons then it cannot be vegetative, reproductive, or germinated in the previous year after the last census

Cotyledons == recruit in year t   
Small could be recruit in year t or t-1 after last census (so unmarked)  
and recruit could be from reproductive in year t-1 or t-x (some years in the past)     

_____________________________________________________________________________________________________
Occupied in state m: $\phi^[m]$    
The state vector must sum to 1   
All individuals will be at least vegetative and possibly reproductive at some point during the growing season
   $\phi^1$: Vegetative    
   $\phi^2$: Reproductive
   1-$\phi^1$-$\phi^2$-$\phi^3$: unoccupied (not alive)     
        
$\phi_0$ = [1-$\phi^1$-$\phi^2$      $\phi^1$       $\phi^2$]
Ambiguity of observations only extends in one direction -->    

Observed state: 0=unobserved, 1=Vegetative and no reproduction, 2=flower or fruit    
True state: 0=unoccupied- not alive, 1=vegetative, 2=reproductive   
$$
    \begin{bmatrix}
    1                           & 0             & 0\\
    1- p^(1,1)_j                & p^(1,1)_j     & 0\\
    1- p^(1,1)_j - p^(2,2)_j    & p^(1,2)_j     & p^(2,2)_j\\
    \end{bmatrix}
$$
    
    
Observing 1 means 0 isn't true, observing 2 means 1 and 0 aren't true state   
$h_i$ = 120, then the state dependent probability vector is:    
 

$$ 
    p_120 =  
    \begin{bmatrix}
    0 \\
    0 \\
    p^(1,2)_1  p^(2,2)_2  (1-p^(1,2)_1 - p^(2,2)_2 
    \end{bmatrix}
$$


Reparameterized as sets of conditional binomial probabilities; so that $p^[m]$ is the probability of correctly observing the true state $m$, and $\beta^(l,m)$ is the the probability of observing the state $l$ when the true state is $m$ and $l < m$       
      1. These binomials allow for covariates    
$$
    \begin{bmatrix}
    1                                   & 0             & 0\\
    1 - p^[1]_j                         & p^[1]_j         & 0\\
    (1 - \beta^(1,1)_j) (1 - p^[2]_j)    & \beta^(1,2)_j (1-p^[2]_j)  & p^[2]_j\\
    \end{bmatrix}
$$
  
Multi-years  

In a season $t$, the true state of individual, $i$ is a random variable $(z_{i,t})$ from one trial of a multinomial distriubtion (or could be from any reparameterization, correct?)    
$z_i,1 ~ mult(\phi_0,1)$

```{r}

```





















following <https://www.montana.edu/rotella/documents/502/lab07RMark.html> 

```{r}
library(RMark)


```



Prepare data for for RMark 
```{r}

```

