---
title: "Capture histories"
author: "Michelle DePrenger-Levin"
date: "2023-10-11"
output: html_document
---


```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(RMark)
```


Format all years to capture histories  
```{r}
# currentyr <- as.numeric(format(as.Date(Sys.Date(),format="%Y-%m-%d"), "%Y"))
currentyr <- 2023

##To change if someone other than Michelle is running code
userpath <- "C:/Users/DePrengm/"

rawdatapath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/", currentyr, "_Astragalus-microcymbus_RawData.csv", collapse = '', sep = '')

asmi.raw <- read.csv(rawdatapath, na.strings = "na")

## plot 89 at site 1 was lost in 2020
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 89 & asmi.raw$year > 2019),]

# Site 26 had two plots added later; Plots 238 and 300 were added in 1996 with no seedlings! good!;  plot 598 was added in 2004, no seedlings, good!  
table(asmi.raw$AsMi_plot_id[asmi.raw$AsMi_plot_id %in% c(238,300,598)], 
      asmi.raw$status[asmi.raw$AsMi_plot_id %in% c(238,300,598)],
      asmi.raw$year[asmi.raw$AsMi_plot_id %in% c(238,300,598)])

# Need to change all the seedling in 2014 for sites 1 and 2 to appropriate vegetative or reproductive 
asmi.raw$status[asmi.raw$AsMi_site_id<3 & asmi.raw$year == 2014 & asmi.raw$flower == 0] <- "vegetative"
asmi.raw$status[asmi.raw$AsMi_site_id<3 & asmi.raw$year == 2014 & asmi.raw$flower == 1] <- "reproductive"

# No longer adding climate data to the database
asmi.raw <- asmi.raw[,grep(paste(c("Temp","Rain","Snow","Aug.Jul"),collapse="|"), names(asmi.raw),
                           value = TRUE, invert = TRUE)]
asmi.raw$Browsing...Status[asmi.raw$Browsing...Status == "mammal" ] <- "Mammal"
asmi.raw[asmi.raw$Browsing...Status ==""& !is.na(asmi.raw$Browsing...Status),]
asmi.raw[is.na(asmi.raw$AsMi_data_id),]
asmi.raw <- asmi.raw[!is.na(asmi.raw$AsMi_data_id),] # remove any rows with NA for data_id

# Which plot # go with which site
plotsXsite <- as.data.frame(table(asmi.raw$AsMi_plot_id,asmi.raw$AsMi_site_id))
plotsXsite[plotsXsite$Freq > 0,]

# one blank row at the end of the csv download, gone by 2021
asmi.raw[asmi.raw$status == "",]
# asmi.raw <- asmi.raw[asmi.raw$status != "",]
asmi.raw$status <- factor(asmi.raw$status)
table(asmi.raw$status)

#reset factors of browsing to eliminate " "
# asmi.raw$Browsing...Status[asmi.raw$Browsing...Status == "mammal"] <- "Mammal"
asmi.raw$Browsing...Status <- factor(asmi.raw$Browsing...Status)

#reset factors for fence
asmi.raw$fence <- factor(asmi.raw$fence)
table(asmi.raw$fence)

asmi.raw[asmi.raw$length>90 & !is.na(asmi.raw$length),] 
asmi.raw$length[asmi.raw$length == 921] <- 21

# If there are fruit, it flowered and needs to be reproductive, not vegetative
asmi.raw[asmi.raw$status=="vegetative" & asmi.raw$fruit>0,] # 5 times in 2020
wrongAsMidataid <- asmi.raw$AsMi_data_id[asmi.raw$status=="vegetative" & asmi.raw$fruit>0]
asmi.raw$flower[asmi.raw$AsMi_data_id %in% wrongAsMidataid] <- 1
asmi.raw$status[asmi.raw$AsMi_data_id %in% wrongAsMidataid] <- "reproductive"
asmi.raw[asmi.raw$AsMi_data_id %in% wrongAsMidataid,]

# Years of study in the abstract
length(1995:currentyr)
```


Capture histories   
additional matrix of which plants were fenced which years, 0/1 for open/fence and columns are years
```{r}

str(asmi.raw) ## n = 1, y = 2

CJS_asmi <- asmi.raw %>%
  mutate(status == as.character(status)) %>%
  dplyr::mutate(status = if_else((status == "seedling" & flower == 1), "reproductive", as.character(status))) %>%
  mutate(fence = as.numeric(fence)-1) %>%
  mutate(fence = if_else(year < 2006, 0, fence)) %>%
  mutate(fence = if_else(year == 2006 & AsMi_plot_id %in% c(300, 512), 0, fence)) %>%
  mutate(fence = if_else(year > 2015, 0, fence)) %>%
  tidyr::complete(AsMi_tag_id, year, fill = list(length = 0, fence = 0)) %>%
  group_by(AsMi_site_id, AsMi_plot_id, AsMi_tag_id) %>%
  dplyr::summarise(first = min(year),
            dh = if_else(length > 0, 1, 0),
            year = year,
            status = status,
            fence = fence,
            length = length
            ) %>%
  ungroup() %>%
  tidyr::complete(nesting(AsMi_tag_id,AsMi_site_id,AsMi_plot_id,first), year = min(year):max(year), 
           fill = list(dh = 0, status = "dead", fence = 0),
           explicit = FALSE) %>%
  filter(!is.na(AsMi_site_id))  ## Work around but shouldn't be adding a new first for each

CJS_asmi %>%
  filter(!is.na(length))

ch_asmi <- CJS_asmi %>%
  group_by(AsMi_tag_id, AsMi_site_id, AsMi_plot_id, first) %>%
  dplyr::mutate(first.ln = length[year == first]) %>%
  dplyr::arrange(year, .by_group = TRUE) %>%
  dplyr::summarise(ch = paste0(dh, collapse = "")) %>%
  mutate(freq = "1;")

fence_matrix <- CJS_asmi %>%
  dplyr::select(c(AsMi_tag_id:year, fence)) %>%
  group_by(AsMi_tag_id, AsMi_site_id, AsMi_plot_id, first) %>%
  dplyr::arrange(year, .by_group = TRUE) %>%
  pivot_wider(names_from = year, values_from = fence)
  
identical(ch_asmi$AsMi_tag_id, fence_matrix$AsMi_tag_id)

## Cooch and White Chapter 2 MARK
## write.table(ch_asmi, file = "", sep = "    ", quote=FALSE, col.names=FALSE, row.names=FALSE)
## Filter out capture histories that are all 0s
ch_asmi %>%
  ungroup() %>%
  dplyr::select(ch:freq) %>%
  filter(grepl("1", ch)) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/channual.inp", 
              sep = "    ", quote=FALSE, col.names=FALSE, row.names=FALSE)


## Group by Site
ch_asmi %>%
  ungroup() %>%
  dplyr::select(AsMi_site_id:freq) %>%
  filter(grepl("1", ch)) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/channualXplot.inp", 
              sep = "    ", quote=FALSE, col.names=FALSE, row.names=FALSE)




## Multistate just use S,V,R for the states (or just V, R, 0 for non-detection)
ch_asmiMS <- CJS_asmi %>%
  mutate(dh = case_when(status == "seedling" ~ "S", 
                        status == "vegetative" ~ "V",
                        status == "reproductive" ~ "R",
                        status %in% c("dormant","dead") ~ "0")) %>%
  group_by(AsMi_tag_id, AsMi_site_id, AsMi_plot_id, first) %>%
  dplyr::mutate(first.ln = rep(length[year == first], length.out = n()))  %>%
  ungroup() %>%
  group_by(AsMi_tag_id, AsMi_site_id, AsMi_plot_id,first,first.ln) %>%
  dplyr::arrange(year, .by_group = TRUE) %>%  ## fine, they're in the correct order
  dplyr::summarise(ch = paste0(dh, collapse = "")) %>%
  ungroup() %>%
  mutate(freq = "1;")

## Multi-state by Site
ch_asmiMS %>%
  ungroup() %>%
  dplyr::select(AsMi_site_id:freq) %>%
  filter(grepl("1", ch)) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/chMSXplot.inp", 
              sep = "    ", quote=FALSE, col.names=FALSE, row.names=FALSE)



## Robust design
# 3 primary occasions (2013-2015) and 9 primary occasions within, just skipped "." a bunch for 2013, skipped some for others
### time.intervals

save(ch_asmi, file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/capturehistoryannual.Rdata")
save(fence_matrix, file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/fencematrix.Rdata")

```

Pradel model for lambda (population growth rate) and recruitment (phi) and detection (p)
```{r}

asmi29 <- convert.inp("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/channual.inp")

asmi29.proc <- process.data(asmi29, model = "Pradel")

## Gamma p 
mod.dot <- mark(asmi29.proc, model.parameters = list(Gamma = list(formula = ~1),
                                                     p = list(formula = ~1)))

## Phi p f
asmi29.proc2 <- process.data(asmi29, model = "Pradrec")
mod.dot2 <- mark(asmi29.proc2, model.parameters = list(Phi = list(formula = ~1),
                                                       p = list(formula = ~1)))
asmi29.ddl <- make.design.data(asmi29.proc2)

mod.time <- mark(asmi29.proc2, model.parameters = list(Phi = list(formula = ~ time),
                                                       p = list(formula = ~ 1)))

mod.time$pims
mod.time$results$real
### LAMBDA!!!
mod.time$results$derived

asmi.lambda <- mod.time$results$derived$`Lambda Population Change`

ggplot(asmi.lambda, aes(1995:2022, estimate)) +
  geom_point()+
  # geom_line() +
  theme_bw() +
  geom_errorbar(aes(1995:2022, ymin = lcl, ymax = ucl),
                color = "red")+
  xlab("Year") +
  ylab(expression(lambda))+
  geom_hline(yintercept = 1, linetype = "dotted")
  # geom_ribbon(aes(ymin = lcl, ymax = ucl), fill = "red", alpha=0.25)

asmi.loglambda <- mod.time$results$derived$`log(Lambda) Population Change`

ggplot(asmi.loglambda, aes(1995:2022, estimate)) +
  geom_point()+
  # geom_line() +
  theme_bw() +
  geom_errorbar(aes(1995:2022, ymin = lcl, ymax = ucl),
                color = "red")+
  xlab("Year") +
  ylab(expression("log"~lambda))+
  geom_hline(yintercept = 0, linetype = "dotted")


## More likely recruitment changes over time (f) as well as survival 
mod.timetime <- mark(asmi29.proc2, model.parameters = list(Phi = list(formula = ~ time),
                                                           p = list(formula = ~ 1),
                                                           f = list(formula = ~ time)))

asmi.lambdatimetime <- mod.timetime$results$derived$`Lambda Population Change`
ggplot(asmi.lambdatimetime, aes(1995:2022, estimate)) +
  geom_point()+
  # geom_line() +
  theme_bw() +
  geom_errorbar(aes(1995:2022, ymin = lcl, ymax = ucl),
                color = "red")+
  xlab("Year") +
  ylab(expression(lambda))+
  geom_hline(yintercept = 1, linetype = "dotted")

asmi.loglambdatimetime <- mod.timetime$results$derived$`log(Lambda) Population Change`
ggplot(asmi.loglambdatimetime, aes(1995:2022, estimate)) +
  geom_point()+
  # geom_line() +
  theme_bw() +
  geom_errorbar(aes(1995:2022, ymin = lcl, ymax = ucl),
                color = "red")+
  xlab("Year") +
  ylab(expression(lambda))+
  geom_hline(yintercept = 0, linetype = "dotted")


```

Individual covariates can be used to model phi and p in the Pradel models. However, the biological meaning of modeling lambda as a function of an individual covariate is not clear. Intuitively, it makes more sense to model f and gamma as functions of individual covariates, even though these parameters can be combined with phi to provide a derived estimate of lambda.


Robust design, three years, single stage   
```{r}
# obsmodel
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Robust3years.Rdata")

## 3 = Reproductive, 2 = seedling, 1 = vegetative (can combine 1 and 2)
## Need to string together all 
# obsmodelsingle <- obsmodel %>%
#   mutate(across(Occ1:Occ10, ~ ifelse(.x > 0, 1, .x))) %>%
#   unite(ch, Occ1:Occ10, sep = "") %>%
#   mutate(freq = "1;") %>%
#   dplyr::select(ch, freq)

asmiRDinp <- asmiRD %>%
  ungroup()%>%
  dplyr::select(ch) %>%
  mutate(freq = "1;")

write.table(asmiRDinp, 
            file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/asmiRobust.inp",
            sep = "    ", quote = FALSE, col.names = FALSE, row.names = FALSE)

asmi.rdinp <- convert.inp("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/asmiRobust.inp")

time.intervals <- c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0)

asmi.rd <- process.data(asmi.rdinp, model = "Robust", time.intervals = time.intervals)
## Check primary occasions
asmi.rd$nocc
asmi.rd$nocc.secondary
  


run.robust <- function()
{
  ## 3 primary sessions with 10 secondary sessions (I could have just had different secondary session lengths)
  
  ## Apparent survival varies by year
  S.time = list(formula ~ time)
  
  
  # p varies by primary not among secondary; session:time would be both; p=c from share=TRUE c being overdispersion
  p.session = list(formula = ~ session, share = TRUE)
  
  Phi.dot = list(formula ~ 1) 
  Phi.time = list(formula ~ time) 
  p.dot = list(formula ~ 1)
  p.time = list(formula ~ time)
  
  # Gamma being the reverse time estimates: probability that an individual at time t+1 was in population at time t, recruitment
  Gamma
  
  ## RDHUggins: condition on the encounter history of not seeing an individual. divide 1-not seeing from all other histories. no individual covariates
  ## Can do model = "Pradel"
  ## "Pradrec
  ## "Multistrata" will automatically identify number of strata and labels based on unique alphabetic codes used. 
}

```



Multistate mark-recapture   
http://www.wizcity.com/Computers/Characters/GreekUTF8.php 
```{r}
grViz("digraph integrated_pop {
      
      # digraph = directed, graph = undirected, graphID
      graph [compound = true, color = blue, rankdir = TB]
      
      # node definitions with substituted label text
      node [shape = circle, fontname = Helvetica]
      # Demographic monitoring
      a [label = '@@1-1'] # seedling
      b [label = '@@1-2'] # veg
      c [label = '@@1-3'] # rep
      
      # name must start with 'cluster_'
      subgraph cluster_demography {
        # graph [style = dashed, color = Sienna]
      {rank = min; b}
      {rank = same; a;c}
      # {rank = min; b}
      # {rank = max; a;b;c}
        a -> b [label = '@@2-2', fontsize=25]
        a -> c [label = '@@2-3', fontsize=25]
        b -> b [label = '@@3-2', fontsize=25]
        b -> c [label = '@@3-3', fontsize=25]
        c -> a [style = dashed, color = LimeGreen, label = 'F', fontsize=25]
        c -> b [color = red, label = '@@4-2', fontsize=25]
        c -> c [label = '@@4-3', fontsize=25]
        
        label = 'Demography'
      }
      
      }
      
      [1]: c('Seedling','Vegetative','Reproductive')
      [2]: 'S&#968;'   # &#966;' # phi # &#x2082; subscript # seedling to
      [3]: 'S&#968;'  # &#966;' # veg to
      [4]: 'S&#968;'  # &#966;' # reprod to
      ")

## Export to save as impage 
```










## Open population capture-recapture, population growth rate   
```{r}

library(openCR)
## Fits non-spatial open-population models of CJS and JSSA (JSSA or POPAN) types. Recruitment, Pollock's robust design. Conditional called a Pradel-Link-Barker model  
## Uses secr for spatial  
# From robust design - secondary sampling within primary sessions. 


```


