---
title: "Marked CJS AsMi"
author: "Michelle DePrenger-Levin"
date: "2023-10-07"
output: html_document
---

```{r}
rm(list=ls())
library(marked)
library(ggplot2)
library(dplyr)
library(tidyr)
data("dipper")
```

https://cran.r-project.org/web/packages/marked/vignettes/markedVignette.html  
```{r}
moddip <- crm(dipper) ## constant phi and p, ch and sex factor; calls process.data, make.design.data, sjc
moddip.h <- crm(dipper, hessian = TRUE) # To get precision estimates; but to check convergence without penalty of computer the hessian, then don't use. 
moddip.h2 <- cjs.hessian(moddip) ## to compute and store in the model the variance-covariance, standard error, 95% normal confidence and estimate on link scale (logit)


dipper.proc=process.data(dipper)
dipper.ddl=make.design.data(dipper.proc)
fit.models=function()
 {
   Phi.sex=list(formula=~sex)
   Phi.time=list(formula=~time)
   p.sex=list(formula=~sex)
   p.dot=list(formula=~1)
   cml=create.model.list(c("Phi","p"))
   results=crm.wrapper(cml,data=dipper.proc, ddl=dipper.ddl,
                       external=FALSE,accumulate=FALSE)
   return(results)
 }
dipper.models=fit.models()
dipper.models # non-zero convergence value means model did not converge
dipper.models[[2]] # to get one of the models out with estimates


# covariates
dipper$weight=round(runif(nrow(dipper),0,9),0)+1
Flood=matrix(rep(c(0,1,1,0,0,0),each=nrow(dipper)),ncol=6)
colnames(Flood)=paste("Flood",1:6,sep="")
dipper=cbind(dipper,Flood)
# Add td covariate, but exclude first release as a capture - go through and make the first one zero
# splitCH and process.ch are functions in the marked package
td=splitCH(dipper$ch)
td=td[,1:6]
releaseocc=process.ch(dipper$ch)$first
releaseocc=cbind(1:length(releaseocc),releaseocc)
releaseocc=releaseocc[releaseocc[,2]<nchar(dipper$ch[1]),] #got rid of last first capture
td[releaseocc]=0
colnames(td)=paste("td",2:7,sep="")
dipper=cbind(dipper,td)

# Process data
dipper.proc=process.data(dipper)

# Create design data with static and time varying covariates
design.Phi=list(static=c("weight"),time.varying=c("Flood"))
design.p=list(static=c("sex"),time.varying=c("td"),
              age.bins=c(0,1,20))
design.parameters=list(Phi=design.Phi,p=design.p)
ddl=make.design.data(dipper.proc,parameters=design.parameters)
names(ddl$Phi)
names(ddl$p)

Phi.sfw=list(formula=~Flood+weight)
p.ast=list(formula=~age+sex+td)
model=crm(dipper.proc,ddl,hessian=TRUE,
          model.parameters=list(Phi=Phi.sfw,p=p.ast))

## create range of values to compute predicted phi and p values to plot for Flood and non-flood years for a range of weights  
newdipper=expand.grid(sex=c("Female","Male"),weight=1:10,Flood1=0,
                      Flood2=1,Flood3=1,Flood4=0,Flood5=0,Flood6=0,td2=0,td3=c(0,1),
                      td4=c(0,1),td5=c(0,1),td6=c(0,1),td7=c(0,1)) 
reals=predict(model,newdata=newdipper,se=TRUE)
reals$Phi$Flood=factor(reals$Phi$Flood,labels=c("Non-flood","Flood"))
ggplot(reals$Phi,aes(weight,estimate,ymin=lcl,ymax=ucl))+
  geom_errorbar(width=0.2)+geom_point()+geom_line()+
  xlab("\nWeight")+ylab("Survival\n")+facet_grid(Flood~.) 

## Bayesian with 95% highest posterior density interval
# Add Flood covariate
# Flood=matrix(rep(c(0,1,1,0,0,0),each=nrow(dipper)),ncol=6)
# colnames(Flood)=paste("Flood",1:6,sep="")
# dipper=cbind(dipper,Flood)
design.parameters=list(Phi=list(time.varying="Flood"))
model.parameters=list(Phi=list(formula=~Flood),
                      p=list(formula=~time+sex))
MCMCfit=crm(dipper,model="probitCJS",
            model.parameters=model.parameters,
            design.parameters=design.parameters,
            burnin=1000,iter=5000)

MCMCfit
```


 Seedling == 2    
 Reproductive == 3    
 Vegetative == 1    
 Not seen == 0     
 
 Within season    
 Assume dormancy is just another way that detection is less than zero, cannot assess separately from other less than perfect detection    
 Assume dormancy represents temporary emigration and can be distinguished from other forms if pick a period within which temporary emigration won't happen (among late season surveys?) maybe collapse 
```{r}

load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Observations3years.Rdata")

## Length of dormancy 
obsmodel %>%
  dplyr::select(Site:'4') %>%
  filter(Year == 2013)
  
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_2013.Rdata")
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_201415.Rdata")

## Make it look like dipper data
CJS_AsMi <- obsmodel %>%
  group_by(Year, Tag, Site, Plot) %>%
  mutate(state = ifelse(any(c_across('1':'9') == 3), "Reproductive", "Vegetative")) %>%
  # mutate(state = ifelse(any(c_across('1':'9') == 3), "Reproductive",
  #                       ifelse(any(c_across('1':'9') == 2), "Seedling", "Vegetative"))) %>%
  mutate(begin.time = min(which(c_across('1':'9')>0))) %>%
  filter(!is.infinite(begin.time)) %>%  ## There were a few cases with no observations in 2014
  mutate(across('1':'9', ~ ifelse(.x > 0, 1, 0))) %>%
  unite(ch, '1':'9', sep = "") %>%
  mutate(state = as.factor(state)) %>%
  filter(Year != 2013)

## Covariates: timing of seeing cotyledons, timing of reproduction
coty <- function(x) ifelse(x==2,1,0)
cotyledons <- obsmodel %>%
  group_by(Year, Tag, Site, Plot) %>%
  mutate(begin.time = min(which(c_across('1':'9')>0))) %>%
  filter(!is.infinite(begin.time)) %>%  ## There were a few cases with no observations in 2014
  mutate_at(vars('1':'9'), .funs = coty) %>%
  rename_with(~ paste0('Coty',.), '1':'9') %>%
  filter(Year != 2013)

cotyledons %>%
  pivot_longer(Coty1:Coty9, names_to = "Day", names_prefix = "Coty", values_to = "Cotyledons") %>%
  ggplot( aes(Day, Cotyledons)) +
    geom_point() +
    facet_wrap(~Year)

reprod <- function(x) ifelse(x==3,1,0)
reproduction <- obsmodel %>%
  group_by(Year, Tag, Site, Plot) %>%
  mutate(begin.time = min(which(c_across('1':'9')>0))) %>%
  filter(!is.infinite(begin.time)) %>%  ## There were a few cases with no observations in 2014
  mutate_at(vars('1':'9'), .funs = reprod) %>%
  rename_with(~ paste0('Repro',.), '1':'9') %>%
  filter(Year != 2013)

reproduction %>%
  pivot_longer(Repro1:Repro9, names_to = "Day", names_prefix = "Repro", values_to = "Reproduction") %>%
  ggplot( aes(Day, Reproduction)) +
    geom_point() +
    facet_wrap(~Year)



## currently has 2014 and 2015
AsMidf <- as.data.frame(CJS_AsMi[,c("ch","state")])

model <- crm(AsMidf)
## capture histories
model$data$data %>%
  arrange(state,freq)
sum(model$data$freq)

model.hessian <- cjs.hessian(model)


AsMi.proc <- process.data(AsMidf)
AsMi.ddl <- make.design.data(AsMi.proc)
Phi.state <- list(formula = ~ state)
model <- crm(AsMi.proc, AsMi.ddl, model.parameters = list(Phi = Phi.state), accumulate = FALSE)

## Compare several models
fit.models=function()
 {
   Phi.state = list(formula= ~ state)
   # Phi.time = list(formula= ~ time)
   # Phi.statetime = list(formula= ~ time + state)
   Phi.dot = list(formula = ~ 1)
   p.state = list(formula= ~ state)
   p.time = list(formula = ~ time)
   p.statetime = list(formula = ~ time + state)
   p.dot = list(formula= ~ 1)
   cml = create.model.list(c("Phi","p"))
   results=crm.wrapper(cml,data = AsMi.proc, ddl = AsMi.ddl,
                       external=FALSE,accumulate=FALSE)
   return(results)
 }
AsMi.models=fit.models()



```


2013 only
```{r}
CJS_AsMi_2013 <- obsmodel %>%
  group_by(Year, Tag, Site, Plot) %>%
  mutate(state = ifelse(any(c_across('1':'9') == 3), "Reproductive", "Vegetative")) %>%
  mutate(begin.time = min(which(c_across('1':'9')>0))) %>%
  filter(!is.infinite(begin.time)) %>%  ## There were a few cases with no observations in 2014
  mutate(across('1':'9', ~ ifelse(.x > 0, 1, 0))) %>%
  unite(ch, '1':'4', sep = "") %>%
  mutate(state = as.factor(state)) %>%
  filter(Year == 2013)

AsMi2013 <- as.data.frame(CJS_AsMi_2013[,c("ch","state")])

model <- crm(AsMi2013)
## capture histories
model$data$data %>%
  arrange(state,freq)
sum(model$data$freq)

model.hessian <- cjs.hessian(model)


AsMi.proc <- process.data(AsMi2013)
AsMi.ddl <- make.design.data(AsMi.proc)
Phi.state <- list(formula = ~ state)
model <- crm(AsMi.proc, AsMi.ddl, model.parameters = list(Phi = Phi.state), accumulate = FALSE)

## Compare several models
fit.models=function()
 {
   Phi.state = list(formula= ~ state)
   # Phi.time = list(formula= ~ time)
   # Phi.statetime = list(formula= ~ time + state)
   Phi.dot = list(formula = ~ 1)
   p.state = list(formula= ~ state)
   p.time = list(formula = ~ time)
   p.statetime = list(formula = ~ time + state)
   p.dot = list(formula= ~ 1)
   cml = create.model.list(c("Phi","p"))
   results=crm.wrapper(cml,data = AsMi.proc, ddl = AsMi.ddl,
                       external=FALSE,accumulate=FALSE)
   return(results)
 }
AsMi.models=fit.models()


model13 <- crm(AsMi2013,
             model.parameters = list(Phi = list(formula = ~ state), 
                                     p = list(formula = ~ state)))
model.hessian <- cjs.hessian(model13)

## Dot
model13.dot <- crm(AsMi2013,
             model.parameters = list(Phi = list(formula = ~ 1), 
                                     p = list(formula = ~ 1)))
model.hessian.dot <- cjs.hessian(model13.dot)

```


2014  
```{r}
CJS_AsMi_2014 <- obsmodel %>%
  group_by(Year, Tag, Site, Plot) %>%
  mutate(state = ifelse(any(c_across('1':'9') == 3), "Reproductive", "Vegetative")) %>%
  # mutate(state = ifelse(any(c_across('1':'9') == 3), "Reproductive",
  #                       ifelse(any(c_across('1':'9') == 2), "Seedling", "Vegetative"))) %>%
  mutate(begin.time = min(which(c_across('1':'9')>0))) %>%
  filter(!is.infinite(begin.time)) %>%  ## There were a few cases with no observations in 2014
  mutate(across('1':'9', ~ ifelse(.x > 0, 1, 0))) %>%
  unite(ch, '1':'9', sep = "") %>%
  mutate(state = as.factor(state)) %>%
  filter(Year == 2014)


AsMi2014 <- as.data.frame(CJS_AsMi_2014[,c("ch","state")])
AsMi.proc <- process.data(AsMi2014)
AsMi.ddl <- make.design.data(AsMi.proc)
Phi.state <- list(formula = ~ state)
model <- crm(AsMi.proc, AsMi.ddl, model.parameters = list(Phi = Phi.state), accumulate = FALSE)

## Compare several models
fit.models=function()
 {
   Phi.state = list(formula= ~ state)
   # Phi.time = list(formula= ~ time)
   # Phi.statetime = list(formula= ~ time + state)
   Phi.dot = list(formula = ~ 1)
   p.state = list(formula= ~ state)
   p.time = list(formula = ~ time)
   p.statetime = list(formula = ~ time + state)
   p.dot = list(formula= ~ 1)
   cml = create.model.list(c("Phi","p"))
   results=crm.wrapper(cml,data = AsMi.proc, ddl = AsMi.ddl,
                       external=FALSE,accumulate=FALSE)
   return(results)
 }
AsMi.models=fit.models()


model14 <- crm(AsMi2014,
             model.parameters = list(Phi = list(formula = ~ state), 
                                     p = list(formula = ~ state)))
model.hessian <- cjs.hessian(model14)

## Dot
model14.dot <- crm(AsMi2014,
             model.parameters = list(Phi = list(formula = ~ 1), 
                                     p = list(formula = ~ 1)))
model.hessian14.dot <- cjs.hessian(model14.dot)

```


```{r}
CJS_AsMi_2015 <- obsmodel %>%
  group_by(Year, Tag, Site, Plot) %>%
  mutate(state = ifelse(any(c_across('1':'9') == 3), "Reproductive", "Vegetative")) %>%
  # mutate(state = ifelse(any(c_across('1':'9') == 3), "Reproductive",
  #                       ifelse(any(c_across('1':'9') == 2), "Seedling", "Vegetative"))) %>%
  mutate(begin.time = min(which(c_across('1':'9')>0))) %>%
  filter(!is.infinite(begin.time)) %>%  ## There were a few cases with no observations in 2014
  mutate(across('1':'9', ~ ifelse(.x > 0, 1, 0))) %>%
  unite(ch, '1':'9', sep = "") %>%
  mutate(state = as.factor(state)) %>%
  filter(Year == 2015)


AsMi2015 <- as.data.frame(CJS_AsMi_2015[,c("ch","state")])
AsMi.proc <- process.data(AsMi2015)
AsMi.ddl <- make.design.data(AsMi.proc)
Phi.state <- list(formula = ~ state)
model <- crm(AsMi.proc, AsMi.ddl, model.parameters = list(Phi = Phi.state), accumulate = FALSE)

## Compare several models
fit.models=function()
 {
   Phi.state = list(formula= ~ state)
   # Phi.time = list(formula= ~ time)
   # Phi.statetime = list(formula= ~ time + state)
   Phi.dot = list(formula = ~ 1)
   p.state = list(formula= ~ state)
   p.time = list(formula = ~ time)
   p.statetime = list(formula = ~ time + state)
   p.dot = list(formula= ~ 1)
   cml = create.model.list(c("Phi","p"))
   results=crm.wrapper(cml,data = AsMi.proc, ddl = AsMi.ddl,
                       external=FALSE,accumulate=FALSE)
   return(results)
 }
AsMi.models=fit.models()


model15 <- crm(AsMi2015,
             model.parameters = list(Phi = list(formula = ~ state), 
                                     p = list(formula = ~ time + state)))
model.hessian15 <- cjs.hessian(model15)


model15 <- crm(AsMi2015,
             model.parameters = list(Phi = list(formula = ~ state), 
                                     p = list(formula = ~ state)))
model.hessian15 <- cjs.hessian(model15)

## Dot
model15.dot <- crm(AsMi2015,
             model.parameters = list(Phi = list(formula = ~ 1), 
                                     p = list(formula = ~ 1)))
model.hessian15.dot <- cjs.hessian(model15.dot)

```




```{r}
AsMidf <- cbind(AsMidf, as.matrix(cotyledons[,c(5:13)]), as.matrix(reproduction[,c(5:13)]))
## Added cotyledon and reproduction covariates

## Add td covariate as "trap dependence" like if tagged, will be more likely to be seen again. Detection depends on if the individual was recaptured (trapped) at previous time. Will differ if previously marked or not
td <- splitCH(AsMidf$ch)
td <- td[,1:8] ## A 0/1 value of capture from previous occasion
## Could get td for the second to last. 
releaseocc=process.ch(AsMidf$ch)$first
releaseocc=cbind(1:length(releaseocc),releaseocc)
releaseocc=releaseocc[releaseocc[,2]<nchar(AsMidf$ch[1]),]  ## for all the first, select only ones that weren't the last occasion
td[releaseocc]=0  ## For each row and column, make the first time marked be zero
colnames(td)=paste("td",2:9,sep="")


fit.models=function()
 {
   Phi.state = list(formula= ~ state)
   # Phi.time = list(formula= ~ time)
   # Phi.statetime = list(formula= ~ time + state)
   Phi.dot = list(formula = ~ 1)
   p.state = list(formula= ~ state)
   p.time = list(formula = ~ time)
   p.statetime = list(formula = ~ time + state)
   p.dot = list(formula= ~ 1)
   cml = create.model.list(c("Phi","p"))
   results=crm.wrapper(cml,data = AsMi.proc, ddl = AsMi.ddl,
                       external=FALSE,accumulate=FALSE)
   return(results)
 }
AsMi.models=fit.models()


asmi.proc2 <- process.data(AsMidf)
design.Phi <- list(time.varying = c("Repro"))
design.p <- list(time.varying = c("Repro","td"))
design.parameters = list(Phi = design.Phi, p = design.p)
ddl <- make.design.data(asmi.proc2, parameters = design.parameters)
## Where did age, cohort, fix, and order come from? 
names(ddl$Phi)
names(ddl$p)

## Define models for survival and detection to fit  
Phi.sfw=list(formula = ~ Repro)
p.ast=list(formula = ~ Repro)
model=crm(asmi.proc2,ddl,hessian=TRUE,
          model.parameters=list(Phi=Phi.sfw,p=p.ast))




```

Below we create a range of data values to compute predicted ϕ
 values and then plot the results for Flood and non-Flood years for a range of weights. Not surprising that the slope for weight is nearly 0 because the weight values were generated randomly.  
 ??? Really want to plot along survey times
```{r}
newAsMi <- expand.grid( state = c("Vegetative","Seedling","Reproductive"), 
  Repro1=c(0,1), Repro2=c(0,1),Repro3=c(0,1),Repro4=c(0,1),Repro5=c(0,1),
  Repro6=c(0,1),Repro7=c(0,1),Repro8=c(0,1),Repro9=c(0,1), td2=0,td3=c(0,1),
  td4=c(0,1),td5=c(0,1),td6=c(0,1),td7=c(0,1),td8=c(0,1),td9=c(0,1)) 
reals=predict(model,newdata=newAsMi,se=TRUE)
reals$Phi$Repro=factor(reals$Phi$Repro,labels=c("Vegetative","Reproductive"))
ggplot(reals$Phi,aes(Repro, estimate,ymin=lcl,ymax=ucl))+
  geom_errorbar(width=0.2)+
  geom_point()+
  xlab("\nState")+ylab("Detection\n")

reals$p$Repro=factor(reals$p$Repro,labels=c("Vegetative","Reproductive"))
ggplot(reals$p,aes(Repro, estimate,ymin=lcl,ymax=ucl))+
  geom_errorbar(width=0.2)+
  geom_point()+
  xlab("\nState")+ylab("Detection\n")


```
 
Using MCMC CJS model (probitCJS) similar to MLE.   
```{r}
# Add reproduction covariate
head(AsMidf)
design.parameters <- list(Phi = list(time.varying = "Repro"))
model.parameters = list(Phi=list(formula = ~ Repro),
                        p = list(formula = ~ time))
MCMCfit <- crm(AsMidf, model = "probitCJS",
               model.parameters = model.parameters,
               design.parameters = design.parameters,
               burnin=1000, iter=5000)

MCMCfit$results$reals

MCMCfit$results$reals$p %>%
  ggplot(  aes(time, mean,ymin=CI.lower,ymax=CI.upper))+
  geom_errorbar()+
  geom_point()+
  ylab("Detection")

```



Format of covariates   
```{r}
data(dipper)
# Add a dummy weight field which are random values from 1 to 10
set.seed(123)
dipper$weight=round(runif(nrow(dipper),0,9),0)+1  ## Static, could this be longest length in AsMi? 
# Add Flood covariate
Flood=matrix(rep(c(0,1,1,0,0,0),each=nrow(dipper)),ncol=6)  ## Matrix of individuals X (capture events - 1) 
# If td was added to the time-varying argument for ϕ and Flood was added for p, then we would have also needed to add td1 and Flood7 due to the different time labels for ϕ and p. 
colnames(Flood)=paste("Flood",1:6,sep="")
dipper=cbind(dipper,Flood)
# Add td covariate, but exclude first release as a capture
# splitCH and process.ch are functions in the marked package
td=splitCH(dipper$ch)  # To split character string vector of capture histories into a matrix, append to original dataset
td=td[,1:6]
releaseocc=process.ch(dipper$ch)$first
releaseocc=cbind(1:length(releaseocc),releaseocc)
releaseocc=releaseocc[releaseocc[,2]<nchar(dipper$ch[1]),]
td[releaseocc]=0
colnames(td)=paste("td",2:7,sep="")
dipper=cbind(dipper,td)
> # show names
names(dipper)
```


Among years    

 Compare variable and constant detection, variable and constant survival, differences among fenced and unfenced
```{r}

## ch_asmi
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/capturehistoryannual.Rdata")

## fence_matrix 
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/fencematrix.Rdata")


```



# RMark   
   1. Mark individuals that had a tag prior to 2013, those that are newly tagged. 




Recruitment with Jolly-Seber  <https://sites.warnercnr.colostate.edu/gwhite/recruitment-parameters-jolly-seber-models/>     
phi = survival rate; p = probability of capture     
Pradel: 
        gamma = probablity that individual at time i had not entered population between i and i-1. gamma(i+1) = 1- B(i)/N(i+1)   
        Lambda = rate of change; lambda(i) = N(i+1)/N(i) (with log link function)   
        f = fecundity, f(i) = number of new in population at time i per number in the population at time i-1; N(i+1) = N(i)f(i) + N(i)phi(i)       
        
Schwarz and Arnason (1996):     
        Super population (N) and probability of entry (pent in MARK, beta in their paper)    
        
Link-Barker (2003):     
        

```{r}

```

