---
title: "AsMi confidence intervals from point estimates"
author: "Michelle DePrenger-Levin"
date: "2023-08-15"
output: html_document
---

```{r}
library(popbio)
library(popdemo)
library(dplyr)
library(tidyr)
library(ggplot2)
library(popbio)
library(devtools)
library(sciplot)
library(plotrix)
library(scales)
library(digest)
library(RCurl)
library(prism)
# prism_set_dl_dir("Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/PRISM_asmiclimate")
# library(rgdal) Will go away in 2023 https://r-spatial.org/r/2022/04/12/evolution.html and https://github.com/r-spatial/evolution rgdal: version: 1.6-2, (SVN revision 1183)

library(raster)
library(Rmisc)

library(car) 

# AIC
require(AICcmodavg)
library(lme4)
library(lattice)

#Plots and figures
library(DiagrammeR)
library(patchwork)

library(MASS)
library(HDInterval)

```


Load MPM from Annual report tables and figures.Rmd  

```{r}
## all.pva
load("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_AnnualReports/2023_Astragalus-microcymbus_AnnualReport/allpva2023-11-12.Rdata")

load("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/stage-fate2023.Rdata")

```

Jensen's inequality via Fiske et al 2008 (via Fox and Kendall)   
Small sample sizes will lead to biased estimates of lambda as a result of increased sampling variance. 
```{r}
#binomial distribution means
p <- seq(0.01,0.99, 0.01)
sampleSize <- seq(1,100,10)
varBinom <- (p*(1-p))/sampleSize[4]

plot(p, varBinom, type = "l")

binomDF <- do.call(rbind,lapply(sampleSize[-1], function(x){
  # pHetero <- lapply(p, function(i) { rbeta(1,round(i*x,0), round((1-i)*x,0))}) ## This would still be homogenous
  varHetero <- unlist(lapply(p, function(i) var(unlist(lapply(1:x, function(rep) rbeta(1,round(i*x,0), round((1-i)*x,0)) ) ))))
  
  data.frame(sampleSize = x, p, varBinom = (p*(1-p))/x,
             varHetero)
}))

binomDF %>%
  filter(sampleSize < 200) %>%
ggplot(  aes(p, varBinom, color = as.factor(sampleSize)))+
  geom_point() +
  stat_smooth(aes(p, varHetero))+
  theme_bw()


lams <- c(0.3,rep(1.1, 9))
#arithmetic mean
mean(lams) # grows 2%/year
#geometric mean
prod(lams)^(1/10)
exp(mean(log(lams)))

mean(log(lams)) # geometric mean less than 
log(mean(lams)) # arithmetic mean
```



```{r}

## Which plots go with which sites
A_site1 <- mean(do.call(c,list(all.pva$"1", all.pva$"22", all.pva$"32", all.pva$"42", all.pva$"52")))
A_site2 <- mean(do.call(c,list(all.pva$"16", all.pva$"58", all.pva$"89", all.pva$"90", all.pva$"96", all.pva$"98")))
A_site5 <- mean(do.call(c,list(all.pva$"606", all.pva$"607", all.pva$"608", all.pva$"609", all.pva$"610")))
A_site15 <- mean(do.call(c,list(all.pva$"8", all.pva$"578", all.pva$"581", all.pva$"799")))
A_site19 <- mean(do.call(c,list(all.pva$"300", all.pva$"512", all.pva$"513", all.pva$"514", all.pva$"515")))
A_site26 <- mean(do.call(c,list(all.pva$"238", all.pva$"480", all.pva$"598", all.pva$"611", all.pva$"614")))

do.call(rbind,lapply(list(A_site1,A_site2,A_site5,A_site15,A_site19,A_site26), function(x) lambda(x)))
## Expected lambdas per site


## Per year


site1.mpm <- do.call(list,list(all.pva$"1", all.pva$"22", all.pva$"32", all.pva$"42", all.pva$"52"))
site1year <- lapply(2015:2021, function(yr){
  mpmyr <- lapply(site1.mpm, function(m) {
    m[grepl(as.character(yr), names(m))]
    })
  mean(lapply(mpmyr, '[[', 1))
})

site2.mpm <- list(all.pva$"16", all.pva$"58", all.pva$"89", all.pva$"90", all.pva$"96", all.pva$"98")  ## 89 removed
site2year <- lapply(2015:2020, function(yr){
  mpmyr <- lapply(site2.mpm, function(m) {
    m[grepl(as.character(yr), names(m))]
    })
  mean(lapply(mpmyr, '[[', 1))
})

26*26
## by year
lapply(all.pva, function(ap) names(ap))
names(do.call(c,all.pva))

allpva <- do.call(c,all.pva)
allpva[1:2]

peryear <- lapply(1996:2022, function(x) {
  allpva[grepl(x, names(allpva))]
})

mean(do.call(list,peryear[[1]]))

x<-peryear[[3]]
lapply(peryear, function(x){ 
  yr <- do.call(list,x)
  lambda(mean(yr))
  })

all.pvamat <- matrix(unlist(all.pva), ncol = length(do.call(c,all.pva)))
all.pvamat[,1:4]

  s <- sqrt(dim(all.pvamat)[1]) ## number of stage classes
  n <- dim(all.pvamat)[2] ## number of matrixes
  Abar <- numeric(s^2) ## an s X s matrix and/or array of all zeros ??
  Exy <- numeric(s^4)
  
  C <- kronecker(all.pvamat[,1], all.pvamat[,1]) * n / (n-1)
  # C <- (Exy - kronecker(Abar, Abar)) * n / (n - 1)
  C <- matrix(C, nrow = s^2)
  Abar <- matrix(all.pvamat[,1], nrow = s)
  
  V <- Conj(solve(W, tol = 1e-100))
  
  v <- abs(Re(V[lmax, ]))
  S <- v %o% w 
  
  ## code from eigen.analysis for lambda and Sensitivity matrix
  ev <- eigen(Abar)
  lmax <- which(Re(ev$values) == max(Re(ev$values)))
  lambda <- Re(ev$values[lmax])
  W <- ev$vectors
  w <- abs(Re(W[, lmax]))
  

```


error is variance^2 = e_i * T(sum e_i)

Morris and Doak Stochastic log growth rate by simulation   
```{r}
library(Rage)
library(popbio)
# 1. Estimate the mean and variance of vital rates, growth, survival, and retrogression

matU1 <- all.pva[[1]][[1]]
mats1 <- mpm_split(matU1)
vitrat <- vital_rates(mats1$matU, mats1$matF, mats1$matC, splitStages = "ontogeny", weights = "SSD",
                      matrixStages = c("active","active","active","dorm"))

vr_growth(mats1$matU)

## mean for each vital rate (not matrix element) for each MPM
meanvitalrates <- do.call(rbind,lapply(all.pva, function(m){
  out <- do.call(rbind,lapply(m, function(x){
    mats <- mpm_split(x)
    vitrat <- vital_rates(mats$matU, mats$matF, mats$matC, splitStages = "ontogeny", weights = "SSD")
    data.frame(JuvSurv = vitrat$survJuv, JuvRetr = vitrat$retrJuv, JuvGrowth = vitrat$progJuv,
               AduSurv = vitrat$survAdu, AduRetr = vitrat$retrAdu)
  }))
  out
  }))

meanvitalrates <- meanvitalrates %>%
  mutate(PlotYear = row.names(.)) %>%
  separate(PlotYear, into = c("Plot","Year"))%>%
  mutate(Year = as.numeric(Year))

MeanVitRat <- meanvitalrates %>%
  group_by(Year) %>%
  dplyr::summarise(across(JuvSurv:AduRetr, mean))

VarVitRat <- meanvitalrates %>%
  group_by(Year) %>%
  dplyr::summarise(across(JuvSurv:AduRetr, var))

# 2. Correct for sampling variation  Kendall()
?Kendall
```



Wisdom Mills and Doak 2000 Life stage simulation analysis: estimating vital-rate effects on population growth for conservation   
vital-rate uncertainty, variation, and covariation cannot be ignored   
vital rates drawn randomly from prob distribution - reflects spatial and covariation. 
randomly selected vital rates used for time-invariant matrix population model   
Sample many matrices, distribution of all vital rates   
Not time varying like Taljapurkar   
does not examine time varying effects, must be calculated at stable stage distribution (like LTREs)
1. selected reproductive rates from log-normal, selected all other rates from beta-prob distribution adhering to prespecificed variance-covariance structure among vital rates

```{r}

## Population size per year and site
popsz.yearsite <- asmi.raw %>%
  group_by(year, AsMi_site_id) %>%
  filter(length > 0) %>%
  dplyr::summarise(N = n()) %>%
  arrange(year, AsMi_site_id) %>%
  unite(Site.Year, c(AsMi_site_id,year), sep = ".")


mpmdim <- dim(all.pva[[1]][[1]])[1]

## Gather all mpms into one large matrix, columns are a matrix
all.mpm <- do.call(c, all.pva)
names.mpm <- do.call(c,lapply(all.pva, function(x){
  lapply(x, function(y) names(y))
}))
matrices <- matrix(unlist(all.mpm), ncol = length(all.mpm))
named.mpms <- data.frame(matrices)
names(named.mpms) <- names(names.mpm)

A <- matrix(named.mpms[,1],nrow = mpmdim)

named.mpms[which(colSums(matrices)==0)] ## ??? Why?? 

asmi.all2 %>%
  filter(plot %in% c(581, 609)) %>%
  group_by(year, plot, site) %>%
  dplyr::summarise(n = n()) %>%
ggplot( aes(year, n, color = as.factor(plot)))+
  geom_point()+
  geom_line()

asmi.all2 %>%
  filter(plot %in% c(581, 609)) %>%
  group_by(year, plot, site) %>%
  dplyr::summarise(n = n()) %>%
  print(n=43)

# named.mpms[1]  ## 1.2015
popsz <- popsz.yearsite$N[which(grepl(names(named.mpms)[1],popsz.yearsite$Site.Year))]

# popsz <- 100
ssd <- stable.stage(A)
n0 <- round(ssd*popsz,0)

maxt <- 100
n <- c(NA)
## Projection matrix
nt <- round(A * n0,0)

Px <- seq(0.01,0.99, by = 0.01)
likdis <- lapply(1:4, function(i){
    dbinom(x = nt[i],size = n0[i], Px)
    })  ## Dormant always survive

s <- 4 ## number of classes
n <- dim(matrices)[2]
prob <- rep(1 / n, n)
## sample survival rates from distribution Here's where I need a vcov matrix to select among the vital rates that result in distribution
## code from eigen.analysis for lambda and Sensitivity matrix
  Abar <- numeric(s^2) ## an s X 1 array of all zeros 
  Exy <- numeric(s^4)

  ## for each matrix, add values to Abar and Exy weighted by probabilities in prob
  for (i in 1:n) {
    A <- matrices[, i]
    Exy <- Exy + prob[i] * kronecker(A, A)
    Abar <- Abar + prob[i] * A
  }
  ## Covariance matrix
  C <- (Exy - kronecker(Abar, Abar)) * n / (n - 1)
  C <- matrix(C, nrow = s^2)
  Abar <- matrix(Abar, nrow = s)  
  ev <- eigen(Abar)
  lmax <- which(Re(ev$values) == max(Re(ev$values)))  ## which are the real numbers??
  lambda <- Re(ev$values[lmax])
  # lambda <- lambda(Abar)  ## same
  W <- ev$vectors
  w <- abs(Re(W[, lmax]))
  
  ## Causes problems when... tol too high
  V <- Conj(solve(W, tol = 1e-30))
  
  
  v <- abs(Re(V[lmax, ]))  ## reproduction, multiple left and right eigen
  S <- v %o% w  # is scalar quantity, the sum of the product of the values, transposes the second 
 Svec <- as.vector(S)
 tau2 <-  t(Svec) %*% C %*% Svec  
 loglam <- log(lambda) - tau2/(2*lambda^2)
 lams <- exp(loglam)
 
 
## tau2/lambdahat2 is structured equivalent of sigma2 for unstructured
 varmpm <- tau2/(lambda^2)

 
plot(Px, likdis[[1]], type = "l", ylim = c(0,1))
lines(Px, likdis[[2]], col = "red")
lines(Px, likdis[[3]], col="blue")
lines(Px, likdis[[4]], col="green")

# CI95 <- lapply( which(n0 >0), function(i){ 
#     prop.test(x = nt[i],n = n0[i], conf.level = 0.95)
#     }) 

library(Hmisc)
CI95 <- do.call(rbind,lapply( which(n0 >0), function(i){ 
    binconf(x = nt[i],n = n0[i], alpha = 0.05)
    }) )

abline(v = CI95[,1], col = c("red","blue","green"))
abline(v = CI95[,2], col = c("red","blue","green"), lty = "dotted")
abline(v = CI95[,3], col = c("red","blue","green"), lty = "dotted")



```

Morris and Doak 2002, pg 237. The quantity tau^2/lambda^2 (of the mean matrix) approximates the temporal (or spatial) variance of the log population growth rate caused by environmental stochasticity    
```{r}

all.mpm <- do.call(c, all.pva)
names.mpm <- do.call(c,lapply(all.pva, function(x){
  lapply(x, function(y) names(y))
}))
matrices <- matrix(unlist(all.mpm), ncol = length(all.mpm))
named.mpms <- data.frame(matrices)
names(named.mpms) <- names(names.mpm)

lambdatau <- function(mats){
  
  # matrices <- matrix(unlist(mats), ncol = length(all.mpm))
  matrices <- mats
  s <- 4 ## number of classes
  n <- dim(matrices)[2]
  prob <- rep(1 / n, n)
## sample survival rates from distribution Here's where I need a vcov matrix to select among the vital rates that result in distribution
## code from eigen.analysis for lambda and Sensitivity matrix
  Abar <- numeric(s^2) ## an s X 1 array of all zeros 
  Exy <- numeric(s^4)

  ## for each matrix, add values to Abar and Exy weighted by probabilities in prob
  for (i in 1:n) {
    A <- matrices[, i]
    Exy <- Exy + prob[i] * kronecker(A, A)
    Abar <- Abar + prob[i] * A
  }
  ## Covariance matrix
  C <- (Exy - kronecker(Abar, Abar)) * n / (n - 1)
  C <- matrix(C, nrow = s^2)
  Abar <- matrix(Abar, nrow = s)  
  ev <- eigen(Abar)
  lmax <- which(Re(ev$values) == max(Re(ev$values)))  ## which are the real numbers??
  lambda <- Re(ev$values[lmax])
  # lambda <- lambda(Abar)  ## same
  W <- ev$vectors
  w <- abs(Re(W[, lmax]))
  
  ## Causes problems when... tol too low
  V <- Conj(solve(W, tol = 1e-300))
  v <- abs(Re(V[lmax, ]))  ## reproduction, multiple left and right eigen
  S <- v %o% w  # is scalar quantity, the sum of the product of the values, transposes the second 
  Svec <- as.vector(S)
  tau2 <-  t(Svec) %*% C %*% Svec  
  loglam <- log(lambda) - tau2/(2*lambda^2)
  lams <- exp(loglam)
  
  ## tau2/lambdahat2 is structured equivalent of sigma2 for unstructured
  varmpm <- tau2/(lambda^2)
  out <- data.frame(lambdabar = lambda, tau2 = tau2, varmpm = varmpm)
  return(out)
}

## remove the first and last years, year indicates the start of the interval 
year.var <- do.call(rbind,lapply(1996:2021, function(yr){
  data.frame(Year = yr, lambdatau(named.mpms[,which(grepl(yr,names(named.mpms)))]))
}))


```

From Astragalus-microcymbus_capturehistories   
```{r}
# currentyr <- as.numeric(format(as.Date(Sys.Date(),format="%Y-%m-%d"), "%Y"))
currentyr <- 2023

##To change if someone other than Michelle is running code
userpath <- "C:/Users/DePrengm/"

rawdatapath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/", currentyr, "_Astragalus-microcymbus_RawData.csv", collapse = '', sep = '')

asmi.raw <- read.csv(rawdatapath, na.strings = "na")

## plot 89 at site 1 was lost in 2020
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 89 & asmi.raw$year > 2019),]

# Site 26 had two plots added later; Plots 238 and 300 were added in 1996 with no seedlings! good!;  plot 598 was added in 2004, no seedlings, good!  
table(asmi.raw$AsMi_plot_id[asmi.raw$AsMi_plot_id %in% c(238,300,598)], 
      asmi.raw$status[asmi.raw$AsMi_plot_id %in% c(238,300,598)],
      asmi.raw$year[asmi.raw$AsMi_plot_id %in% c(238,300,598)])

# Need to change all the seedling in 2014 for sites 1 and 2 to appropriate vegetative or reproductive 
asmi.raw$status[asmi.raw$AsMi_site_id<3 & asmi.raw$year == 2014 & asmi.raw$flower == 0] <- "vegetative"
asmi.raw$status[asmi.raw$AsMi_site_id<3 & asmi.raw$year == 2014 & asmi.raw$flower == 1] <- "reproductive"

# No longer adding climate data to the database
asmi.raw <- asmi.raw[,grep(paste(c("Temp","Rain","Snow","Aug.Jul"),collapse="|"), names(asmi.raw),
                           value = TRUE, invert = TRUE)]
asmi.raw$Browsing...Status[asmi.raw$Browsing...Status == "mammal" ] <- "Mammal"
asmi.raw[asmi.raw$Browsing...Status ==""& !is.na(asmi.raw$Browsing...Status),]
asmi.raw[is.na(asmi.raw$AsMi_data_id),]
asmi.raw <- asmi.raw[!is.na(asmi.raw$AsMi_data_id),] # remove any rows with NA for data_id

# Which plot # go with which site
plotsXsite <- as.data.frame(table(asmi.raw$AsMi_plot_id,asmi.raw$AsMi_site_id))
plotsXsite[plotsXsite$Freq > 0,]

# one blank row at the end of the csv download, gone by 2021
asmi.raw[asmi.raw$status == "",]
# asmi.raw <- asmi.raw[asmi.raw$status != "",]
asmi.raw$status <- factor(asmi.raw$status)
table(asmi.raw$status)

#reset factors of browsing to eliminate " "
# asmi.raw$Browsing...Status[asmi.raw$Browsing...Status == "mammal"] <- "Mammal"
asmi.raw$Browsing...Status <- factor(asmi.raw$Browsing...Status)

#reset factors for fence
asmi.raw$fence <- factor(asmi.raw$fence)
table(asmi.raw$fence)

asmi.raw[asmi.raw$length>90 & !is.na(asmi.raw$length),] 
asmi.raw$length[asmi.raw$length == 921] <- 21

# If there are fruit, it flowered and needs to be reproductive, not vegetative
asmi.raw[asmi.raw$status=="vegetative" & asmi.raw$fruit>0,] # 5 times in 2020
wrongAsMidataid <- asmi.raw$AsMi_data_id[asmi.raw$status=="vegetative" & asmi.raw$fruit>0]
asmi.raw$flower[asmi.raw$AsMi_data_id %in% wrongAsMidataid] <- 1
asmi.raw$status[asmi.raw$AsMi_data_id %in% wrongAsMidataid] <- "reproductive"
asmi.raw[asmi.raw$AsMi_data_id %in% wrongAsMidataid,]


str(asmi.raw) ## n = 1, y = 2

CJS_asmi <- asmi.raw %>%
  dplyr::mutate(fence = as.numeric(fence)-1) %>%
  dplyr::mutate(fence = if_else(year < 2006, 0, fence)) %>%
  dplyr::mutate(fence = if_else(year == 2006 & AsMi_plot_id %in% c(300, 512), 0, fence)) %>%
  dplyr::mutate(fence = if_else(year > 2015, 0, fence)) %>%
  tidyr::complete(AsMi_tag_id, year, fill = list(length = 0, fence = 0)) %>%
  group_by(AsMi_site_id, AsMi_plot_id, AsMi_tag_id) %>%
  dplyr::summarise(first = min(year),
            dh = if_else(length > 0, 1, 0),
            year = year,
            status = status,
            fence = fence,
            length = length
            ) %>%
  ungroup() %>%
  tidyr::complete(nesting(AsMi_tag_id,AsMi_site_id,AsMi_plot_id,first), year = min(year):max(year), 
           fill = list(dh = 0, status = "dead", fence = 0),
           explicit = FALSE) %>%
  filter(!is.na(AsMi_site_id))  ## Work around but shouldn't be adding a new first for each

CJS_asmi %>%
  filter(!is.na(length))

ch_asmi <- CJS_asmi %>%
  group_by(AsMi_tag_id, AsMi_site_id, AsMi_plot_id, first) %>%
  dplyr::mutate(first.ln = length[year == first]) %>%
  dplyr::arrange(year, .by_group = TRUE) %>%
  dplyr::summarise(ch = paste0(dh, collapse = "")) %>%
  mutate(freq = "1;")

fence_matrix <- CJS_asmi %>%
  dplyr::select(c(AsMi_tag_id:year, fence)) %>%
  group_by(AsMi_tag_id, AsMi_site_id, AsMi_plot_id, first) %>%
  dplyr::arrange(year, .by_group = TRUE) %>%
  pivot_wider(names_from = year, values_from = fence)
  
identical(ch_asmi$AsMi_tag_id, fence_matrix$AsMi_tag_id)

## Cooch and White Chapter 2 MARK
## write.table(ch_asmi, file = "", sep = "    ", quote=FALSE, col.names=FALSE, row.names=FALSE)
## Filter out capture histories that are all 0s
ch_asmi %>%
  ungroup() %>%
  dplyr::select(ch:freq) %>%
  filter(grepl("1", ch)) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/channual.inp", 
              sep = "    ", quote=FALSE, col.names=FALSE, row.names=FALSE)


save(fence_matrix, file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/fencematrix.Rdata")

## Multistate
CJS_asmi <- asmi.raw %>%
  mutate(status == as.character(status)) %>%
  dplyr::mutate(status = if_else((status == "seedling" & flower == 1), "reproductive", as.character(status))) %>%
  mutate(fence = as.numeric(fence)-1) %>%
  mutate(fence = if_else(year < 2006, 0, fence)) %>%
  mutate(fence = if_else(year == 2006 & AsMi_plot_id %in% c(300, 512), 0, fence)) %>%
  mutate(fence = if_else(year > 2015, 0, fence)) %>%
  tidyr::complete(AsMi_tag_id, year, fill = list(length = 0, fence = 0)) %>%
  group_by(AsMi_site_id, AsMi_plot_id, AsMi_tag_id) %>%
  dplyr::summarise(first = min(year),
            dh = if_else(length > 0, 1, 0),
            year = year,
            status = status,
            fence = fence,
            length = length
            ) %>%
  ungroup() %>%
  tidyr::complete(nesting(AsMi_tag_id,AsMi_site_id,AsMi_plot_id,first), year = min(year):max(year), 
           fill = list(dh = 0, status = "dead", fence = 0),
           explicit = FALSE) %>%
  filter(!is.na(AsMi_site_id))  ## Work around but shouldn't be adding a new first for each
## Multistate just use S,V,R for the states (or just V, R, 0 for non-detection)
ch_asmiMS <- CJS_asmi %>%
  mutate(dh = case_when(status == "seedling" ~ "S", 
                        status == "vegetative" ~ "V",
                        status == "reproductive" ~ "R",
                        status %in% c("dormant","dead") ~ "0")) %>%
  group_by(AsMi_tag_id, AsMi_site_id, AsMi_plot_id, first) %>%
  dplyr::mutate(first.ln = rep(length[year == first], length.out = n()))  %>%
  ungroup() %>%
  group_by(AsMi_tag_id, AsMi_site_id, AsMi_plot_id,first,first.ln) %>%
  dplyr::arrange(year, .by_group = TRUE) %>%  ## fine, they're in the correct order
  dplyr::summarise(ch = paste0(dh, collapse = "")) %>%
  ungroup() %>%
  mutate(freq = "1;")

## Multi-state 
ch_asmiMS %>%
  ungroup() %>%
  dplyr::select(ch:freq) %>%
  filter(grepl("S", ch)|grepl("V", ch)|grepl("R",ch)) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/chMS.inp", 
              sep = "    ", quote=FALSE, col.names=FALSE, row.names=FALSE)

## Multi-state by Site
ch_asmiMS %>%
  ungroup() %>%
  dplyr::select(AsMi_site_id:freq) %>%
  filter(grepl("S", ch)|grepl("V", ch)|grepl("R",ch)) %>%
  dplyr::rename(Site = AsMi_site_id) %>%
  dplyr::select(c(Site,ch,freq)) %>%
  dplyr::mutate(Site = as.factor(Site)) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/chMSXplot.inp", 
              sep = "    ", quote=FALSE, col.names=FALSE, row.names=FALSE)



ch_asmiMS <- CJS_asmi %>%
  group_by(AsMi_tag_id, AsMi_site_id, AsMi_plot_id, first) %>%
  dplyr::mutate(first.ln = length[year == first]) %>%
  dplyr::arrange(year, .by_group = TRUE) %>%
  dplyr::summarise(ch = paste0(dh, collapse = "")) %>%
  mutate(freq = "1;")

```

Pradel model for lambda (population growth rate) and recruitment (phi) and detection (p)
```{r}

# Years of study in the abstract
length(1995:currentyr)
asmi29 <- convert.inp("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/channual.inp")

asmi29.proc <- process.data(asmi29, model = "Pradel")

## Gamma p 
mod.dot <- mark(asmi29.proc, model.parameters = list(Gamma = list(formula = ~1),
                                                     p = list(formula = ~1)))

## Phi p f
asmi29.proc2 <- process.data(asmi29, model = "Pradrec")
mod.dot2 <- mark(asmi29.proc2, model.parameters = list(Phi = list(formula = ~1),
                                                       p = list(formula = ~1)))
asmi29.ddl <- make.design.data(asmi29.proc2)

mod.time <- mark(asmi29.proc2, model.parameters = list(Phi = list(formula = ~ time),
                                                       p = list(formula = ~ 1)))

mod.time$pims
mod.time$results$real
### LAMBDA!!!
mod.time$results$derived

asmi.lambda <- mod.time$results$derived$`Lambda Population Change`

ggplot(asmi.lambda, aes(1995:2022, estimate)) +
  geom_point()+
  # geom_line() +
  theme_bw() +
  geom_errorbar(aes(1995:2022, ymin = lcl, ymax = ucl),
                color = "red")+
  xlab("Year") +
  ylab(expression(lambda))+
  geom_hline(yintercept = 1, linetype = "dotted")
  # geom_ribbon(aes(ymin = lcl, ymax = ucl), fill = "red", alpha=0.25)

asmi.loglambda <- mod.time$results$derived$`log(Lambda) Population Change`

ggplot(asmi.loglambda, aes(1995:2022, estimate)) +
  geom_point()+
  # geom_line() +
  theme_bw() +
  geom_errorbar(aes(1995:2022, ymin = lcl, ymax = ucl),
                color = "red")+
  xlab("Year") +
  ylab(expression("log"~lambda))+
  geom_hline(yintercept = 0, linetype = "dotted")


## More likely recruitment changes over time (f) as well as survival 
mod.timetime <- mark(asmi29.proc2, model.parameters = list(Phi = list(formula = ~ time),
                                                           p = list(formula = ~ 1),
                                                           f = list(formula = ~ time)))

asmi.lambdatimetime <- mod.timetime$results$derived$`Lambda Population Change`
asmi.lambdatimetime2 <- asmi.lambdatimetime %>%
  mutate(Year = 1995:2022) %>%
  relocate(Year, .before = estimate) %>%
  mutate(met = "CJS")

year.var %>%
  dplyr::rename(estimate = lambdabar) %>%
  dplyr::rename(se = tau2) %>%
  dplyr::mutate(lcl = estimate - varmpm,
                ucl = estimate + varmpm) %>%
  dplyr::select(c(Year:se,lcl:ucl)) %>%
  mutate(met = "MPM") %>%
  bind_rows(asmi.lambdatimetime2) %>%
ggplot(   aes(Year, estimate, color = met)) +
  geom_point(position = position_dodge(width = 0.5))+
  # geom_line() +
  theme_bw() +
  geom_errorbar(position = position_dodge(width = 0.5), aes(ymin = lcl, ymax = ucl))+
  xlab("Year") +
  ylab(expression(lambda))+
  geom_hline(yintercept = 1, linetype = "dotted") 
 


asmi.loglambdatimetime <- mod.timetime$results$derived$`log(Lambda) Population Change`
asmi.loglambdatimetime2 <- asmi.loglambdatimetime %>%
  mutate(Year = 1995:2022) %>%
  relocate(Year, .before = estimate) %>%
  mutate(met = "CJS")
plotdata <- year.var %>%
  dplyr::mutate(lambdabar = log(lambdabar)) %>%
  dplyr::rename(estimate = lambdabar) %>%
  dplyr::rename(se = tau2) %>%
  dplyr::mutate(lcl = estimate - varmpm,
                ucl = estimate + varmpm) %>%
  dplyr::select(c(Year:se,lcl:ucl)) %>%
  mutate(met = "MPM") %>%
  bind_rows(asmi.loglambdatimetime2) 

p1 <- ggplot(plotdata, aes(Year, estimate, color = met)) +
  geom_point(position = position_dodge(width = 0.5))+
  # geom_line() +
  theme_bw() +
  geom_errorbar(position = position_dodge(width = 0.5), aes(ymin = lcl, ymax = ucl))+
  xlab("Year") +
  ylab(expression(log~lambda))+
  scale_color_manual("Method", values = c("black","grey70"))+
  geom_hline(yintercept = 0, linetype = "dotted")


ggsave(p1, filename = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/Chapter AsMi/Images/lambdasCJSMPM.jpg",
         width=250, height=125,units='mm', dpi=300 )

```



## Multistate just use S,V,R for the states (or just V, R, 0 for non-detection)
```{r}

groupDATAFRAME <- asmi.raw %>%
  group_by(AsMi_site_id, AsMi_plot_id) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::select(AsMi_site_id:AsMi_plot_id) %>%
  dplyr::rename(Site = AsMi_site_id) %>%
  dplyr::rename(Plot = AsMi_plot_id) %>%
  as.data.frame()

groupSitedf <- data.frame(Site = unique(asmi.raw$AsMi_site_id))

# asmiMSplot <- convert.inp("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/chMSXplot.inp", 
#                       # group.df = groupSitedf,
#                       covariates = c("Site"))


asmiMS <- convert.inp("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/chMS.inp")
asmiMSproc <- process.data(asmiMS, model = "Multistrata")
asmiMSddl <- make.design.data(asmiMSproc)
mod.dotMS <- mark(asmiMSproc, asmiMSddl)

mod.dotMS$results$real
summary(mod.dotMS, show.fixed = TRUE)
mod.dotMS$model
mod.dotMS$parameters  ## S, p, Psi
mod.dotMS$results$real
```


## Robust design
# 3 primary occasions (2013-2015) and 9 primary occasions within, just skipped "." a bunch for 2013, skipped some for others
### time.intervals






################## Failed
Devenish Nelson et al 2010 Conf Int     
```{r}

##Correlation structure of growth, survival, reproduction, retrogression following Doak et al 1994 
# survival 
head(asmi.all2)
asmi.all2 %>%
  mutate(surv = ifelse(fate == "dead",0,1)) %>%
  group_by(stage, surv) %>%
  dplyr::summarise(n = n()) %>%
  group_by(stage) %>%
  dplyr::summarise(n[surv == 1]/sum(n))


## Point estimate of lambda as dominant eigenvalue
lapply(all.pva, function(x){ 
  lapply(x, function(y){ 
    lambda(y)
    })
  })

## Resample, parametric bootstrap, from 10,000 replicate projection matrices each element drawn from corresponding likelihood distribution
## Did not use right shifted Poisson because a reproductive adult could have all individuals fail??
# split into birth and death matrices
U <- all.pva$'1'[[1]]
## Set population size and numbers 
N <- 100
start <- stable.stage(U)
n <- matrix(rep(0,nrow(U)), nrow = 1, byrow = TRUE)
n <- round(N*start,0) 
Ubinom <- mapply(as.vector(U[-1,]), n, function(x,y) dbinom())
dbinom()

boot.lambdas <- lapply(all.pva, function(x){
  lapply(x, function(y){
    
  })
})




```


