---
title: "MPM_lambda_survival_transitions"
author: "Michelle DePrenger-Levin"
date: "2023-12-02"
output: html_document
---


```{r}

rm(list=ls())

library(plyr)
library(dplyr)
library(popbio)
library(devtools)
library(ggplot2)
library(RMark)
```


MPM   
```{r}

## Get 2 stage pva to make MPMs
source_url("https://raw.githubusercontent.com/DenverBotanicGardens/PVA/master/StagePVA2stages_dormancyoptional.R")

load("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/stage-fate2023.Rdata")

## StagePVA2stages_dormancyoptional.R but dormancy isn't optional yet :)
MPM.plots <- StagePVA(asmi.all2)

SitePlot <- asmi.all2 %>%
  distinct(plot,site) %>%
  mutate(plot = as.factor(plot))%>%
  mutate(site = as.factor(site))

MPMlam <- data.frame(Lambda = do.call(c,lapply(MPM.plots, function(x) do.call(c, lapply(x, function(y) lambda(y))))))

MPMlamdf <- MPMlam %>%
  mutate(PlotYear = row.names(MPMlam)) %>%
  tidyr::separate(PlotYear, into = c("Plot","Year"))%>%
  dplyr::mutate(Plot = as.factor(Plot)) %>%
  dplyr::left_join(SitePlot, by = c("Plot" = "plot")) %>%
  filter(!is.na(Lambda)) %>%
  arrange(site,Plot,Year)

p1 <- MPMlamdf %>%
  filter(site %in% c(5,15,19,26)) %>%
  filter(Lambda != 0) %>%
  ggplot(  aes(Plot, log(Lambda)))+
  geom_boxplot()+
  geom_jitter(height = 0, width = 0.2)+
  theme_bw()+
  facet_wrap(~site, scales = "free_x")+
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey40")+
  ylab(expression(log~lambda))

ggsave(p1, filename = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/Chapter AsMi/Images/lambdas4sitesPlots.jpg",
         width=125, height=100,units='mm', dpi=300 )

### Tuljapukar's log stochastic lambda by plots in a site across site over all years, need to change 
  # V <- Conj(solve(W, tol = 1e-300))  add tolerance to make it work 
# For Pradel comparison, exclude sites 1 and 2, and three plots that are added 1996 and 2004
site1.mpm <- do.call(list,list(MPM.plots$"1", MPM.plots$"22", MPM.plots$"32", 
                               MPM.plots$"42", MPM.plots$"52"))
MPMsensitiviessite1 <- lapply(2015:2021, function(yr){
  mpmyr <- do.call(c,lapply(site1.mpm, function(m) {
    m[grepl(as.character(yr), names(m))]
    }) )

  ms1 <- mean(mpmyr)
})

lapply(MPMsensitiviessite1, function(x) sensitivity(x))
stoch1 <- stoch.growth.rate(MPMsensitiviessite1)

site2.mpm <- list(MPM.plots$"16", MPM.plots$"58", MPM.plots$"89", MPM.plots$"90", MPM.plots$"96", MPM.plots$"98")  ## 89 removed in 2020
MPMsensitiviessite2 <- lapply(2015:2021, function(yr){
  mpmyr <- do.call(c,lapply(site2.mpm, function(m) {
    m[grepl(as.character(yr), names(m))]
    }) )

  ms1 <- mean(mpmyr)
})
stoch2 <- stoch.growth.rate(MPMsensitiviessite2)


site5.mpm <- list(MPM.plots$"606", MPM.plots$"607", MPM.plots$"608", MPM.plots$"609", MPM.plots$"610")  
MPMsensitiviessite5 <- lapply(1996:2021, function(yr){
  mpmyr <- do.call(c,lapply(site5.mpm, function(m) {
    m[grepl(as.character(yr), names(m))]
    }) )

  ms1 <- mean(mpmyr)
})
stoch5 <- stoch.growth.rate(MPMsensitiviessite5)

site15.mpm <- list(MPM.plots$"8", MPM.plots$"578", MPM.plots$"581", MPM.plots$"799")  
MPMsensitiviessite15 <- lapply(1996:2021, function(yr){
  mpmyr <- do.call(c,lapply(site15.mpm, function(m) {
    m[grepl(as.character(yr), names(m))]
    }) )

  ms1 <- mean(mpmyr)
})
stoch15 <- stoch.growth.rate(MPMsensitiviessite15)

site19.mpm <- list(MPM.plots$"512", MPM.plots$"513", MPM.plots$"514", MPM.plots$"515", MPM.plots$'300')  
MPMsensitiviessite19 <- lapply(1996:2021, function(yr){
  mpmyr <- do.call(c,lapply(site19.mpm, function(m) {
    m[grepl(as.character(yr), names(m))]
    }) )

  ms1 <- mean(mpmyr)
})
stoch19 <- stoch.growth.rate(MPMsensitiviessite19)

site26.mpm <- list(MPM.plots$"480", MPM.plots$"611", MPM.plots$"614", MPM.plots$"238", MPM.plots$"598")  
MPMsensitiviessite26 <- lapply(1996:2021, function(yr){
  mpmyr <- do.call(c,lapply(site26.mpm, function(m) {
    m[grepl(as.character(yr), names(m))]
    }) )

  ms1 <- mean(mpmyr)
})
stoch26 <- stoch.growth.rate(MPMsensitiviessite26)

stoch.log.lam <- do.call(rbind, lapply(list(# stoch1,stoch2,
  stoch5,stoch15,stoch19,stoch26),
                                       function(x){
                                         data.frame(Approx = x$approx, Sim = x$sim,
                                                    lcl = x$sim.CI[1], ucl = x$sim.CI[2])
                                       }))

stoch.log.lam %>%
  mutate(Site = c(#"Site1","Site2",
    "Site5","Site15","Site19","Site26")) %>%
  # mutate(Sim = if_else(is.nan(Sim), Approx, Sim),
  #        lcl = if_else(is.nan(lcl), Approx, lcl),
  #        ucl = if_else(is.nan(ucl), Approx, ucl)) %>%
ggplot(  aes(Site, Sim, color = as.factor(Site)))+
  geom_point(position = position_dodge(width = 1))+
  geom_errorbar(aes(ymin = lcl, ymax = ucl),position = position_dodge(width = 1))+
  theme_bw()+
  ylab(expression(lambda))+
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey20")


```



Bootstrap lambda from Annual report tables and figures
```{r}
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/bootlambda.Rdata")
## From DormancyDetection_MPMMark-recpature.Rmd:  lambdaHDI
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel.Rdata")

paramsAll <- row.names(asmi.models$Phi.timePlot.p.dot.f.timePlot$design.matrix)
params <- paramsAll[!grepl("f ", paramsAll)]

# lmdas <- asmi.models$Phi.timePlot.p.dot.f.timePlot$results$derived$`Lambda Population Change` %>%
lmdas <- asmi.models$Phi.Plot.p.dot.f.Plot$results$derived$`Lambda Population Change` %>%
  bind_cols(Params = params[-length(params)]) %>%
  separate(Params, into = c("parameter","plot","age","time")) %>%
  mutate(time = as.character(time)) %>%
  mutate(time = gsub("t", "",time)) %>%
  mutate(time = as.numeric(time))

PlotSite <- asmi.all2 %>%
  distinct(site, plot)

lambdaHDI1 <- lambdaHDI %>%
  mutate(Plot = as.numeric(as.character(Plot))) %>%
  dplyr::rename(plot = Plot) %>%
  left_join(PlotSite) %>%
  dplyr::select(c(estimate:upper,plot,site)) %>%
  dplyr::rename(lcl = lower) %>%
  dplyr::rename(ucl = upper) %>%
  filter(site %in% c(5,15,19,26)) %>%
  mutate(Method = "MPM")
PradelLambda <- lmdas %>%
  mutate(plot = gsub("g","", plot)) %>%
  mutate(plot = as.numeric(plot)) %>%
  left_join(PlotSite) %>%
  filter(site %in% c(5,15,19,26)) %>%
  dplyr::select(c(estimate,lcl:ucl,plot,site)) %>%
  mutate(Method = "Pradel")

PradelLambda %>%
  bind_rows(lambdaHDI1) %>%
  ggplot( aes(as.factor(plot), estimate, color = Method))+
    geom_point(position = position_dodge(width = 1))+
    geom_errorbar(aes(ymin = lcl, ymax = ucl),position = position_dodge(width = 1))+
    theme_bw()+
    scale_color_manual(values = c("black","grey60")) +
  facet_wrap(~site, scales = "free_x")
  

```









```{r}
currentyr <- 2023

##To change if someone other than Michelle is running code
userpath <- "C:/Users/DePrengm/"

rawdatapath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/", currentyr, "_Astragalus-microcymbus_RawData.csv", collapse = '', sep = '')

asmi.raw <- read.csv(rawdatapath, na.strings = "na")

## plot 89 at site 1 was lost in 2020
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 89 & asmi.raw$year > 2019),]
asmi.raw <- asmi.raw[!is.na(asmi.raw$AsMi_site_id),]

asmiplots <- asmi.raw %>%
  distinct(AsMi_site_id, AsMi_plot_id) 
  # filter(complete.cases(.))
## Keeping old 2014 naming conventions
ASMIF4 <- 


```


Sensitivities, elasticities   
s1 probability a veg survives
s2 prob a rep survives

g1 prob veg survives and grows to Rep
g2 prob rep survives and grows

f1 prob veg survives, grows, produces fruit next year (transitions to reproductive)
f2 prob rep survives, maintains, produces fruits next year

r1 number fruit produced by veg (zero) and number that survive as veg next year
r2 number fruit produced by rep and number that survive as veg next year
```{r}
# Survival rates (S) 
asmiA <- 

vr <- c(prop.table(table()))
```


# Compare to recruitment and survival, lambda from Pradel model       
Pradel model for lambda (population growth rate) and recruitment (phi) and detection (p)    
```{r}
currentyr <- 2023
userpath <- "C:/Users/DePrengm/"
rawdatapath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/", currentyr, "_Astragalus-microcymbus_RawData.csv", collapse = '', sep = '')

asmi.raw <- read.csv(rawdatapath, na.strings = "na")

Pradel_asmi <- asmi.raw %>%
  filter(!(AsMi_tag_id %in% c(1177,1171,2746,3903))) %>%
  filter(!is.na(year)) %>%  
  filter(AsMi_site_id %in% c(5,15,19,26)) %>% ## exclude Cebolla Creek that was added 2014
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence)) %>%
  group_by(AsMi_tag_id) %>%
  filter(any(length > 0)) %>%
  ungroup() %>%
  mutate(Obs = case_when(length > 0 ~ 1, 
                         length == 0 ~ 0,
                         is.na(length) ~ 0)) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  tidyr::pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id)) %>%
  tidyr::unite(ch, Year1995:Year2022, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id)) %>%
  filter(grepl("1",ch))

dmPlot <- model.matrix(~ -1 + Plot, Pradel_asmi)
dimnames(dmPlot)[[2]][length(dimnames(dmPlot)[[2]])]
plotPradeldf <- Pradel_asmi %>%
  distinct(Plot)


Pradel_asmi %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlot) %>%
  mutate(Plot799 = paste(Plot799, ";", sep = "")) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel_asmi_mark.inp", sep = " ", 
              col.names = FALSE, row.names = FALSE)


asmiPradel.inp <- convert.inp("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel_asmi_mark.inp", 
                        group.df = plotPradeldf,
                        covariates = NULL,
                        use.comments = FALSE)

asmiproc <- process.data(asmiPradel.inp, model = "Pradrec", groups = "Plot", begin.time = 1995)
asmiddl <- make.design.data(asmiproc)

## Fix p to zero for plots not surveyed 
### filter(AsMi_plot_id != 598) %>% ## exclude the one plot added in 2004
### filter(!(AsMi_plot_id %in% c(300,238))) %>% ## Since two plots, 300, 238 were added in 1996
pindex <- asmiddl$p
p598indexApp3 <- as.numeric(row.names(asmiddl$p[(asmiddl$p$time %in% c("1995","1996","1997","1998","1999",
                                                                      "2000","2001","2002","2003") &
                                                   asmiddl$p$group == "598"),]))
p598idx <- pindex %>%
  mutate(time = as.numeric(as.character(time)))%>%
  filter(time < 2004 & group == "598") %>%
  mutate(idx = as.numeric(row.names(.)))
p300_238idx <- pindex %>%
  filter(time == 1995 & group %in% c("300","238")) %>%
  mutate(idx = as.numeric(row.names(.)))


run.asmi <- function(){
  Phi.dot = list(formula =  ~ 1)
  Phi.time = list(formula =  ~ time)
  Phi.Plot = list(formula =  ~ Plot)  ## so each plot is offset
  Phi.timePlot = list(formula =  ~ time + Plot)
  
  p.dot = list(formula =  ~ 1)
  # p.time = list(formula = ~ time)
  # p.Plot = list(formula = ~ -1 + Plot)
  # p.timePlot = list(formula = ~ time + Plot)
  
  f.dot = list(formula =  ~ 1)
  f.time = list(formula = ~ time)
  f.Plot = list(formula = ~ Plot)
  f.timePlot = list(formula = ~ time + Plot)
  
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiproc, ddl = asmiddl)
  
  return(asmi.results)
}

asmi.models <- run.asmi()
save(asmi.models, file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel.Rdata")
load( "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel.Rdata")

### LAMBDA
nrow(asmi.models$Phi.timePlot.p.dot.f.timePlot$results$derived$`Lambda Population Change`)
paramsAll <- row.names(asmi.models$Phi.timePlot.p.dot.f.timePlot$design.matrix)
params <- paramsAll[!grepl("f ", paramsAll)]
length(params[-length(params)])

lmdas <- asmi.models$Phi.timePlot.p.dot.f.timePlot$results$derived$`Lambda Population Change` %>%
  bind_cols(Params = params[-length(params)]) %>%
  separate(Params, into = c("parameter","plot","age","time")) %>%
  mutate(time = as.character(time)) %>%
  mutate(time = gsub("t", "",time)) %>%
  mutate(time = as.numeric(time))

allL <- allLambdas %>%
  mutate(lcl = Lambda, ucl = Lambda) %>%
  dplyr::rename(estimate = Lambda) %>%
  mutate(Method = "MPM") %>%
  relocate(lcl, .after = estimate) %>%
  relocate(ucl, .after = lcl) %>%
  mutate(Year = as.numeric(Year)) %>%
  filter(plot %in% c("606","607","608","609","610",
                     "8","578","581","799",
                     "512","513","514","515",
                     "480","611","614")) %>%
  mutate(plot = as.factor(plot))
```




