---
title: "Multistate CJS for annual analysis"
output: html_document
date: "2024-10-28"
---


```{r}
library(RMark)
library(dplyr)
library(tidyr)
library(patchwork)
library(ggplot2)
```


```{r}
currentyr <- as.numeric(format(as.Date(Sys.Date(),format="%Y-%m-%d"), "%Y"))

##To change if someone other than Michelle is running code
userpath <- "C:/Users/DePrengm/"

rawdatapath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/", currentyr, "_Astragalus-microcymbus_RawData.csv", collapse = '', sep = '')

asmi.raw <- read.csv(rawdatapath, na.strings = "na")

# need to remove all plot 89, site 1 in 2020 and beyond
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 89 & asmi.raw$year > 2019),]


# For the path and start of the name of each file
savepath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_AnnualReports/", currentyr, "_Astragalus-microcymbus_AnnualReport/", collapse = '', sep = '')

# Need to change all the seedling in 2014 for sites 1 and 2 to appropriate vegetative or reproductive 
asmi.raw$status[asmi.raw$AsMi_site_id<3 & asmi.raw$year == 2014 & asmi.raw$flower == 0] <- "vegetative"
asmi.raw$status[asmi.raw$AsMi_site_id<3 & asmi.raw$year == 2014 & asmi.raw$flower == 1] <- "reproductive"

# No longer adding climate data to the database
asmi.raw <- asmi.raw[,grep(paste(c("Temp","Rain","Snow","Aug.Jul"),collapse="|"), names(asmi.raw),
                           value = TRUE, invert = TRUE)]

asmi.raw$Browsing...Status[asmi.raw$Browsing...Status == "mammal" ] <- "Mammal"
asmi.raw$status <- factor(asmi.raw$status)
asmi.raw$length[asmi.raw$length == 921] <- 21

# If there are fruit, it flowered and needs to be reproductive, not vegetative
asmi.raw[asmi.raw$status=="vegetative" & asmi.raw$fruit>0,] # 5 times in 2020
wrongAsMidataid <- asmi.raw$AsMi_data_id[asmi.raw$status=="vegetative" & asmi.raw$fruit>0]
asmi.raw$flower[asmi.raw$AsMi_data_id %in% wrongAsMidataid] <- 1
asmi.raw$status[asmi.raw$AsMi_data_id %in% wrongAsMidataid] <- "reproductive"

asmi.raw <- asmi.raw %>%
  mutate(status = as.character(status)) %>%
  dplyr::mutate(status = case_when((AsMi_tag_id == 2551 & year == 2015) ~ "vegetative",
                            TRUE ~ status)) %>%
  mutate(status = as.factor(status))

## Add Creek

asmi.raw$Creek <- "SouthBeaver"
asmi.raw$Creek[asmi.raw$AsMi_site_id < 3] <- "Cebolla" 
asmi.raw$Creek <- as.factor(asmi.raw$Creek)

```



Multi-state model   
```{r}
load("Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/asmi.all2.Rdata") 


```



#process data for RMark
crdms.data<-process.data(crdms,model="CRDMS",time.interval=t.int,
strata.labels=c("1","U"))
#change Psi parameters that are obtained by subtraction
crdms.ddl<-make.design.data(crdms.data,
parameters=list(Psi=list(subtract.stratum=c("1","1"))))
levels are R then V because alphabetical

By default the remain in stratum are calculated by subtraction. (One minus sum of point estimates of other transitions)   
Computing the standard error and CI is harder. Instead, use TransitionMatrix for that. 



Add climate variables, time varying  
Add fencing as a factor



##### Should Cebolla and South Beaver be in separate models instead of fixing detection to zero for 1995:2013

##########################################################################################################################################
```{r}

load( paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.climate", currentyr,".Rdata", sep=""))
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.annual", currentyr,".Rdata", sep=""))
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.season", currentyr,".Rdata", sep=""))

```



```{r}



CJS_asmiCebolla <- asmi.raw %>%
  filter(!is.na(year)) %>%  
  filter(Creek == "Cebolla") %>%
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence,Creek)) %>%
  group_by(AsMi_tag_id) %>%
  filter(any(length > 0)) %>%
  mutate(first.cap.stage = status[which.min(year[length>0])]) %>%
  ungroup() %>%
  mutate(Obs = case_when(status %in% c("seedling","vegetative") ~ "V",
                         status == "reproductive" ~ "R",
                         status %in% c("dormant","dead") ~ "0",
                         TRUE ~ "0")) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence,first.cap.stage, Creek)) %>%
  ### UPDATE TO CURRENT YEAR ######
  unite(ch, Year2014:Year2024, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id),
         Creek = as.factor(Creek)) %>%
  relocate(Creek, .after = Plot) %>%
  filter((grepl("R",ch) | grepl("V",ch)))


# Different end times (lost plot 89 in 2019)
plotdfCebolla <- CJS_asmiCebolla %>%
  distinct(Plot)
dmPlotCebolla <- model.matrix(~ -1 + Plot, CJS_asmiCebolla)
dimnames(dmPlotCebolla)[[2]][length(dimnames(dmPlotCebolla)[[2]])]  # Plot98

CJS_asmiCebolla %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlotCebolla) %>%
  mutate(Plot98 = paste(Plot98, ";", sep = "")) %>%
  write.table(file = paste(savepath, "Multi_asmi_markCebolla.inp", sep = ""), sep = " ",
              col.names = FALSE, row.names = FALSE)

asmiMultiCebolla <- convert.inp(paste(savepath, "Multi_asmi_markCebolla.inp", sep = ""),
                        group.df = plotdfCebolla,
                        covariates = NULL,
                        use.comments = FALSE)



asmiMultiCebolla.proc <- process.data(asmiMultiCebolla, begin.time = 2014, model = "Multistrata", groups = "Plot",
                               strata.labels = c("R","V"))
## A vector of strata (one for each strata) that is the tostratum that is computed by subtraction
asmiMultiCebolla.ddl <- make.design.data(asmiMultiCebolla.proc)


WinterPPTCebolla <-  asmi.season %>%
  filter(Prev12 >= 2014) %>%
  filter(site %in% c("Cebolla Mid","Cebolla North")) %>%
  ungroup() %>%
  filter(Variable == "ppt" & season == "winter") %>%
  rename(WinterPPT = Value) %>%
  dplyr::select(Plot, Prev12, WinterPPT) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))


asmiMultiCebolla.ddl$S <- asmiMultiCebolla.ddl$S %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPTCebolla) 
  
asmiMultiCebolla.ddl$p <- asmiMultiCebolla.ddl$p %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPTCebolla) 
  
asmiMultiCebolla.ddl$Psi <- asmiMultiCebolla.ddl$Psi %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPTCebolla) 


SummerMaxCebolla <- asmi.season %>%
  filter(Prev12 >= 2014) %>%
  filter(site %in% c("Cebolla Mid","Cebolla North")) %>%
  ungroup() %>%
  filter(Variable == "tmax" & season == "springSummer") %>%
  rename(SummerMax = Value) %>%
  dplyr::select(Plot, Prev12, SummerMax) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))


asmiMultiCebolla.ddl$S <- asmiMultiCebolla.ddl$S %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(SummerMaxCebolla)

asmiMultiCebolla.ddl$p <- asmiMultiCebolla.ddl$p %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(SummerMaxCebolla)

asmiMultiCebolla.ddl$Psi <- asmiMultiCebolla.ddl$Psi %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(SummerMaxCebolla)


# By sites, have the plots the same over site 
PlotSiteCebolla <- CJS_asmiCebolla %>%
  distinct(Site, Plot)

asmiMultiCebolla.ddl$S <- asmiMultiCebolla.ddl$S %>%
  left_join(PlotSiteCebolla)
asmiMultiCebolla.ddl$p <- asmiMultiCebolla.ddl$p %>%
  left_join(PlotSiteCebolla)
asmiMultiCebolla.ddl$Psi <- asmiMultiCebolla.ddl$Psi %>%
  left_join(PlotSiteCebolla)


############### Cebolla
# ps.asmi <- ps.asmi[!(ps.asmi$Plot == 89 & ps.asmi$Year > 2019),]
# cohort is the year the plant was first seen
# time is the current observation (recapture)

pindexCebolla <- asmiMultiCebolla.ddl$p 
## Need to stop at '2018'to 2019 because no 2020
p.Cebolla89 <- pindexCebolla %>%
  mutate(time = as.numeric(as.character(time))) %>%
  mutate(group = as.numeric(as.character(group))) %>%
  filter((group %in% c(89) & 
            time >= 2019))

p.Cebolla.missing2024 <- pindexCebolla %>%
  filter((group %in% c("52") &
            time %in% c("2024")))

p.Cebollaindices <- unique(c(p.Cebolla89$par.index, 
                             p.Cebolla.missing2024$par.index)) 

p.Cebollavalues <- rep(0, length(p.Cebollaindices))

########################################################################################################
### Fix detection to 1 for the final survey
## Just for 2024
## All the 89s are already zeros for 2019 and later so make them 1 in 2018

# p.OneinfinalCebolla <- pindexCebolla %>%
#   filter(group == "89" & time == "2018" | 
#            time == "2024" & !( group %in% c("89","52") )| 
#            group == "52" & time == "2023" )
# 
# p1.Cebollaindices <- p.OneinfinalCebolla$par.index
# p1.Cebollavalues <- rep(1, length(p1.Cebollaindices))
# 
# ## put the zeros and ones together
# p.Cebollaindices <- c(p.Cebollaindices, p1.Cebollaindices)
# p.Cebollavalues <- c(p.Cebollavalues, p1.Cebollavalues)
```

# Cebolla Creek
```{r}


asmiCebolla <- function(){
  S.stratumSite = list(formula = ~ stratum:Site)
  S.stratumSiteWinter = list(formula = ~ stratum:WinterPPT:Site) 
  S.stratumSiteSummer = list(formula = ~ stratum:SummerMax:Site) 
  S.stratumSiteSummerWinter = list(formula = ~ stratum:WinterPPT:SummerMax:Site) 

  p.strata = list(formula = ~ stratum,
               fixed = list(index = p.Cebollaindices, value = p.Cebollavalues))
  Psi.stratumSite = list(formula = ~ -1 + stratum:tostratum:Site)
  Psi.stratumSitewinter = list(formula = ~ -1 + stratum:tostratum:WinterPPT:Site)
  Psi.stratumSitesummer = list(formula = ~ -1 + stratum:tostratum:SummerMax:Site)
  Psi.stratumSitewintersummer = list(formula = ~ -1 + stratum:tostratum:SummerMax:WinterPPT:Site)
  
  asmi.model.list <- create.model.list("Multistrata")
  asmi.results <- mark.wrapper(asmi.model.list, 
                               data = asmiMultiCebolla.proc, 
                               ddl = asmiMultiCebolla.ddl)
  
  return(asmi.results)
}

asmi.modelCebolla <- asmiCebolla()
save(asmi.modelCebolla, file = paste(savepath, "asmimodelCebolla.Rdata", sep = ""))
```

```{r}
load(paste(savepath, "asmimodelCebolla.Rdata", sep = ""))

asmi.modelCebolla

# Psi.tableCebolla <- get.real(asmi.modelCebolla$S.stratumSite.p.strata.Psi.stratumSite, "Psi", vcv = TRUE)
Psi.tableCebolla <- get.real(asmi.modelCebolla$S.stratumSiteWinter.p.strata.Psi.stratumSitewinter, "Psi", vcv = TRUE)

PsivaluesCebolla <- Psi.tableCebolla$estimates
PsivaluesCebolla$stratum <- factor(PsivaluesCebolla$stratum, levels = c("V","R"))
PsivaluesCebolla$tostratum <- factor(PsivaluesCebolla$tostratum, levels = c("V","R"))

## Don't differ by plot so just pick one
PsivaluesSite <- PsivaluesCebolla %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) 


### 
TM.CebollaMid <- TransitionMatrix(PsivaluesSite[PsivaluesSite$Plot == 1 & PsivaluesSite$time == 2014,], vcv.real = Psi.tableCebolla$vcv.real)
TM.CebollaNorth <- TransitionMatrix(PsivaluesSite[PsivaluesSite$Plot == 16,], vcv.real = Psi.tableCebolla$vcv.real)

## Detection
p.strataSite <- get.real(asmi.modelCebolla$S.stratumSiteWinter.p.strata.Psi.stratumSitewinter, "p", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  rename(p.estimates = estimate,
         p.lcl = lcl,
         p.ucl = ucl) %>%
  select(group, Plot, time, stratum,  p.estimates, p.lcl, p.ucl)


## Transitions
Psi.strataSiteWinter <- get.real(asmi.modelCebolla$S.stratumSiteWinter.p.strata.Psi.stratumSitewinter, "Psi", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  rename(Psi.estimates = estimate,
         Psi.lcl = lcl,
         Psi.ucl = ucl) %>%
  select(group, Plot, time, stratum, tostratum, Psi.estimates, Psi.lcl, Psi.ucl) %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPTCebolla) 

stratatostratalabels <- c("R" = "Vegetative to Reproductive", "V" = "Reproductive to Vegetative")

CC.Phi <- ggplot(Psi.strataSiteWinter, aes(WinterPPT, Psi.estimates, color = Plot)) +
  geom_point(position = position_dodge(width = 5)) +
  geom_errorbar(aes(ymin = Psi.lcl, ymax = Psi.ucl),position = position_dodge(width = 5))+
  facet_wrap(~tostratum, labeller = labeller(tostratum = stratatostratalabels))+
  theme_bw() +
  ylab("Transition rate") +
  scale_color_manual("Site", values = c("darkblue","skyblue"),
                     labels = c("1" = "Cebolla Mid",
                                "16" = "Cebolla North"))

S.strataSiteWinter <- get.real(asmi.modelCebolla$S.stratumSiteWinter.p.strata.Psi.stratumSitewinter, "S", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  rename(S.estimates = estimate,
         S.lcl = lcl,
         S.ucl = ucl) %>%
  select(group, Plot, time, stratum, S.estimates, S.lcl, S.ucl) %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPTCebolla) 

CC.S <- ggplot(S.strataSiteWinter, aes(WinterPPT, S.estimates, color = Plot)) +
  geom_point(position = position_dodge(width = 2)) +
  geom_errorbar(aes(ymin = S.lcl, ymax = S.ucl),position = position_dodge(width = 2))+
  facet_wrap(~stratum)+
  theme_bw() +
  ylab("Survival rate")+
  scale_color_manual("Site", values = c("darkblue","skyblue"),
                     labels = c("1" = "Cebolla Mid",
                                "16" = "Cebolla North"))

ggsave(paste(savepath, "CebollaCreek_survival.jpg", sep = ""),
       CC.S/CC.Phi,
width=210, height=200,units='mm', dpi=300)


################# By Site
## transitions
Psi.strataSite <- get.real(asmi.modelCebolla$S.stratumSite.p.strata.Psi.stratumSite, "Psi", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  rename(Psi.estimates = estimate,
         Psi.lcl = lcl,
         Psi.ucl = ucl) %>%
  select(group, Plot, stratum, tostratum, Psi.estimates, Psi.lcl, Psi.ucl)

## Survival
S.strataSite <- get.real(asmi.modelCebolla$S.stratumSite.p.strata.Psi.stratumSite, "S", se = TRUE) %>%  
  distinct(estimate, lcl, ucl, .keep_all = TRUE)  %>%
  rename(S.estimates = estimate,
         S.lcl = lcl,
         S.ucl = ucl) %>%
  select(group, Plot, stratum, S.estimates, S.lcl, S.ucl)


## Select a high winter ppt and a low winter ppt and compare
## Point estimates for MPM
MPMCebolla <- Psi.strataSite %>%
  left_join(S.strataSite) %>%
  mutate(MPMremain = (1-Psi.estimates) * S.estimates) %>%
  mutate(MPMtransition = Psi.estimates * S.estimates)

## Need to multiply recruitment for the upper right
MPMSiteMid <- MPMCebolla %>% filter(group == "1") 
CebollaMPMMid <- matrix(c(  MPMSiteMid$MPMremain[MPMSiteMid$stratum == "V"], MPMSiteMid$MPMtransition[MPMSiteMid$stratum == "R"],
                         MPMSiteMid$MPMtransition[MPMSiteMid$stratum == "V"], MPMSiteMid$MPMremain[MPMSiteMid$stratum == "R"]),
                     nrow = 2,
                     byrow = TRUE)

MPMSiteNorth<- MPMCebolla %>% filter(group == "16") 
CebollaMPMNorth <- matrix(c(  MPMSiteNorth$MPMremain[MPMSiteNorth$stratum == "V"], MPMSiteNorth$MPMtransition[MPMSiteNorth$stratum == "R"],
                         MPMSiteNorth$MPMtransition[MPMSiteNorth$stratum == "V"], MPMSiteNorth$MPMremain[MPMSiteNorth$stratum == "R"]),
                     nrow = 2,
                     byrow = TRUE)

CebollaMPMMid
CebollaMPMNorth

```












South Beaver Creek
```{r}
CJS_asmiSouthBeaver <- asmi.raw %>%
  filter(!is.na(year)) %>%  
  filter(Creek == "SouthBeaver") %>%
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence,Creek)) %>%
  group_by(AsMi_tag_id) %>%
  filter(any(length > 0)) %>%
  mutate(first.cap.stage = status[which.min(year[length>0])]) %>%
  ungroup() %>%
  mutate(Obs = case_when(status %in% c("seedling","vegetative") ~ "V",
                         status == "reproductive" ~ "R",
                         status %in% c("dormant","dead") ~ "0",
                         TRUE ~ "0")) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence,first.cap.stage, Creek)) %>%
  ### UPDATE TO CURRENT YEAR ######
  unite(ch, Year1995:Year2024, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id),
         Creek = as.factor(Creek)) %>%
  relocate(Creek, .after = Plot) %>%
  filter((grepl("R",ch) | grepl("V",ch)))


## Need to split by plot because different time frames and different times having fencing, different start times
plotdfSouthBeaver <- CJS_asmiSouthBeaver %>%
  distinct(Plot)
dmPlotSouthBeaver <- model.matrix(~ -1 + Plot, CJS_asmiSouthBeaver)
dimnames(dmPlotSouthBeaver)[[2]][length(dimnames(dmPlotSouthBeaver)[[2]])]  # Plot799



CJS_asmiSouthBeaver %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlotSouthBeaver) %>%
  mutate(Plot799 = paste(Plot799, ";", sep = "")) %>%
   ### FOR DBG0077
  # write.table(file = "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/Multi_asmi_mark.inp",sep = " ",
  write.table(file = paste(savepath, "Multi_asmi_markSouthBeaver.inp", sep = ""), sep = " ",
              col.names = FALSE, row.names = FALSE)

asmiMulti <- convert.inp(paste(savepath, "Multi_asmi_markSouthBeaver.inp", sep = ""),
                        group.df = plotdfSouthBeaver,
                        covariates = NULL,
                        use.comments = FALSE)


asmiMultiSouthBeaver.proc <- process.data(asmiMulti, begin.time = 1995, model = "Multistrata", groups = "Plot",
                               strata.labels = c("R","V"))
## A vector of strata (one for each strata) that is the tostratum that is computed by subtraction
asmiMultiSouthBeaver.ddl <- make.design.data(asmiMultiSouthBeaver.proc)


# By default the remain in stratum are calculated by subtraction. (One minus sum of point estimates of other transitions)   
# Computing the standard error and CI is harder. Instead, use TransitionMatrix for that. 
# Add climate variables, time varying  
# Add fencing as a factor


##Winter precipitation, summer maximum temp
WinterPPT <- asmi.season %>%
  filter(Prev12 >= 1995) %>%
  ungroup() %>%
  filter(Variable == "ppt" & season == "winter") %>%
  rename(WinterPPT = Value) %>%
  dplyr::select(Plot, Prev12, WinterPPT) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))

asmiMultiSouthBeaver.ddl$S <- asmiMultiSouthBeaver.ddl$S %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPT) 
  
asmiMultiSouthBeaver.ddl$p <- asmiMultiSouthBeaver.ddl$p %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPT) 
  
asmiMultiSouthBeaver.ddl$Psi <- asmiMultiSouthBeaver.ddl$Psi %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPT) 



  
SummerMax <- asmi.season %>%
  filter(Prev12 >= 1995) %>%
  ungroup() %>%
  filter(Variable == "tmax" & season == "springSummer") %>%
  rename(SummerMax = Value) %>%
  dplyr::select(Plot, Prev12, SummerMax) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))

asmiMultiSouthBeaver.ddl$S <- asmiMultiSouthBeaver.ddl$S %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(SummerMax)

asmiMultiSouthBeaver.ddl$p <- asmiMultiSouthBeaver.ddl$p %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(SummerMax)

asmiMultiSouthBeaver.ddl$Psi <- asmiMultiSouthBeaver.ddl$Psi %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(SummerMax)



# By sites, have the plots the same over site 
PlotSiteSBC <- CJS_asmiSouthBeaver %>%
  distinct(Site, Plot)

asmiMultiSouthBeaver.ddl$S <- asmiMultiSouthBeaver.ddl$S %>%
  left_join(PlotSiteSBC)
asmiMultiSouthBeaver.ddl$p <- asmiMultiSouthBeaver.ddl$p %>%
  left_join(PlotSiteSBC)
asmiMultiSouthBeaver.ddl$Psi <- asmiMultiSouthBeaver.ddl$Psi %>%
  left_join(PlotSiteSBC)

### South Beaver only
# Which sites/plots were fenced when; when Open == 1, when fenced == 0; switched back, that didn't fix it
PlotSiteYrFence <- asmi.all2 %>%
  filter(site > 3) %>%
  distinct(year,plot,site,fenced)
### Use PlotSiteYrFence to specify years when a plot was fenced
PlotSiteYrFence %>%
  filter(fenced == "y")
## Add  
plot2381995 <- PlotSiteYrFence %>%
  filter(plot %in% c(300,238) &
           year == 1996) %>%
  mutate(year = 1995)
plot606 <- PlotSiteYrFence %>%
  filter(plot %in% c(606) &
           year == 2021) 
plot606 <- plot606 %>%
  bind_rows(plot606) %>%
  mutate(year = c(2022, 2023))

add2024 <- PlotSiteYrFence %>%
  filter(year == 2023) %>%
  mutate(year = 2024)
PlotSiteYrFence <- PlotSiteYrFence %>%
  bind_rows(add2024) %>%
  bind_rows(plot2381995) %>%
  bind_rows(plot606)

asmiMultiSouthBeaver.ddl$S$fence <-  ifelse(asmiMultiSouthBeaver.ddl$S$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                               asmiMultiSouthBeaver.ddl$S$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiMultiSouthBeaver.ddl$S$fence <- as.factor(asmiMultiSouthBeaver.ddl$S$fence)

asmiMultiSouthBeaver.ddl$p$fence <-  ifelse(asmiMultiSouthBeaver.ddl$p$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                               asmiMultiSouthBeaver.ddl$p$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiMultiSouthBeaver.ddl$p$fence <- as.factor(asmiMultiSouthBeaver.ddl$p$fence)

  
asmiMultiSouthBeaver.ddl$Psi$fence <-  ifelse(asmiMultiSouthBeaver.ddl$Psi$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                               asmiMultiSouthBeaver.ddl$Psi$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiMultiSouthBeaver.ddl$Psi$fence <- as.factor(asmiMultiSouthBeaver.ddl$Psi$fence)

### Need to note, no surveys for some plots 1995, before 2004, before 2014, and after 2019
# i.e. fix indices as detection == 0 on occasions where there were no observations
############### South Beaver
# ps.asmi <- ps.asmi[!(ps.asmi$Plot == 598 & ps.asmi$Year < 2004),]
# ps.asmi <- ps.asmi[!(ps.asmi$Plot == 300 & ps.asmi$Year < 1996),]
# ps.asmi <- ps.asmi[!(ps.asmi$Plot == 238 & ps.asmi$Year < 1996),]

pindex <- asmiMultiSouthBeaver.ddl$p 
p.idAddedlater <- pindex %>%
  mutate(cohort = as.numeric(as.character(cohort))) %>%
  mutate(time = as.numeric(as.character(time))) %>%
  mutate(group = as.numeric(as.character(group))) %>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(598) & cohort < 2004 & time < 2005))
p.idAdded1996 <- pindex %>%
  mutate(cohort = as.numeric(as.character(cohort))) %>%
  mutate(time = as.numeric(as.character(time))) %>%
  mutate(group = as.numeric(as.character(group))) %>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(300,238) & cohort < 1996 & time < 1997))  

  
## Fixed indices
p.indices <- 
  unique(c(p.idAdded1996$par.index, 
           p.idAddedlater$par.index))
p.values <- rep(0, length(p.indices))


########################################################################################################
### Fix detection to 1 for the final survey
# p.OneinfinalSouthBeaver <- pindex %>%
#   filter(time == "2024")
# 
# p1.SouthBeaverindices <- p.OneinfinalSouthBeaver$par.index
# p1.SouthBeavervalues <- rep(1, length(p1.SouthBeaverindices))


#########################################################################################################


p.SouthBeaverindices <- c(p.indices) #, p1.SouthBeaverindices)
p.SouthBeavervalues <- c(p.values) #, p1.SouthBeavervalues)


```


South Beaver
```{r}


asmiSouthBeaver <- function(){
  S.stratumSiteFence = list(formula = ~ stratum:Site:fence)
  S.stratumSiteWinter = list(formula = ~ stratum:WinterPPT:Site:fence) 
  S.stratumSiteSummer = list(formula = ~ stratum:SummerMax:Site:fence) 
  S.stratumSiteSummerWinter = list(formula = ~ stratum:WinterPPT:SummerMax:Site:fence) 

  p.strata = list(formula = ~ stratum,
               fixed = list(index = p.SouthBeaverindices, value = p.SouthBeavervalues))
  
  Psi.stratumSiteFence = list(formula = ~ -1 + stratum:tostratum:Site:fence)
  Psi.stratumSitewinter = list(formula = ~ -1 + stratum:tostratum:WinterPPT:Site:fence)
  Psi.stratumSitesummer = list(formula = ~ -1 + stratum:tostratum:SummerMax:Site:fence)
  Psi.stratumSitewintersummer = list(formula = ~ -1 + stratum:tostratum:SummerMax:WinterPPT:Site:fence)
  
  asmi.model.list <- create.model.list("Multistrata")
  asmi.results <- mark.wrapper(asmi.model.list, 
                               data = asmiMultiSouthBeaver.proc, 
                               ddl = asmiMultiSouthBeaver.ddl)
  
  return(asmi.results)
}

asmi.modelSouthBeaver <- asmiSouthBeaver()
save(asmi.modelSouthBeaver, file = paste(savepath, "asmimodelSouthBeaver_v2.Rdata", sep = ""))

```


###########################################################################################################################
###########################################################################################################################
###########################################################################################################################


### Annual building of matrix models
These are factors that should influence survival and transitions, less so detection. Just have detection vary by strata, check factors for the rest, account for fencing in all
 Look at Fletcher c-hat in section 5.8 Chapter 5
 c-hat greater than 1 means overdispersion, variation from age or stage that isn't in the model


<http://www.phidot.org/forum/viewtopic.php?f=54&t=3410> 
"Note that I have this structured so that probability of remaining in the stratum is done by subtraction. That is the only way I can conceive of it working because you can't set the covariate value for the one computed by subtraction. One could also have ~stratum+distance such that probability of remaining is strata dependent but because Psi values sum to 1, this will happen to some degree without a stratum specific intercept. This is made clear by writing down formula for strata A and strata B using the values I assign in code below and beta is slope for distance and intercept is the Psi intercept value. 

Psi A to B = exp(intercept + 10*beta)/[1+exp(intercept + 10*beta)+exp(intercept + 20*beta)]
Psi A to C = exp(intercept + 20*beta)/[1+exp(intercept + 10*beta)+exp(intercept + 20*beta)]
Psi A to A = 1/[1+exp(intercept + 10*beta)+exp(intercept + 20*beta)]

Psi B to A = exp(intercept + 10*beta)/[1+exp(intercept + 10*beta)+exp(intercept + 15*beta)]
Psi B to C = exp(intercept + 15*beta)/[1+exp(intercept + 10*beta)+exp(intercept + 15*beta)]
Psi B to B = 1/[1+exp(intercept + 10*beta)+exp(intercept + 15*beta)]"    

these are linear models by distance. Without distance between sites
Psi A to B = exp(intercept/  (1+exp(intercept) / exp(intercept))  )   = exp(intercept)/1+exp(intercept)
Psi A to A = 1/exp(intercept)

Not in the model but after getting the estimate    

 Really, detection should depend on observers - but we have so many different people, didn't record well who did what, we semi-changed the strategy of digging around with metal detector or just digging in the dirt to find every tag. As plants died, tags might not mean there is a plant so we switched to looking for plants, then figuring out which tag they are associated with... So maybe time is a good proxy for 'trap' measures.   


