---
title: "Multistate CJS for annual analysis"
output: html_document
date: "2024-10-28"
---


```{r}
library(RMark)
library(dplyr)
library(tidyr)
library(patchwork)
library(ggplot2)
```


```{r}
currentyr <- as.numeric(format(as.Date(Sys.Date(),format="%Y-%m-%d"), "%Y"))

##To change if someone other than Michelle is running code
userpath <- "C:/Users/DePrengm/"

rawdatapath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/", currentyr, "_Astragalus-microcymbus_RawData.csv", collapse = '', sep = '')

asmi.raw <- read.csv(rawdatapath, na.strings = "na")

# need to remove all plot 89, site 1 in 2020 and beyond
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 89 & asmi.raw$year > 2019),]


# For the path and start of the name of each file
savepath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_AnnualReports/", currentyr, "_Astragalus-microcymbus_AnnualReport/", collapse = '', sep = '')

# Need to change all the seedling in 2014 for sites 1 and 2 to appropriate vegetative or reproductive 
asmi.raw$status[asmi.raw$AsMi_site_id<3 & asmi.raw$year == 2014 & asmi.raw$flower == 0] <- "vegetative"
asmi.raw$status[asmi.raw$AsMi_site_id<3 & asmi.raw$year == 2014 & asmi.raw$flower == 1] <- "reproductive"

# No longer adding climate data to the database
asmi.raw <- asmi.raw[,grep(paste(c("Temp","Rain","Snow","Aug.Jul"),collapse="|"), names(asmi.raw),
                           value = TRUE, invert = TRUE)]

asmi.raw$Browsing...Status[asmi.raw$Browsing...Status == "mammal" ] <- "Mammal"
asmi.raw$status <- factor(asmi.raw$status)
asmi.raw$length[asmi.raw$length == 921] <- 21

# If there are fruit, it flowered and needs to be reproductive, not vegetative
asmi.raw[asmi.raw$status=="vegetative" & asmi.raw$fruit>0,] # 5 times in 2020
wrongAsMidataid <- asmi.raw$AsMi_data_id[asmi.raw$status=="vegetative" & asmi.raw$fruit>0]
asmi.raw$flower[asmi.raw$AsMi_data_id %in% wrongAsMidataid] <- 1
asmi.raw$status[asmi.raw$AsMi_data_id %in% wrongAsMidataid] <- "reproductive"

asmi.raw <- asmi.raw %>%
  mutate(status = as.character(status)) %>%
  dplyr::mutate(status = case_when((AsMi_tag_id == 2551 & year == 2015) ~ "vegetative",
                            TRUE ~ status)) %>%
  mutate(status = as.factor(status))

## Add Creek

asmi.raw$Creek <- "SouthBeaver"
asmi.raw$Creek[asmi.raw$AsMi_site_id < 3] <- "Cebolla" 
asmi.raw$Creek <- as.factor(asmi.raw$Creek)

```



Multi-state model   
```{r}
load("Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/asmi.all2.Rdata") 



CJS_asmi <- asmi.raw %>%
  filter(!is.na(year)) %>%  
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence,Creek)) %>%
  group_by(AsMi_tag_id) %>%
  filter(any(length > 0)) %>%
  mutate(first.cap.stage = status[which.min(year[length>0])]) %>%
  ungroup() %>%
  mutate(Obs = case_when(status %in% c("seedling","vegetative") ~ "V",
                         status == "reproductive" ~ "R",
                         status %in% c("dormant","dead") ~ "0")) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence,first.cap.stage, Creek)) %>%
  ### UPDATE TO CURRENT YEAR ######
  unite(ch, Year1995:Year2024, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id),
         Creek = as.factor(Creek)) %>%
  relocate(Creek, .after = Plot) %>%
  filter((grepl("R",ch) | grepl("V",ch)))


# Any factor is a group variable, any numeric can be a covariate
plotSiteCreek <- CJS_asmi %>%
  distinct(Plot,Site,Creek)
dmPlot <- model.matrix(~ -1 + Plot, CJS_asmi)
dmSite <- model.matrix(~ -1 + Site, CJS_asmi)
dmCreek <- model.matrix(~ -1 + Creek, CJS_asmi)
dimnames(dmCreek)[[2]][length(dimnames(dmCreek)[[2]])] ## CreekSouthBeaver


## Need to split by plot because different time frames and different times having fencing
plotdf <- CJS_asmi %>%
  distinct(Plot)
dmPlot <- model.matrix(~ -1 + Plot, CJS_asmi)
dimnames(dmPlot)[[2]][length(dimnames(dmPlot)[[2]])]

CJS_asmi %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlot) %>%
  mutate(Plot799 = paste(Plot799, ";", sep = "")) %>%
   ### FOR DBG0077
  write.table(file = "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/Multi_asmi_mark.inp",sep = " ",
  # write.table(file = paste(savepath, "Multi_asmi_mark.inp", sep = ""), sep = " ",
              col.names = FALSE, row.names = FALSE)

asmiMulti <- convert.inp(
  "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/Multi_asmi_mark.inp",
  # paste(savepath, "Multi_asmi_mark.inp", sep = ""),
                        group.df = plotdf,
                        covariates = NULL,
                        use.comments = FALSE)

## By plot, needed since plots have different lengths 
asmiMulti.proc <- process.data(asmiMulti, begin.time = 1995, model = "Multistrata", groups = "Plot")
asmiMulti.ddl <- make.design.data(asmiMulti.proc)
```




Add climate variables, time varying  
Add fencing as a factor
```{r}

load( paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.climate", currentyr,".Rdata", sep=""))
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.annual", currentyr,".Rdata", sep=""))
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.season", currentyr,".Rdata", sep=""))


##Winter precipitation, summer maximum temp
WinterPPT <- asmi.season %>%
  ungroup() %>%
  filter(Variable == "ppt" & season == "winter") %>%
  rename(WinterPPT = Value) %>%
  dplyr::select(Plot, Prev12, WinterPPT) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))

asmiMulti.ddl$S <- asmiMulti.ddl$S %>%
  left_join(WinterPPT, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
asmiMulti.ddl$p <- asmiMulti.ddl$p %>%
  left_join(WinterPPT, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
asmiMulti.ddl$Psi <- asmiMulti.ddl$Psi %>%
  left_join(WinterPPT, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
SummerMax <- asmi.season %>%
  ungroup() %>%
  filter(Variable == "tmax" & season == "springSummer") %>%
  rename(SummerMax = Value) %>%
  dplyr::select(Plot, Prev12, SummerMax) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))

asmiMulti.ddl$S <- asmiMulti.ddl$S %>%
  left_join(SummerMax, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
asmiMulti.ddl$p <- asmiMulti.ddl$p %>%
  left_join(SummerMax, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
asmiMulti.ddl$Psi <- asmiMulti.ddl$Psi %>%
  left_join(SummerMax, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))

SummerPPT <- asmi.season %>%
  ungroup() %>%
  filter(Variable == "ppt" & season == "springSummer") %>%
  rename(SummerPPT = Value) %>%
  dplyr::select(Plot, Prev12, SummerPPT) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))

asmiMulti.ddl$S <- asmiMulti.ddl$S %>%
  left_join(SummerPPT, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
asmiMulti.ddl$p <- asmiMulti.ddl$p %>%
  left_join(SummerPPT, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
asmiMulti.ddl$Psi <- asmiMulti.ddl$Psi %>%
  left_join(SummerPPT, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))

# Which sites/plots were fenced when; when Open == 1, when fenced == 0; switched back, that didn't fix it
PlotSiteYrFence <- asmi.all2 %>%
  distinct(year,plot,site,fenced)
### Use PlotSiteYrFence to specify years when a plot was fenced
PlotSiteYrFence %>%
  filter(fenced == "y")
add2024 <- PlotSiteYrFence %>%
  filter(year == 2023) %>%
  mutate(year = 2024)
PlotSiteYrFence <- PlotSiteYrFence %>%
  bind_rows(add2024)

asmiMulti.ddl$S$Psience <-  ifelse(asmiMulti.ddl$S$time %in% PlotSiteYrFence$year[PlotSiteYrFence$Psienced == "y"] &
                               asmiMulti.ddl$S$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$Psienced == "y"],1,0)
asmiMulti.ddl$S$Psience <- as.factor(asmiMulti.ddl$S$Psience)

asmiMulti.ddl$p$Psience <-  ifelse(asmiMulti.ddl$p$time %in% PlotSiteYrFence$year[PlotSiteYrFence$Psienced == "y"] &
                               asmiMulti.ddl$p$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$Psienced == "y"],1,0)
asmiMulti.ddl$p$Psience <- as.factor(asmiMulti.ddl$p$Psience)

  
asmiMulti.ddl$Psi$Psience <-  ifelse(asmiMulti.ddl$Psi$time %in% PlotSiteYrFence$year[PlotSiteYrFence$Psienced == "y"] &
                               asmiMulti.ddl$Psi$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$Psienced == "y"],1,0)
asmiMulti.ddl$Psi$Psience <- as.factor(asmiMulti.ddl$Psi$Psience)
```


If want to get estimates by Site or Creek, need to add this
```{r}
# Add groups
asmiMulti.ddl$S <- asmiMulti.ddl$S %>%
  mutate(Site = case_when(Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 1] ~ 1,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 2] ~ 2,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 5] ~ 5,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 15] ~ 15,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 19] ~ 19,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 26] ~ 26)) %>%
  mutate(Site = as.factor(Site)) %>%
  mutate(Creek = case_when(Plot %in% plotSiteCreek$Plot[plotSiteCreek$Creek == "SouthBeaver"] ~ "SouthBeaver",
                           Plot %in% plotSiteCreek$Plot[plotSiteCreek$Creek == "Cebolla"] ~ "Cebolla")) %>%
  mutate(Creek = as.factor(Creek))

asmiMulti.ddl$p <- asmiMulti.ddl$p %>%
  mutate(Site = case_when(Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 1] ~ 1,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 2] ~ 2,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 5] ~ 5,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 15] ~ 15,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 19] ~ 19,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 26] ~ 26)) %>%
  mutate(Site = as.factor(Site)) %>%
  mutate(Creek = case_when(Plot %in% plotSiteCreek$Plot[plotSiteCreek$Creek == "SouthBeaver"] ~ "SouthBeaver",
                           Plot %in% plotSiteCreek$Plot[plotSiteCreek$Creek == "Cebolla"] ~ "Cebolla")) %>%
  mutate(Creek = as.factor(Creek))
asmiMulti.ddl$Psi <- asmiMulti.ddl$Psi %>%
  mutate(Site = case_when(Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 1] ~ 1,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 2] ~ 2,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 5] ~ 5,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 15] ~ 15,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 19] ~ 19,
                          Plot %in% plotSiteCreek$Plot[plotSiteCreek$Site == 26] ~ 26)) %>%
  mutate(Site = as.factor(Site)) %>%
  mutate(Creek = case_when(Plot %in% plotSiteCreek$Plot[plotSiteCreek$Creek == "SouthBeaver"] ~ "SouthBeaver",
                           Plot %in% plotSiteCreek$Plot[plotSiteCreek$Creek == "Cebolla"] ~ "Cebolla")) %>%
  mutate(Creek = as.factor(Creek))

```
    
```{r}
# Which sites/plots were fenced when; when Open == 1, when fenced == 0; switched back, that didn't fix it
PlotSiteYrFence <- asmi.all2 %>%
  distinct(year,plot,site,fenced)
## Add last year
addLast <- PlotSiteYrFence %>%
  filter(year == (currentyr-1))
PlotSiteYrFence <- PlotSiteYrFence %>%
  bind_rows(addLast)

## Time varying fencing; as a dummy variable, not a factor
asmiMulti.ddl$S$fence <- ifelse(asmiMulti.ddl$S$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                                  asmiMulti.ddl$S$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiMulti.ddl$S$fence <- as.factor(asmiMulti.ddl$S$fence) ## Should be numeric, and it is.

asmiMulti.ddl$Psi$fence <- ifelse(asmiMulti.ddl$Psi$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                                  asmiMulti.ddl$Psi$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiMulti.ddl$Psi$fence <- as.factor(asmiMulti.ddl$Psi$fence)

asmiMulti.ddl$p$fence <- ifelse(asmiMulti.ddl$p$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                                  asmiMulti.ddl$p$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiMulti.ddl$p$fence <- as.factor(asmiMulti.ddl$p$fence)


### Need to note, no surveys for some plots 1995, before 2004, before 2014, and after 2019
# i.e. fix indices as detection == 0 on occasions where there were no observations

# ps.asmi <- ps.asmi[!(ps.asmi$Plot == 89 & ps.asmi$Year > 2019),]
# ps.asmi <- ps.asmi[!(ps.asmi$Site < 3 & ps.asmi$Year < 2014),]
# ps.asmi <- ps.asmi[!(ps.asmi$Plot == 598 & ps.asmi$Year < 2004),]
# ps.asmi <- ps.asmi[!(ps.asmi$Plot == 300 & ps.asmi$Year < 1996),]
# ps.asmi <- ps.asmi[!(ps.asmi$Plot == 238 & ps.asmi$Year < 1996),]

## Should this have been time?! not cohort?? No, cohort and time 
plotsSite <- CJS_asmi %>%
  distinct(Plot,Site)
pindex <- asmiMulti.ddl$p 
p.idAddedlater <- pindex %>%
  mutate(cohort = as.numeric(as.character(cohort))) %>%
  mutate(group = as.numeric(as.character(group))) %>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(598) & cohort < 2004))
p.idAdded1996 <- pindex %>%
  mutate(cohort = as.numeric(as.character(cohort))) %>%
  mutate(group = as.numeric(as.character(group))) %>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(300,238) & cohort < 1996))  
p.Cebolla <- pindex %>%
  mutate(cohort = as.numeric(as.character(cohort))) %>%
  mutate(group = as.numeric(as.character(group))) %>%
  filter((group %in% c(1,22,32,42,52,16,58,89,90,96,98) & 
            cohort < 2013))
## Need to stop at '2018'to 2019 because no 2020
p.Cebolla89 <- pindex %>%
  mutate(cohort = as.numeric(as.character(cohort))) %>%
  mutate(group = as.numeric(as.character(group))) %>%
  filter((group %in% c(89) & 
            cohort >= 2019))
## None of them will be 2024 because it's going into 2024, need to be the 2023? And get rid of the '2024' to 2025
p.Cebolla.missing2024 <- pindex %>%
  mutate(cohort = as.numeric(as.character(cohort))) %>%
  mutate(group = as.numeric(as.character(group))) %>%
  filter((group %in% c("52") &
            cohort %in% c(2024)))
p.Cebolla.missing2024_time <- pindex %>%
  mutate(cohort = as.numeric(as.character(cohort))) %>%
  mutate(group = as.numeric(as.character(group))) %>%
  filter((group %in% c(52) & time %in% c("2024")))
  
## Fixed indices
p.indices <- 
  unique(c(p.idAdded1996$par.index, p.idAddedlater$par.index,
               p.Cebolla$par.index, p.Cebolla89$par.index,
               p.Cebolla.missing2024$par.index,p.Cebolla.missing2024_time$par.index))
p.values <- rep(0, length(p.indices))
## 10 repeats
length(p.indices) -
length(unique(p.indices))

## All the plot 52 in 2024 times! Oh, from the ones where the cohort was 1995:2012 (which is through year 2013)
## Added unique above to fix
pindex %>%
  filter(par.index %in% p.indices[which(duplicated(p.indices))])

```


Should detection be consistent across strata, fencing, Creeks? 
```{r}

# nest tostratum within stratum and fence within tostratum
## Make a function to compare multiple models

## First check best model for detection 
multi.asmi <- function(){
  S.dot = list(formula = ~ stratum)
  
  p.dot = list(formula = ~1,
               fixed = list(index = p.indices, value = p.values))
  p.fence = list(formula = ~ fence,
               fixed = list(index = p.indices, value = p.values))
  p.creek = list(formula = ~ Creek,
               fixed = list(index = p.indices, value = p.values))
  p.strata = list(formula = ~ stratum,
               fixed = list(index = p.indices, value = p.values))
  p.stratafence = list(formula = ~ stratum*fence,
               fixed = list(index = p.indices, value = p.values))
  p.stratafenceCreek = list(formula = ~ stratum*fence*Creek,
               fixed = list(index = p.indices, value = p.values))
  p.Time = list(formula = ~ Time,
               fixed = list(index = p.indices, value = p.values))
  p.Timestrata = list(formula = ~ stratum*Time,
               fixed = list(index = p.indices, value = p.values))
  p.TimeCreek = list(formula = ~ Time*Creek,
               fixed = list(index = p.indices, value = p.values))
  p.TimeStrataCreek = list(formula = ~ Time*stratum*Creek,
               fixed = list(index = p.indices, value = p.values))
  
  Psi.dot = list(formula = ~ -1 + stratum:tostratum)
  
  asmi.model.list <- create.model.list("Multistrata")
  asmi.results <- mark.wrapper(asmi.model.list, 
                               data = asmiMulti.proc, 
                               ddl = asmiMulti.ddl)
  
  return(asmi.results)
}

asmi.modelsMulti20241031 <- multi.asmi()

save(asmi.modelsMulti20241031, file = "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/multi_20241031.rdata")
```


```{r}

## Took 4 days to run
# load("Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/multi_20241031.rdata")
# save(asmi.modelsMulti20241031, file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/asmi_modelsMulti20241031.Rdata")
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/asmi_modelsMulti20241031.Rdata")

asmi.modelsMulti20241031$model.table

p.table <- get.real(asmi.modelsMulti20241031$S.dot.p.TimeStrataCreek.Psi.dot, "p", se = TRUE)
p.table %>%
  filter(fixed != "Fixed") %>%
  left_join(plotSiteCreek) %>%
  group_by(Time,Creek,stratum) %>%
  summarise(estimates = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(estimates != 0) %>%
ggplot(  aes(Time, estimates, color = stratum)) +
  geom_point(position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.9,position = position_dodge(width = 0.9)) +
  theme_bw()+
  facet_wrap(~Creek, scales = "free")


p.table <- get.real(asmi.modelsMulti20241031$S.dot.p.strata.Psi.dot, "p", se = TRUE)
p.table %>%
  filter(fixed != "Fixed") %>%
  left_join(plotSiteCreek) %>%
  group_by(Time,Creek,stratum) %>%
  summarise(estimates = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(estimates != 0) %>%
ggplot(  aes(Time, estimates, color = stratum)) +
  geom_point(position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.9,position = position_dodge(width = 0.9)) +
  theme_bw()+
  facet_wrap(~Creek, scales = "free")


S.table <- get.real(asmi.modelsMulti20241031$S.dot.p.TimeStrataCreek.Psi.dot, "S", se = TRUE)
S.table %>%
  filter(fixed != "Fixed") %>%
  left_join(plotSiteCreek) %>%
  group_by(Time,Creek,stratum) %>%
  summarise(estimates = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(estimates != 0) %>%
ggplot(  aes(Time, estimates, color = stratum)) +
  geom_point(position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.9,position = position_dodge(width = 0.9)) +
  theme_bw()+
  facet_wrap(~Creek, scales = "free")


## R to V, and V to R
Psi.table <- get.real(asmi.modelsMulti20241031$S.dot.p.TimeStrataCreek.Psi.dot, "Psi", se = TRUE)
Psi.table %>%
  filter(fixed != "Fixed") %>%
  left_join(plotSiteCreek) %>%
  group_by(Time,Creek,stratum) %>%
  summarise(estimates = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(estimates != 0) %>%
ggplot(  aes(Time, estimates, color = stratum)) +
  geom_point(position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.9,position = position_dodge(width = 0.9)) +
  theme_bw()+
  facet_wrap(~Creek, scales = "free")



```



### Annual building of matrix models
These are factors that should influence survival and transitions, less so detection. Just have detection vary by strata, check factors for the rest, account for fencing in all
```{r}

starttime <- Sys.Date()

run.Multi <- function(){
  p.strata = list(formula = ~ stratum,
                  fixed = list(index = p.indices, value = p.values))
  
  S.strata = list(formula = ~ stratum)
  S.strataCreek = list(formula = ~ -1 + stratum*Creek)
  S.strataCreekFence = list(formula = ~ -1 + stratum*Creek*fence)
  S.strataCreekWinterPPT = list(formula = ~ -1 + stratum*Creek*WinterPPT )
  S.strataCreekSummerPPT = list(formula = ~ -1 + stratum*Creek*SummerPPT )
  S.strataCreekSummerMax = list(formula = ~ -1 + stratum*Creek*SummerMax )
  S.strataCreekSummerMaxWinterPPT = list(formula = ~ -1 + stratum*Creek*SummerMax*WinterPPT )
  
  
  Psi.strata = list(formula = ~ -1 + stratum:tostratum)
  Psi.strataCreek = list(formula = ~ -1 + stratum:tostratum*Creek)
  Psi.strataCreekFence = list(formula = ~ -1 + stratum:tostratum*Creek*fence)
  Psi.strataCreekWinterPPT = list(formula = ~ -1 + stratum:tostratum*Creek*WinterPPT )
  Psi.strataCreekSummerPPT = list(formula = ~ -1 + stratum:tostratum*Creek*SummerPPT )
  Psi.strataCreekSummerMax = list(formula = ~ -1 + stratum:tostratum*Creek*SummerMax )
  Psi.strataCreekSummerMaxWinterPPT = list(formula = ~ -1 + stratum:tostratum*Creek*SummerMax*WinterPPT )
  

  asmi.model.list <- create.model.list("Multistrata")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiMulti.proc, ddl = asmiMulti.ddl)
  
  return(asmi.results)
}

asmiMultimodels <- run.Multi()
endtime <- Sys.Date()

save(asmiMultimodels, file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/",
                           starttime, "to", endtime,
                           "multistrata.Rdata",sep=""))

```




```{r}
###### Need fencing in all and stratum in all. Detection of reproductive will be different than vegetative
multi.mark2 <- mark(asmiMulti.proc, asmiMulti.ddl, begin.time = 1995, model = "Multistrata",
                   model.parameters = list(S = list(formula = ~ -1 + stratum:fence),
                                           p = list(formula = ~ -1 + stratum:fence,
                                                    fixed = list(index = p.indices, value = p.values)),
                                           Psi = list(formula = ~ -1 + stratum:tostratum:fence)))

save(multi.mark2, file = "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/multimarkStratumFence2024.rdata")

multi.mark.nofence <- mark(asmiMulti.proc, asmiMulti.ddl, begin.time = 1995, model = "Multistrata",
                   model.parameters = list(S = list(formula = ~ 1),
                                           p = list(formula = ~ 1,
                                                    fixed = list(index = p.indices, value = p.values)),
                                           Psi = list(formula = ~ -1 + stratum:tostratum)))

## To save from DBG0077
save(multi.mark.nofence, file = "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/multimarkNoFence2024.rdata")


## Look at Fletcher c-hat in section 5.8 Chapter 5
## c-hat greater than 1 means overdispersion, variation from age or stage that isn't in the model
```


x:y means that nests y within x, so for a given x, what is the y. So should probably have time:fence, not the fence first. and stratum:Creek:fence   
Not all plots (or Creek) or year, have fence, so that should be the last one.  
Let's see if it makes a difference   
```{r}
## First check best model for detection 
multi.asmi.v2 <- function(){
  S.dot = list(formula = ~1)
  
  p.dot = list(formula = ~1,
               fixed = list(index = p.indices, value = p.values))
  p.strata = list(formula = ~ -1 + stratum,
               fixed = list(index = p.indices, value = p.values))
  p.time = list(formula = ~ time,
               fixed = list(index = p.indices, value = p.values))
  p.fence = list(formula = ~ -1 + fence,
               fixed = list(index = p.indices, value = p.values))
  p.creek = list(formula = ~ -1 + Creek,
               fixed = list(index = p.indices, value = p.values))
  p.creektime = list(formula = ~ -1 + Creek + time,
               fixed = list(index = p.indices, value = p.values))
  ## This is the one that ended up as -1 + fence:time when written as: -1 + fence + time
  p.fencetime = list(formula = ~ -1 + time:fence,
               fixed = list(index = p.indices, value = p.values))
  p.stratacreek = list(formula = ~ -1 + strata:Creek,
               fixed = list(index = p.indices, value = p.values))
  p.stratatime = list(formula = ~ -1 + strata:time,
               fixed = list(index = p.indices, value = p.values))
  p.stratafencetime = list(formula = ~ -1 + strata:time:fence,
               fixed = list(index = p.indices, value = p.values))
  p.stratafence = list(formula = ~ -1 + stratum:fence,
               fixed = list(index = p.indices, value = p.values))
  p.stratafenceCreek = list(formula = ~ -1 + stratum:Creek:fence,
               fixed = list(index = p.indices, value = p.values))
  
  Psi.dot = list(formula = ~1)
  
  asmi.model.list <- create.model.list("Multistrata")
  asmi.results <- mark.wrapper(asmi.model.list, 
                               data = asmiMulti.proc, 
                               ddl = asmiMulti.ddl)
  
  return(asmi.results)
}

asmi.modelsMulti.v2 <- multi.asmi.v2()

save(asmi.modelsMulti.v2, file = "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/multimarkDetection2024.rdata")

```



```{r}

## Would need time for year when the climate was and can add climate variables
foo <- asmiMulti.ddl
foo$p <- merge_design.covariates(foo$p, plotSiteCreek)

load("Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/multimarkNoFence2024.rdata")
multi.mark.nofence$results$beta


## multi.mark
load("Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/multimarkFence2024.rdata")
multi.mark$results$real %>%
  distinct()

```



## If do site all together, need to split out into sites with similar length surveys
```{r}
sitedf <- CJS_asmi %>%
  distinct(Site)
dmSite <- model.matrix(~ -1 + Site, CJS_asmi)
dimnames(dmSite)[[2]][length(dimnames(dmSite)[[2]])]

CJS_asmi %>%
  dplyr::select(ch) %>%
  bind_cols(dmSite) %>%
  mutate(Site26 = paste(Site26, ";", sep = "")) %>%
  write.table(file = paste(savepath, "Multi_asmi_mark_site.inp", sep = ""), sep = " ", 
              col.names = FALSE, row.names = FALSE)

asmiMultisite <- convert.inp(paste(savepath, "Multi_asmi_mark_site.inp", sep = ""), 
                        group.df = sitedf,
                        covariates = NULL,
                        use.comments = FALSE)


asmiMultisite


```











```{r}
load("Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/multimarkDetection2024.rdata")

asmi.modelsMulti.v2$model.table
# Oy, detection changes with each year, that's not super useful.

```


## Really, detection should depend on observers - but we have so many different people, didn't record well who did what, we semi-changed the strategy of digging around with metal detector or just digging in the dirt to find every tag. As plants died, tags might not mean there is a plant so we switched to looking for plants, then figuring out which tag they are associated with... So maybe time is a good proxy for 'trap' measures.   


