---
title: "Pradel and CountPVA"
author: "Michelle DePrenger-Levin"
date: "2023-12-06"
output: html_document
---

Chapter 3: dormancy and detection in plants

# Population-level analyses
Demographic data for Astragalus microcymbus
RMark

```{r}
rm(list=ls())


library(dplyr)
library(tidyr)
library(RMark)
library(boot)
library(ggplot2)
library(lme4)

currentyr <- 2023
userpath <- "C:/Users/DePrengm/"
rawdatapath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/", currentyr, "_Astragalus-microcymbus_RawData.csv", collapse = '', sep = '')
asmi.raw <- read.csv(rawdatapath, na.strings = "na")
# need to remove all plot 89, site 1 in 2020 and beyond
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 89 & asmi.raw$year > 2019),]

# In a Stage-Fate format. stage is year t, fate is year t+1
load("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/stage-fate2023.Rdata")

asmi.raw %>%
  filter(is.na(year))

asmi.raw <- asmi.raw %>%
  filter(!is.na(year))

```



Check out count.DD.PVA from Bruce Kendall PVA package in R. Count-based Density-dependent PVA  https://rdrr.io/github/BruceKendall/PVA/man/count.DD.PVA.html     



counts and a count based PVA   
```{r}

### (1) ignoring dormant, only with above ground growth
count.yr <- asmi.raw %>%
  filter(length > 0) %>%
  dplyr::rename(plot = AsMi_plot_id) %>%
  dplyr::rename(site = AsMi_site_id) %>%
  group_by(site, plot, year) %>%
  dplyr::summarise(count = n())

plotsSite1 <- count.yr %>%
  distinct(site,plot)
  
## I think the lwr and upr are most appropriate 
loglams <- do.call(rbind, lapply(plotsSite1$plot, function(plot){
  y.count <- count.yr$count[count.yr$plot == plot]
  x.years <- count.yr$year[count.yr$plot == plot]
  ySpecies <- log(y.count[-1]/y.count[-length(y.count)])
  yr <- sqrt(x.years[-1]-x.years[-length(x.years)])

  
  PVA.lm <- lm(ySpecies ~ -1 + yr)
  mu_sp <- predict(PVA.lm, level= 0.95,
                   interval = "confidence",
                   se.fit = T)$fit[1,]
  
  
  rsq <- summary(PVA.lm)$adj.r.squared
  data.frame(plot = plot, mu = mu_sp[1], lwr = mu_sp[2], upr = mu_sp[3],
             geomean = mean(ySpecies), var.r = var(ySpecies), sd.r = sd(ySpecies), 
             lowerLambda = HDInterval::hdi(ySpecies)[1], upperLambda = HDInterval::hdi(ySpecies)[2])
  }))

## All together with plot per site as random effect
lmer.df <- do.call(rbind, lapply(plotsSite1$plot, function(plot){
  y.count <- count.yr$count[count.yr$plot == plot]
  x.years <- count.yr$year[count.yr$plot == plot]
  ySpecies <- log(y.count[-1]/y.count[-length(y.count)])
  yr <- sqrt(x.years[-1]-x.years[-length(x.years)])

  lmer.df <- data.frame(plot = plot, ySpecies, Year = x.years[-length(x.years)]) %>%
    left_join(plotsSite1)
  lmer.df
}))

lmer.df <- lmer.df %>%
  mutate(plot = as.factor(plot)) %>%
  filter(site %in% c(5,15,19,26)) %>%
  mutate(method = "Count without dormant") %>%
  mutate(site = as.factor(site))

## How many transitions? 
q <- length(diff(sort(unique(lmer.df$Year))))



# PVA.lm <- lmer(ySpecies ~ -1 + Year + (1|site), lmer.df)
# summary(PVA.lm)
PVA.lm <- lm(ySpecies ~ -1 + Year, lmer.df)
summary(PVA.lm)

## sum of squared errors 
sum(PVA.lm$residuals^2)
## mean squared error
mean(PVA.lm$residuals^2) ## this is the mean squared error
## Summary mean square
summary(PVA.lm)$sigma^2 ## sigma^2
sigma(PVA.lm)^2/PVA.lm$df.residual

### residual mean square == sum of squares of residuals divided by the degrees of freedom
mean(PVA.lm$residuals^2)/PVA.lm$df.residual

q <- PVA.lm$df.residual
## CI around mu
SE <- 2.394e-5
PVA.lm$coefficients-(qt(0.05, q)*SE)
PVA.lm$coefficients+(qt(0.05, q)*SE)

## CI around sigma^2
(q*summary(PVA.lm)$sigma^2 )/qchisq(c(0.025,0.975), df = PVA.lm$df.residual)


PVA.lm5 <- lm(ySpecies ~ -1 + Year, lmer.df[lmer.df$site == 5,])
summary(PVA.lm5)
PVA.lm15 <- lm(ySpecies ~ -1 + Year, lmer.df[lmer.df$site == 15,])
summary(PVA.lm15)
PVA.lm19 <- lm(ySpecies ~ -1 + Year, lmer.df[lmer.df$site == 19,])
summary(PVA.lm19)
PVA.lm26 <- lm(ySpecies ~ -1 + Year, lmer.df[lmer.df$site == 26,])
summary(PVA.lm26)
ggplot(lmer.df, aes(Year, ySpecies, color = plot))+
  geom_point()+
  stat_smooth(method = "lm", alpha = 0.3)+
  facet_wrap(~site, scales = "free_x")+
  theme_bw()+
  ylab(expression(log~lambda))+
  scale_color_grey(guide = 'none')+
  ggtitle("Count")

######################################################

loglams %>%
  left_join(plotsSite1) %>%
  mutate(site = as.factor(site),
         plot = as.factor(plot)) %>%
  # group_by(site) %>%
  # dplyr::summarise(sitegeomean = mean(geomean),
  #                  sitevar = var(var.r),
  #                  sitesd = sd(sd.r),
  #                  lowerLambda = HDInterval::hdi(geomean)[1],
  #                  upperLambda = HDInterval::hdi(geomean)[2]) %>% 
  filter(site %in% c(5,15,19,26))%>%
  ggplot( aes(plot, mu))+
    geom_point()+
    # geom_errorbar(aes(ymin = sitegeomean - sitesd, ymax = sitegeomean + sitesd), width = 0.2)+
    # geom_errorbar(aes(ymin = lowerLambda, ymax = upperLambda), width = 0.2)+
    geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2) +
    theme_bw() +
    geom_hline(yintercept = 0, linetype = "dotted") +
    ylab(expression(log~lambda))+
    xlab("Plot") +
  facet_wrap(~site, scales= "free_x")

#######################################################################################################
### (2) with dormants  
count.yr.wdorm <- asmi.raw %>%
  filter(status != "dead") %>%
  dplyr::rename(plot = AsMi_plot_id) %>%
  dplyr::rename(site = AsMi_site_id) %>%
  group_by(site, plot, year) %>%
  dplyr::summarise(count = n())

plotsSite1 <- count.yr.wdorm %>%
  distinct(site,plot)
  
## I think the lwr and upr are most appropriate 
loglamswdorm <- do.call(rbind, lapply(plotsSite1$plot, function(plot){
  y.count <- count.yr.wdorm$count[count.yr.wdorm$plot == plot]
  x.years <- count.yr.wdorm$year[count.yr.wdorm$plot == plot]
  ySpecies <- log(y.count[-1]/y.count[-length(y.count)])
  yr <- sqrt(x.years[-1]-x.years[-length(x.years)])

  
  PVA.lm <- lm(ySpecies ~ -1 + yr)
  mu_sp <- predict(PVA.lm, level= 0.95,
                   interval = "confidence",
                   se.fit = T)$fit[1,]
  
  
  rsq <- summary(PVA.lm)$adj.r.squared
  data.frame(plot = plot, mu = mu_sp[1], lwr = mu_sp[2], upr = mu_sp[3],
             geomean = mean(ySpecies), var.r = var(ySpecies), sd.r = sd(ySpecies), 
             lowerLambda = HDInterval::hdi(ySpecies)[1], upperLambda = HDInterval::hdi(ySpecies)[2])
  }))

## All together with plot per site as random effect
lmer.df.wdorm <- do.call(rbind, lapply(plotsSite1$plot, function(plot){
  y.count <- count.yr.wdorm$count[count.yr.wdorm$plot == plot]
  x.years <- count.yr.wdorm$year[count.yr.wdorm$plot == plot]
  ySpecies <- log(y.count[-1]/y.count[-length(y.count)])
  yr <- sqrt(x.years[-1]-x.years[-length(x.years)])

  lmer.df <- data.frame(plot = plot, ySpecies, Year = x.years[-length(x.years)]) %>%
    left_join(plotsSite1)
  lmer.df
}))


lmer.df.wdorm <- lmer.df.wdorm %>%
  mutate(plot = as.factor(plot)) %>%
  filter(site %in% c(5,15,19,26)) %>%
  mutate(method = "Count with dormant")

# PVA.lm.wdorm <- lmer(ySpecies ~ -1 + Year + (1|site), lmer.df.wdorm)
PVA.lm.wdorm <- lm(ySpecies ~ -1 + Year, lmer.df.wdorm)
summary(PVA.lm.wdorm)
ggplot(lmer.df.wdorm, aes(Year, ySpecies, color = plot))+
  geom_point()+
  stat_smooth(method = "lm", alpha = 0.3)+
  facet_wrap(~site, scales = "free_x")+
  theme_bw()+
  ylab(expression(log~lambda))+
  scale_color_grey(guide = 'none')+
  ggtitle("Count")


q.wd <- PVA.lm.wdorm$df.residual
## CI around mu
PVA.lm.wdorm$coefficients-(qt(0.05, q.wd)*2.394e-5)
PVA.lm.wdorm$coefficients+(qt(0.05, q.wd)*2.394e-5)

## CI around sigma^2
(q*summary(PVA.lm.wdorm)$sigma^2 )/qchisq(c(0.025,0.975), df = PVA.lm.wdorm$df.residual)



lmer.df %>%
  bind_rows(lmer.df.wdorm) %>%
  ggplot(  aes(Year, ySpecies, color = method))+
  geom_point()+
  stat_smooth(method = "lm", alpha = 0.3)+
  facet_wrap(~site, scales = "free_x")+
  theme_bw()+
  ylab(expression(log~lambda))+
  # scale_color_grey(guide = 'none')+
  ggtitle("Count")

######################################################

```



Format data of all years
```{r}
# asmi.raw <- asmi.raw3

## Just add "." for all the ones that we're measured
# assign group at first encounter, age 0 if 'seedling' or vegetative, older if reproductive. 
CJS_asmi <- asmi.raw %>%
  filter(!(AsMi_tag_id %in% c(1177,1171,2746,3903))) %>%
  filter(!is.na(year)) %>%  
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence)) %>%
  group_by(AsMi_tag_id) %>%
  filter(any(length > 0)) %>%
  mutate(first.cap.stage = status[which.min(year[length>0])]) %>%
  ungroup() %>%
  mutate(Obs = case_when(length > 0 ~ 1, 
                         length == 0 ~ 0,
                         is.na(length) ~ 0)) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence,first.cap.stage)) %>%
  unite(ch, Year1995:Year2022, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id)) %>%
  filter(grepl("1",ch))

dmPlot <- model.matrix(~ -1 + Plot, CJS_asmi)  
dmSite <- model.matrix(~ -1 + Site, CJS_asmi)
finalplot <- dimnames(dmPlot)[[2]][length(dimnames(dmPlot)[[2]])]

## By Plot
CJS_asmi %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlot) %>%
  mutate(Plot799 = paste(Plot799, ";", sep = "")) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_asmi_mark.inp", sep = " ", 
              col.names = FALSE, row.names = FALSE)

## By Site
CJS_asmi %>%
  dplyr::select(ch) %>%
  bind_cols(dmSite) %>%
  mutate(Site26 = paste(Site26, ";", sep = "")) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_asmi_mark_site.inp", sep = " ", 
              col.names = FALSE, row.names = FALSE)

fields <- CJS_asmi %>%
  dplyr::select(ch) %>%
  bind_cols(dmSite) %>%
  bind_cols(dmPlot) %>%
  colnames()

plotdf <- CJS_asmi %>%
  distinct(Site,Plot)

sitedf <- CJS_asmi %>%
  distinct(Site)

```



Mark-recapture by plot    
```{r}

## Convert for MARK  
asmi.inp <- convert.inp("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_asmi_mark.inp", 
                        group.df = plotdf,
                        covariates = NULL,
                        use.comments = FALSE)

asmiSite.inp <- convert.inp("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_asmi_mark_site.inp", 
                        group.df = sitedf,
                        covariates = NULL,
                        use.comments = FALSE)
```

CJS   
Took out "." notation, need to get index and fix to zero detection
```{r}
asmi.proc <- process.data(asmi.inp, model = "CJS", groups = "Plot")
asmi.ddl <- make.design.data(asmi.proc)

run.asmi <- function(){
  Phi.dot = list(formula =  ~ 1)
  Phi.time = list(formula =  ~ time)
  Phi.Plot = list(formula =  ~ Plot)  ## so each plot is offset
  Phi.timePlot = list(formula =  ~ time + Plot)
  
  p.dot = list(formula =  ~ 1)
  p.time = list(formula = ~ time)
  p.Plot = list(formula = ~ Plot)
  p.timePlot = list(formula = ~ time + Plot)

  asmi.model.list <- create.model.list("CJS")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmi.proc, ddl = asmi.ddl)
  
  return(asmi.results)
}

asmi.models <- run.asmi()
asmi.models$Phi.timePlot.p.timePlot$results$real

```

## Plot
Pradel model for lambda (population growth rate) and recruitment (phi) and detection (p)    
```{r}
Pradel_asmi <- asmi.raw %>%
  filter(!(AsMi_tag_id %in% c(1177,1171,2746,3903))) %>%
  filter(!is.na(year)) %>%  
  filter(AsMi_site_id %in% c(5,15,19,26)) %>%
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence)) %>%
  group_by(AsMi_tag_id) %>%
  filter(any(length > 0)) %>%
  # mutate(first.cap.stage = status[which.min(year[length>0])]) %>%
  ungroup() %>%
  mutate(Obs = case_when(length > 0 ~ 1, 
                         length == 0 ~ 0,
                         is.na(length) ~ 0)) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              # id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence,first.cap.stage)) %>%
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence)) %>%
  unite(ch, Year1995:Year2022, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id)) %>%
  filter(grepl("1",ch))

table(Pradel_asmi$Site, Pradel_asmi$Plot)

plotsSite <- Pradel_asmi %>%
  distinct(Plot,Site)

dmPlot <- model.matrix(~ -1 + Plot, Pradel_asmi)
dimnames(dmPlot)[[2]][length(dimnames(dmPlot)[[2]])]
plotPradeldf <- Pradel_asmi %>%
  distinct(Plot)
```

Takes a long time
```{r}
## without Cebolla Creek
Pradel_asmi %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlot) %>%
  mutate(Plot799 = paste(Plot799, ";", sep = "")) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel_asmi_mark.inp", sep = " ", 
              col.names = FALSE, row.names = FALSE)


asmiPradel.inp <- convert.inp("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel_asmi_mark.inp", 
                        group.df = plotPradeldf,
                        covariates = NULL,
                        use.comments = FALSE)

asmiproc <- process.data(asmiPradel.inp, model = "Pradrec", groups = "Plot", begin.time = 1995)
asmiddl <- make.design.data(asmiproc)

asmiproc$group.covariates

## Fix p to zero when no survey
pindex <- asmiddl$p
# p.idSite1_2 <- pindex %>%
#   mutate(time = as.numeric(as.character(time)))%>%
#   ## Sites 1 and 2 were added in 2014, nothing can start in any of these years, time
#   filter((group %in% plotsSite$Plot[plotsSite$Site %in% c(1,2)] & time < 2014)) 
p.idAddedlater <- pindex %>%
  mutate(time = as.numeric(as.character(time)))%>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(598) & time < 2004))
p.idAdded1996 <- pindex %>%
  mutate(time = as.numeric(as.character(time)))%>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(300,238) & time < 1996))  

## Fixed indices
p.indices <- c(#p.idSite1_2$par.index, 
  p.idAdded1996$par.index, p.idAddedlater$par.index)
p.values <- rep(0, length(p.indices))


run.asmi <- function(){
  Phi.dot = list(formula =  ~ 1)
  Phi.time = list(formula =  ~ time)
  Phi.Plot = list(formula =  ~ Plot)  ## so each plot is offset
  Phi.timePlot = list(formula =  ~ time + Plot)
  
  p.dot = list(formula =  ~ 1, fixed = list(index = p.indices, value = p.values))

  f.dot = list(formula =  ~ 1)
  f.time = list(formula = ~ time)
  f.Plot = list(formula = ~ Plot)
  f.timePlot = list(formula = ~ time + Plot)
  
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiproc, ddl = asmiddl)
  
  return(asmi.results)
}

asmi.models <- run.asmi()
save(asmi.models, file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/PradelSBC.Rdata")
```


## Examine Pradel models
plot Pradel with Count PVA
```{r}
## Both CC and SBC
# load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel.Rdata")
## South Beaver Creek only
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/PradelSBC.Rdata")

asmi.models
asmi.models$model.table

asmi.models$Phi.dot.p.dot.f.time$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  mutate(time = as.numeric(gsub("t", "", time))) %>%
  filter(parameter == "f") %>%
  # distinct(estimate,se,lcl,ucl) %>%
  ggplot(  aes(time, estimate))+
  geom_point() +
  geom_errorbar(aes(ymin = lcl, ymax = ucl))

asmi.models$Phi.timePlot.p.dot.f.timePlot$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  mutate(time = as.numeric(gsub("t", "", time))) %>%
  filter(parameter == "f") %>%
  # distinct(estimate,se,lcl,ucl) %>%
  ggplot(  aes(time, estimate, color = group))+
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl),position = position_dodge(width = 0.5))+
  theme_bw()

asmi.models$Phi.timePlot.p.dot.f.timePlot$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  mutate(time = as.numeric(gsub("t", "", time))) %>%
  filter(parameter == "f") %>%
  group_by(time) %>%
  dplyr::summarise(meanmu = mean(estimate)) %>%
  ggplot( aes(time, meanmu))+
  geom_point()+
  stat_smooth(method = "lm")

## Average recruitment across plots and years
asmi.models$Phi.timePlot.p.dot.f.timePlot$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  mutate(time = as.numeric(gsub("t", "", time))) %>%
  filter(parameter == "f") %>%
  dplyr::summarise(meanmu = mean(estimate))

## Highest recruitment 
asmi.models$Phi.timePlot.p.dot.f.timePlot$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  mutate(time = as.numeric(gsub("t", "", time))) %>%
  filter(parameter == "f") %>%
  group_by(time) %>%
  dplyr::summarise(meanmu = mean(estimate)) %>%
  filter(meanmu > 1)


asmi.models$Phi.timePlot.p.dot.f.timePlot$results$real %>%
  filter(estimate > 1.5)

asmi.models$Phi.time.p.dot.f.time$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  mutate(time = as.numeric(gsub("t", "", time))) %>%
  filter(parameter == "f") %>%
  filter(estimate > 0.01) %>%
  dplyr::summarise(n = n())

asmi.models$Phi.time.p.dot.f.time$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  mutate(time = as.numeric(gsub("t", "", time))) %>%
  filter(parameter == "f") %>%
  filter(estimate < 0.01) %>%
  dplyr::summarise(n = n())



# To compare to MPM estimation of fecundity
asmi.models$Phi.dot.p.dot.f.dot$results$real

asmi.models$Phi.time.p.dot.f.time$results$derived$`log(Lambda) Population Change` %>%
  distinct(estimate, se, lcl, ucl)

asmi.models$Phi.timePlot.p.dot.f.timePlot$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  filter(parameter == "Phi") %>%
  distinct(estimate,se,lcl,ucl)   

asmi.models$Phi.Plot.p.dot.f.Plot$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  filter(parameter == "p")
  

asmi.models$Phi.dot.p.dot.f.dot$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  filter(parameter == "f")


asmi.models$Phi.dot.p.dot.f.dot$results$derived$`log(Lambda) Population Change` %>%
  distinct(estimate, se, lcl, ucl)

asmi.models$Phi.Plot.p.dot.f.Plot$results$derived$`log(Lambda) Population Change` %>%
  distinct(estimate, se, lcl, ucl)
### LAMBDA
nrow(asmi.models$Phi.timePlot.p.dot.f.timePlot$results$derived$`Lambda Population Change`)
paramsAll <- row.names(asmi.models$Phi.timePlot.p.dot.f.timePlot$design.matrix)
paramsAll[grepl("Phi", paramsAll)]
params <- paramsAll[grepl("Phi", paramsAll)]

lmdasPlot <- asmi.models$Phi.timePlot.p.dot.f.timePlot$results$derived$`log(Lambda) Population Change` %>%
  bind_cols(Params = params) %>%
  separate(Params, into = c("parameter","plot","age","time")) %>%
  mutate(time = as.character(time)) %>%
  mutate(time = gsub("t", "",time)) %>%
  mutate(time = as.numeric(time)) %>%
  mutate(plot = gsub("g","", plot)) %>%
  mutate(Plot = as.factor(plot)) %>%
  left_join(plotsSite) %>%
  mutate(site = as.factor(Site))

# Pradel.lm <- lmer(estimate ~ -1 + time + (1|plot), lmdasPlot) 
# summary(Pradel.lm)
Pradel.lm1 <- lm(estimate ~ -1 + time, lmdasPlot)
summary(Pradel.lm1)

summary(Pradel.lm1)$sigma^2  # mean squared error
q <- Pradel.lm1$df.residual
## CI around mu
SE <- 1.003e-5
Pradel.lm1$coefficients-(qt(0.05, q)*SE)
Pradel.lm1$coefficients+(qt(0.05, q)*SE)

## CI around sigma^2
(q*summary(Pradel.lm1)$sigma^2 )/qchisq(c(0.025,0.975), df = Pradel.lm1$df.residual)

# Pradel.lm2 <- lmer(estimate ~ -1 + time + (1|site), lmdasPlot) 
# summary(Pradel.lm2)
ggplot(lmdasPlot, aes(time, estimate, color = site))+
  geom_point()+
  stat_smooth(method = "lm", alpha = 0.3)+
  facet_wrap(~site)+
  theme_bw()+
  ylab(expression(log~lambda))+
  scale_color_grey(guide = 'none')+
  ggtitle("Pradel")

CountLam <- lmer.df %>%
  dplyr::rename(estimate = ySpecies) %>%
  dplyr::rename(time = Year) %>%
  mutate(Method = "Count without dormant") %>%
  mutate(site = as.factor(site)) %>%
  filter(time < 2022)

CountLamsDorm <- lmer.df.wdorm %>%
  dplyr::rename(estimate = ySpecies) %>%
  dplyr::rename(time = Year) %>%
  mutate(Method = "Count with dormant") %>%
  mutate(site = as.factor(site)) %>%
  filter(time < 2022)

PradelMPMlongplot <- 
  lmdasPlot %>%
  dplyr::select(c(plot, estimate, time, site)) %>%
  mutate(plot = as.factor(plot)) %>%
  mutate(Method = "Pradel") %>%
  bind_rows(CountLam) %>%
  bind_rows(CountLamsDorm) %>%
ggplot(  aes(time, estimate, color = site)) + #, , color = site
             # color = Method, linetype = Method, 
             # shape = Method, group = interaction(plot,Method))) +
  geom_point(color = "grey50", shape = 20, alpha = 0.5, size = 1)+
  stat_smooth(method = "lm", alpha = 0.3, linewidth = 0.25)+
  theme_bw()+
  ylab(expression(log~lambda))+
  xlab("Year")+
  geom_hline(yintercept = 0, linetype = "dotted", color = "black")+
  facet_grid(~Method, scales = "free_y")+
  # facet_grid(Method~site, scales = "free_y")+
  scale_color_manual(guide = "none",values = c("red","blue","darkgreen","orange")) +
  scale_x_continuous(breaks = c( 2000, 2010,  2020))
  
  
ggsave(PradelMPMlongplot, filename = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/Chapter AsMi/Images/lambdasOverTimePradelCountPVA_v2.jpg",
         # width=125, height=120,units='mm', dpi=300 )
         width=150, height=70,units='mm', dpi=300 )


######################## dodge by plot? #########################
lmdasPlot %>%
  dplyr::select(c(plot, estimate, time, site)) %>%
  mutate(plot = as.factor(plot)) %>%
  mutate(Method = "Pradel") %>%
  bind_rows(CountLam) %>%
  bind_rows(CountLamsDorm) %>%
ggplot(  aes(time, estimate, group = site)) + #, , color = site
             # color = Method, linetype = Method, 
             # shape = Method, group = interaction(plot,Method))) +
  geom_point(color = "grey50", shape = 20, alpha = 0.2, size = 1, position = position_dodge(width = 0.01))+
  stat_smooth(method = "lm", alpha = 0.2)+
  theme_bw()+
  ylab(expression(log~lambda))+
  xlab("Year")+
  geom_hline(yintercept = 0, linetype = "dotted", color = "black")+
  facet_grid(~Method, scales = "free_y")+
  # facet_grid(Method~site, scales = "free_y")+
  scale_shape_manual(values = c(18,17,19)) +
  scale_x_continuous(breaks = c(1995, 2005, 2015))

get.real(asmi.models$Phi.timePlot.p.dot.f.timePlot, "p", vcv=TRUE)$estimates %>%
  distinct(estimate, se, lcl, ucl)
  
  
```


## Not this one
```{r}

## Pradel
lmdas <- asmi.models$Phi.timePlot.p.dot.f.timePlot$results$derived$`log(Lambda) Population Change` %>%
  bind_cols(Params = params) %>%
  separate(Params, into = c("parameter","plot","age","time")) %>%
  mutate(time = as.character(time)) %>%
  mutate(time = gsub("t", "",time)) %>%
  mutate(time = as.numeric(time)) %>%
  mutate(plot = gsub("g","", plot)) %>%
  mutate(plot = as.numeric(as.character(plot))) %>%
  group_by(plot) %>%
  dplyr::summarize(meanLambda = mean(estimate),
                   sdLambda = sd(estimate),
                   varLambda = var(estimate),
                   marginerror = qt(0.975, df = (n()-1)*sd(estimate)/sqrt(n())),
                   lcl = meanLambda - marginerror,
                   ucl = meanLambda + marginerror,
                   lowerHDI = HDInterval::hdi(estimate)[1],
                   upperHDI = HDInterval::hdi(estimate[2])) %>%
  mutate(plot = as.factor(plot)) %>%
  left_join(plotsSite)


#################### Lambdas from asmi.models$Phi.timePlot.p.dot.f.timePlot$results$derived$`log(Lambda) Population Change`
lmdas %>%
  mutate(site = as.factor(site)) %>%
  mutate(plot = as.factor(plot)) %>%
  # filter(Site %in% c(5,15,19,26)) %>%
  ggplot( aes(plot, meanLambda)) +
    geom_point()+
    # geom_errorbar(aes(ymin = meanLambda - sdLambda, ymax = meanLambda + sdLambda), width = 0.2)+
    geom_errorbar(aes(ymin = lowerHDI, ymax = upperHDI), width = 0.2) +
    geom_hline(yintercept = 0, linetype = "dotted")+
    theme_bw() +
  facet_wrap(~site, scales = "free_x")+
  ggtitle("Pradel")

# plotsSite <- plotsSite %>%
#   rename(plot = Plot) %>%
#   rename(site = Site)

CountLogLams <- loglams %>%
  mutate(plot = as.factor(plot)) %>%
  left_join(plotsSite) %>%
  filter(site %in% c(5,15,19,26))%>%
  dplyr::rename(estimate = mu) %>%
  # dplyr::rename(lcl = lwr) %>%
  # dplyr::rename(ucl = upr) %>%
  dplyr::rename(lcl = lowerLambda) %>%
  dplyr::rename(ucl = upperLambda) %>%
  mutate(Method = "Count") %>%
  dplyr::select(c(plot, estimate, lcl:ucl, site, Method))

PradelMPMplot <- lmdas %>%
  # mutate(lcl = meanLambda - sdLambda, ucl = meanLambda + sdLambda) %>%
  dplyr::select(plot, meanLambda, lowerHDI, upperHDI, site) %>%
  # dplyr::select(c(plot, estimate, lcl:ucl, site)) %>%
  # rename(estimate = meanLambda) %>%
  rename(estimate = meanLambda) %>%
  rename(lcl = lowerHDI) %>%
  rename(ucl = upperHDI) %>%
  filter(site %in% c(5,15,19,26))%>%
  mutate(plot = as.factor(plot)) %>%
  mutate(Method = "Pradel") %>%
  bind_rows(CountLogLams) %>%
ggplot(  aes(plot, estimate, color = Method)) +
  geom_point(position = position_dodge(width = 0.5))+
  geom_errorbar(aes(ymin = lcl, ymax = ucl),position = position_dodge(width = 0.5), width = 0.3)+
  theme_bw()+
  ylab(expression(log~lambda))+
  xlab("Plot")+
  geom_hline(yintercept = 0, linetype = "dotted")+
  scale_shape_manual(values = c(1,16))+
  facet_wrap(~site, scales = "free_x")+
  scale_color_manual(values = c("black","grey70"))
  
  
ggsave(PradelMPMplot, filename = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/Chapter AsMi/Images/lambdasPradelCountPVA.jpg",
         width=150, height=125,units='mm', dpi=300 )


#############################################################################################################
```


```{r}
## Geometric mean growth rate on real scale 
## Variance from random effects, variance components approach
lmdas %>%
  group_by(time) %>%
  summarise(lamhats = exp(mean(log(estimate)))) %>%
  print(n=30)

asmi.models$Phi.timePlot.p.dot.f.timePlot$results$derived$`log(Lambda) Population Change` %>%
  bind_cols(Params = params[-length(params)]) %>%
  separate(Params, into = c("parameter","plot","age","time")) %>%
  mutate(time = as.character(time)) %>%
  mutate(time = gsub("t", "",time)) %>%
  mutate(time = as.numeric(time)) %>%
  group_by(time) %>%
  summarise(lamhats = exp(mean(estimate)))

realasmi <- get.real(asmi.models$Phi.timePlot.p.dot.f.timePlot, "Phi", vcv=TRUE)
rasmi <- realasmi$estimates$estimate  ## One less than time length? or time length?
vcv <- realasmi$vcv.real
varc <- var.components(rasmi, design = matrix(rep(1,length(rasmi)), ncol=1), vcv)
df <- asmi.models$Phi.timePlot.p.dot.f.timePlot$design.data$Phi
shrinkest <- data.frame(time = 1:27, value = varc$betarand$estimate)
df <- merge(df, shrinkest, by="time")
md <- mark(asmiPradel.inp, model="Pradrec", 
           model.parameters = list(Phi = list(formula = ~ time + Plot,
                                              fixed=list(index=df$par.index,value=df$value)),
                                   f = list(formula = ~ time + Plot)),
           adjust=FALSE,delete=TRUE)

## Recruitment
recruit <- asmi.models$Phi.timePlot.p.dot.f.timePlot$results$real %>%
  bind_cols(Params = paramsAll)%>%
  separate(Params, into = c("parameter","plot","age","time")) %>%
  mutate(time = as.character(time)) %>%
  mutate(time = gsub("t", "",time)) %>%
  mutate(time = as.numeric(time))


## Recruitment is the number of added individuals C times the survival of adults is the change between time t+1 and t+2
recruit %>%
  filter(parameter == "f") %>%
  ggplot(    aes(time, estimate, color = plot)) +
  geom_point(position = position_dodge(width = 0.9))+
  geom_errorbar(aes(ymin = lcl, ymax = ucl),position = position_dodge(width = 0.9))+
  theme_bw()+
  ylab(expression(f[0]))+
  scale_color_discrete(guide = "none")

```

## Site
Pradel model for lambda (population growth rate) and recruitment (phi) and detection (p)    
```{r}
Pradel_asmi <- asmi.raw %>%
  filter(!(AsMi_tag_id %in% c(1177,1171,2746,3903))) %>%
  filter(!is.na(year)) %>%  
  filter(AsMi_site_id %in% c(5,15,19,26)) %>% ## exclude Cebolla Creek that was added 2014
  filter(AsMi_plot_id != 598) %>% ## exclude the one plot added in 2004
  filter(!(AsMi_plot_id %in% c(300,238))) %>% ## Since two plots, 300, 238 were added in 1996
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence)) %>%
  group_by(AsMi_tag_id) %>%
  filter(any(length > 0)) %>%
  mutate(first.cap.stage = status[which.min(year[length>0])]) %>%
  ungroup() %>%
  mutate(Obs = case_when(length > 0 ~ 1, 
                         length == 0 ~ 0,
                         is.na(length) ~ 0)) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence,first.cap.stage)) %>%
  unite(ch, Year1995:Year2022, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id))

dmSite <- model.matrix(~ -1 + Site, Pradel_asmi)
finalSite <- dimnames(dmSite)[[2]][length(dimnames(dmSite)[[2]])]
sitePradeldf <- Pradel_asmi %>%
  distinct(Site)


Pradel_asmi %>%
  dplyr::select(ch) %>%
  bind_cols(dmSite) %>%
  mutate(Site26 = paste(Site26, ";", sep = "")) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel_asmi_mark_site.inp", sep = " ", 
              col.names = FALSE, row.names = FALSE)


asmiPradel.inp <- convert.inp("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel_asmi_mark_site.inp", 
                        group.df = sitePradeldf,
                        covariates = NULL,
                        use.comments = FALSE)

asmiproc <- process.data(asmiPradel.inp, model = "Pradrec", groups = "Site", begin.time = 1995)
asmiddl <- make.design.data(asmiproc)

asmiproc$group.covariates

run.asmi <- function(){
  Phi.dot = list(formula =  ~ 1)
  Phi.time = list(formula =  ~ time)
  Phi.Plot = list(formula =  ~ Site)  ## so each plot is offset
  Phi.timePlot = list(formula =  ~ time + Site)
  
  p.dot = list(formula =  ~ 1)
  
  f.dot = list(formula =  ~ 1)
  f.time = list(formula = ~ time)
  f.Plot = list(formula = ~ Site)
  f.timePlot = list(formula = ~ time + Site)
  
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiproc, ddl = asmiddl)
  
  return(asmi.results)
}

# asmi.models.site <- run.asmi()
# save(asmi.models.site, file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/PradelSite.Rdata")
```



```{r}
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/PradelSite.Rdata")


### LAMBDA  ## Note that now Plot == Site
nrow(asmi.models.site$Phi.timePlot.p.dot.f.timePlot$results$derived$`Lambda Population Change`)
paramsAll <- row.names(asmi.models.site$Phi.timePlot.p.dot.f.timePlot$design.matrix)
params <- paramsAll[!grepl("f ", paramsAll)]
length(params[-length(params)])

# Pradel
lmdas <- asmi.models.site$Phi.timePlot.p.dot.f.timePlot$results$derived$`log(Lambda) Population Change` %>%
  bind_cols(Params = params[-length(params)]) %>%
  separate(Params, into = c("parameter","site","age","time")) %>%
  mutate(time = as.character(time)) %>%
  mutate(time = gsub("t", "",time)) %>%
  mutate(time = as.numeric(time))

# allL <- stoch.log.lam %>%
#   mutate(Sim = if_else(is.nan(Sim), Approx, Sim),
#          lcl = if_else(is.nan(lcl), Approx, lcl),
#          ucl = if_else(is.nan(ucl), Approx, ucl)) %>%
#   dplyr::rename(estimate = Sim) %>%
#   mutate(Method = "MPM") %>%
#   mutate(Year = as.numeric(Year)) %>%
#   dplyr::select(c(estimate:Method)) %>%
#   filter(Year < 2022)

PradelMPMsite <- lmdas %>%
  dplyr::select(c(estimate, lcl:ucl, site, time)) %>%
  dplyr::rename(Site = site) %>%
  mutate(Site = gsub("g","", Site)) %>%
  mutate(Site = as.numeric(Site)) %>%
  dplyr::rename(Year = time) %>%
  mutate(Method = "Pradel") %>%
  bind_rows(allL) %>%
ggplot(  aes(Year, estimate, color = as.factor(Site), shape = Method)) +
  geom_point(position = position_dodge(width = .8))+
  geom_errorbar(aes(ymin = lcl, ymax = ucl),position = position_dodge(width = .8), width = 1)+
  theme_bw()+
  ylab(expression("log"~lambda))+
  xlab("Year")+
  scale_color_manual("Site", values = wes_palette("Darjeeling1"))+
  # scale_color_manual("Method", values = c("black","grey50","grey30","grey80"))+
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey60")+
  scale_shape_manual(values = c(4,16))+
  # facet_wrap(~Site, nrow = 4)+
  theme(legend.position="bottom")
  
  
ggsave(PradelMPMsite, filename = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/Chapter AsMi/Images/lambdasPradelMPMSite.jpg",
         width=300, height=125,units='mm', dpi=300 )


#############################################################################################################


lambda.ftimePlot <- asmi.models.site$Phi.dot.p.dot.f.timePlot$results$derived$`log(Lambda) Population Change`%>%
  bind_cols(Params = params[-length(params)]) %>%
  separate(Params, into = c("parameter","site","age","time")) %>%
  mutate(time = as.character(time)) %>%
  mutate(time = gsub("t", "",time)) %>%
  mutate(time = as.numeric(time))

lambda.ftimePlot %>%
  dplyr::select(c(estimate, lcl:ucl, site, time)) %>%
  dplyr::rename(Site = site) %>%
  mutate(Site = gsub("g","", Site)) %>%
  mutate(Site = as.numeric(Site)) %>%
  dplyr::rename(Year = time) %>%
  mutate(Method = "Pradel") %>%
  bind_rows(allL) %>%
ggplot(  aes(Year, estimate, color = as.factor(Site), shape = Method)) +
  geom_point(position = position_dodge(width = .3))+
  geom_errorbar(aes(ymin = lcl, ymax = ucl),position = position_dodge(width = .3))+
  theme_bw()+
  ylab(expression("log"~lambda))+
  xlab("Year")+
  scale_color_manual("Site", values = wes_palette("Darjeeling1"))+
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey60")+
  scale_shape_manual(values = c(3,16))+
  # facet_wrap(~Site, nrow = 4)+
  theme(legend.position="bottom")

## Recruitment
recruit <- asmi.models.site$Phi.timePlot.p.dot.f.timePlot$results$real %>%
  bind_cols(Params = paramsAll)%>%
  separate(Params, into = c("parameter","site","age","time")) %>%
  mutate(time = as.character(time)) %>%
  mutate(time = gsub("t", "",time)) %>%
  mutate(time = as.numeric(time))


## Recruitment is the number of added individuals C times the survival of adults is the change between time t+1 and t+2
recruitplot <- recruit %>%
  filter(parameter == "f") %>%
  ggplot(    aes(time, estimate, color = site)) +
  geom_point(position = position_dodge(width = 0.9))+
  geom_errorbar(aes(ymin = lcl, ymax = ucl),position = position_dodge(width = 0.9))+
  theme_bw()+
  ylab(expression(f[0]))+
  xlab("Year")+
  scale_color_manual("Site", values = wes_palette("Darjeeling1"))+
  theme(legend.position="bottom")


ggsave(recruitplot, filename = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/Chapter AsMi/Images/recruitmentPradelMPMSite.jpg",
         width=300, height=125,units='mm', dpi=300 )

fandphi <- recruit %>%
  filter(parameter != "p") %>%
  ggplot(    aes(time, estimate, color = site, shape = parameter)) +
  geom_point(position = position_dodge(width = 0.9))+
  geom_errorbar(aes(ymin = lcl, ymax = ucl),position = position_dodge(width = 0.9))+
  theme_bw()+
  ylab("")+
  xlab("Year")+
  scale_color_manual("Site", values = wes_palette("Darjeeling1"))+
  scale_shape_manual("Parameter", labels = c(expression(f[0]), expression(Phi)), values = c(4,16))+
  theme(legend.position="bottom")+
  geom_hline(yintercept = recruit$estimate[recruit$parameter == "p"], color = "grey60", linetype = "dotted")+
  geom_hline(yintercept = recruit$lcl[recruit$parameter == "p"], color = "grey60", linetype = "longdash")+
  geom_hline(yintercept = recruit$ucl[recruit$parameter == "p"], color = "grey60", linetype = "longdash")

ggsave(fandphi, filename = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/Chapter AsMi/Images/recruitmentPhiPradelSite.jpg",
         width=300, height=125,units='mm', dpi=300 )
  
```




## Pradel only over the three years 2013-2015  
```{r}
rm(list=ls())


currentyr <- 2023
userpath <- "C:/Users/DePrengm/"
rawdatapath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/", currentyr, "_Astragalus-microcymbus_RawData.csv", collapse = '', sep = '')
asmi.raw <- read.csv(rawdatapath, na.strings = "na")
# need to remove all plot 89, site 1 in 2020 and beyond
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 89 & asmi.raw$year > 2019),]



Pradel_asmi_3 <- asmi.raw %>%
  filter(!(AsMi_tag_id %in% c(1177,1171,2746,3903))) %>%
  filter(!is.na(year)) %>%  
  filter(year >= 2013 & year <= 2015) %>%
  filter(AsMi_site_id %in% c(5,15,19,26)) %>%
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence)) %>%
  group_by(AsMi_tag_id) %>%
  filter(any(length > 0)) %>%
  # mutate(first.cap.stage = status[which.min(year[length>0])]) %>%
  ungroup() %>%
  mutate(Obs = case_when(length > 0 ~ 1, 
                         length == 0 ~ 0,
                         is.na(length) ~ 0)) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              # id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence,first.cap.stage)) %>%
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence)) %>%
  unite(ch, Year2013:Year2015, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id)) %>%
  filter(grepl("1",ch))

table(Pradel_asmi_3$Site, Pradel_asmi_3$Plot)

dmPlot <- model.matrix(~ -1 + Plot, Pradel_asmi_3)
dimnames(dmPlot)[[2]][length(dimnames(dmPlot)[[2]])]
plotPradeldf <- Pradel_asmi_3 %>%
  distinct(Plot) %>%
  arrange(Plot)

## without Cebolla Creek
Pradel_asmi_3 %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlot) %>%
  mutate(Plot799 = paste(Plot799, ";", sep = "")) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel_asmi_mark_3yrs.inp", sep = " ", 
              col.names = FALSE, row.names = FALSE)


asmiPradel.inp_3 <- convert.inp("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel_asmi_mark_3yrs.inp", 
                        group.df = plotPradeldf,
                        covariates = NULL,
                        use.comments = FALSE)

asmiproc <- process.data(asmiPradel.inp_3, model = "Pradrec", groups = "Plot", begin.time = 2013)
asmiddl <- make.design.data(asmiproc)

asmiproc$group.covariates
plotsSite <- Pradel_asmi_3 %>%
  distinct(Plot,Site)


run.asmi <- function(){
  Phi.dot = list(formula =  ~ 1)
  
  p.dot = list(formula =  ~ 1)

  f.dot = list(formula =  ~ 1)
  f.time = list(formula = ~ time)
  f.Plot = list(formula = ~ Plot)
  f.timePlot = list(formula = ~ time + Plot)
  
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiproc, ddl = asmiddl)
  
  return(asmi.results)
}

asmi.models.3yrs <- run.asmi()
save(asmi.models.3yrs, file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/PradelSBC_3yrs.Rdata")

asmi.models.3yrs$Phi.dot.p.dot.f.dot$results$real

```

```{r}
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/PradelSBC_3yrs.Rdata")
asmi.models.3yrs$Phi.dot.p.dot.f.dot$results$real

```

