---
title: "Pradel for annual report"
author: "Michelle DePrenger-Levin"
date: "2024-10-28"
output: html_document
---


```{r}

currentyr <- as.numeric(format(as.Date(Sys.Date(),format="%Y-%m-%d"), "%Y"))
userpath <- "C:/Users/DePrengm/"
savepath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_AnnualReports/", currentyr, "_Astragalus-microcymbus_AnnualReport/", collapse = '', sep = '')


library(dplyr)
library(tidyr)
library(RMark)
library(boot)
library(ggplot2)
library(lme4)

userpath <- "C:/Users/DePrengm/"
rawdatapath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/", currentyr, "_Astragalus-microcymbus_RawData.csv", collapse = '', sep = '')
asmi.raw <- read.csv(rawdatapath, na.strings = "na")
# need to remove all plot 89, site 1 in 2020 and beyond
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 89 & asmi.raw$year > 2019),]
# Need to remove all plot 52 in 2024, data from tablet was lost, didn't save 
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 52 & asmi.raw$year == 2024),]


# In a Stage-Fate format. stage is year t, fate is year t+1; asmi.all2
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/stage-fate",currentyr,".Rdata",
           sep=""))

asmi.raw %>%
  filter(is.na(year))

asmi.raw <- asmi.raw %>%
  filter(!is.na(year))

```


# Cebolla Creek
```{r}
CJS_asmi_Cebolla <- asmi.raw %>%
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence)) %>%
  # mutate(Creek = case_when(AsMi_site_id < 3 ~ "Cebolla",
  #                          AsMi_site_id > 4 ~ "SouthBeaver")) %>%
  filter(AsMi_site_id < 3) %>%
  ungroup() %>%
  mutate(Obs = case_when(length > 0 ~ 1, 
                         length == 0 ~ 0,
                         is.na(length) ~ 0)) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence)) %>%
  unite(ch, Year2014:Year2024, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id)) %>%
  filter(grepl("1",ch))

dmPlot_Cebolla <- model.matrix(~ -1 + Plot, CJS_asmi_Cebolla)  
dimnames(dmPlot_Cebolla)[[2]][length(dimnames(dmPlot_Cebolla)[[2]])]  # Plot98
```


Mark-recapture by plot    
```{r}

CJS_asmi_Cebolla %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlot_Cebolla) %>%
  mutate(Plot98 = paste(Plot98, ";", sep = "")) %>%
  write.table(file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/",
                           currentyr,
                           "CJS_asmi_markPradelCebolla.inp",sep=""),
              sep = " ", 
              col.names = FALSE, row.names = FALSE)

plotdfCebolla <- CJS_asmi_Cebolla %>%
  distinct(Site,Plot)

## Convert for MARK  
asmi.inpCebolla <- convert.inp(paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/",
                              currentyr,
                              "CJS_asmi_markPradelCebolla.inp",sep=""), 
                        group.df = plotdfCebolla,
                        covariates = NULL,
                        use.comments = FALSE)


asmiprocCebolla <- process.data(asmi.inpCebolla, model = "Pradrec",
                         groups = "Plot",
                         begin.time = 2014)

asmiddlCebolla <- make.design.data(asmiprocCebolla)

## Add site to ddl
asmiddlCebolla$Phi <- asmiddlCebolla$Phi %>%
  left_join(plotdfCebolla)
asmiddlCebolla$f <- asmiddlCebolla$f %>%
  left_join(plotdfCebolla)



## Fix detection, p, to zero when no survey
pindexCebolla <- asmiddlCebolla$p %>%
  left_join(plotdf)

p.Cebolla89 <- pindexCebolla %>%
  mutate(time = as.numeric(as.character(time))) %>%
  mutate(group = as.numeric(as.character(group))) %>%
  filter((group %in% c(89) & 
            time >= 2019))

p.Cebolla.missing2024 <- pindexCebolla %>%
  filter((group %in% c("52") &
            time %in% c("2024")))

p.Cebollaindices <- unique(c(p.Cebolla89$par.index, 
                             p.Cebolla.missing2024$par.index)) 

p.Cebollavalues <- rep(0, length(p.Cebollaindices))

########################################################################################################
### Fix detection to 1 for the final survey
## Just for 2024
## All the 89s are already zeros for 2019 and later so make them 1 in 2018

p.OneinfinalCebolla <- pindexCebolla %>%
  filter(group == "89" & time == "2018" | 
           time == "2024" & !( group %in% c("89","52") )| 
           group == "52" & time == "2023" )

p1.Cebollaindices <- p.OneinfinalCebolla$par.index
p1.Cebollavalues <- rep(1, length(p1.Cebollaindices))

## put the zeros and ones together
p.Cebollaindices <- c(p.Cebollaindices, p1.Cebollaindices)
p.Cebollavalues <- c(p.Cebollavalues, p1.Cebollavalues)



```


### HAVEN'T UPDATED YET
Add climate variables, time varying  
Add fencing as a factor
```{r}

load( paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.climate", currentyr,".Rdata", sep=""))
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.annual", currentyr,".Rdata", sep=""))
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.season", currentyr,".Rdata", sep=""))


##Winter precipitation, summer maximum temp
WinterPPT <- asmi.season %>%
  ungroup() %>%
  filter(Variable == "ppt" & season == "winter") %>%
  rename(WinterPPT = Value) %>%
  dplyr::select(Plot, Prev12, WinterPPT) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))

asmiddl$Phi <- asmiddl$Phi %>%
  left_join(WinterPPT, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
asmiddl$p <- asmiddl$p %>%
  left_join(WinterPPT, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
asmiddl$f <- asmiddl$f %>%
  left_join(WinterPPT, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
SummerMax <- asmi.season %>%
  ungroup() %>%
  filter(Variable == "tmax" & season == "springSummer") %>%
  rename(SummerMax = Value) %>%
  dplyr::select(Plot, Prev12, SummerMax) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))

asmiddl$Phi <- asmiddl$Phi %>%
  left_join(SummerMax, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
asmiddl$p <- asmiddl$p %>%
  left_join(SummerMax, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
asmiddl$f <- asmiddl$f %>%
  left_join(SummerMax, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))

SummerPPT <- asmi.season %>%
  ungroup() %>%
  filter(Variable == "ppt" & season == "springSummer") %>%
  rename(SummerPPT = Value) %>%
  dplyr::select(Plot, Prev12, SummerPPT) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))

asmiddl$Phi <- asmiddl$Phi %>%
  left_join(SummerPPT, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
asmiddl$p <- asmiddl$p %>%
  left_join(SummerPPT, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))
  
asmiddl$f <- asmiddl$f %>%
  left_join(SummerPPT, join_by("time" == "Prev12", 
                              "Plot" == "Plot"))

### By Site and Creek
asmiddl$Phi <- asmiddl$Phi %>%
  left_join(plotSiteCreek)

asmiddl$p <- asmiddl$p %>%
  left_join(plotSiteCreek)

asmiddl$f <- asmiddl$f %>%
  left_join(plotSiteCreek)

```


```{r}

run.Pradrec <- function(){
  Phi.dot = list(formula =  ~ Site)
  f.dot = list(formula =  ~ Site)

  p.dot = list(formula =  ~ 1,
               fixed = list(index = p.Cebollaindices, value = p.Cebollavalues))
  
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiprocCebolla, ddl = asmiddlCebolla)
  
  return(asmi.results)
}

asmiPradelmodelsCebolla <- run.Pradrec()
save(asmiPradelmodelsCebolla, file = paste(savepath, currentyr, "asmiPradelmodelsCebolla.rdata", sep=""))
```

```{r}
asmiPradelmodelsCebolla

f.table <- get.real(asmiPradelmodelsCebolla$Phi.dot.p.dot.f.dot, "f", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  rename(f.estimates = estimate,
         f.lcl = lcl,
         f.ucl = ucl) %>%
  select(group, Plot,f.estimates, f.lcl, f.ucl)

Phi.table <- get.real(asmiPradelmodelsCebolla$Phi.dot.p.dot.f.dot, "Phi", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  rename(Phi.estimates = estimate,
         Phi.lcl = lcl,
         Phi.ucl = ucl) %>%
  select(group, Plot,Phi.estimates, Phi.lcl, Phi.ucl)

f.table %>%
  left_join(Phi.table) %>%
  mutate(Lambda.estimate = f.estimates + Phi.estimates)
```

## Annual
```{r}

run.Pradrecann <- function(){
  Phi.dot = list(formula =  ~ time:Site)
  f.dot = list(formula =  ~ time:Site)

  p.dot = list(formula =  ~ 1,
               fixed = list(index = p.Cebollaindices, value = p.Cebollavalues))
  
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiprocCebolla, ddl = asmiddlCebolla)
  
  return(asmi.results)
}

asmiPradelmodelsCebollaann <- run.Pradrecann()
save(asmiPradelmodelsCebollaann, file = paste(savepath, currentyr, "asmiPradelmodelsCebollaann.rdata", sep=""))
```

```{r}
f.tableann <- get.real(asmiPradelmodelsCebollaann$Phi.dot.p.dot.f.dot, "f", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  rename(f.estimates = estimate,
         f.lcl = lcl,
         f.ucl = ucl) %>%
  select(group, time, Plot,f.estimates, f.lcl, f.ucl)

Phi.tableann <- get.real(asmiPradelmodelsCebollaann$Phi.dot.p.dot.f.dot, "Phi", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  rename(Phi.estimates = estimate,
         Phi.lcl = lcl,
         Phi.ucl = ucl) %>%
  select(group, time,  Plot,Phi.estimates, Phi.lcl, Phi.ucl)

f.tableann %>%
  left_join(Phi.tableann) %>%
  mutate(Lambda.estimate = f.estimates + Phi.estimates) %>%
  mutate(time = as.numeric(as.character(time))) %>%
ggplot(  aes(time, Lambda.estimate, color = group)) +
  geom_point() +
  stat_smooth(method = "lm")

```


First keep survival and recruitment constant and see how detection varies  
Good if we had variables where we'd expect detection to change, like observer or habitat. Could test plot
~ time as a factor, a separate parameter for each occasion   
~ Time as a trend, a continuous variable with a slope and intercept  (see Laake & Rexstad 2008)





### South Beaver Creek  

```{r}
p.idAddedlater <- pindex %>%
  mutate(time = as.numeric(as.character(time)))%>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(598) & time < 2004))
p.idAdded1996 <- pindex %>%
  mutate(time = as.numeric(as.character(time)))%>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(300,238) & time < 1996))  




# Which sites/plots were fenced when; when Open == 1, when fenced == 0; switched back, that didn't fix it
PlotSiteYrFence <- asmi.all2 %>%
  distinct(year,plot,site,fenced)
### Use PlotSiteYrFence to specify years when a plot was fenced
PlotSiteYrFence %>%
  filter(fenced == "y")
add2024 <- PlotSiteYrFence %>%
  filter(year == 2023) %>%
  mutate(year = 2024)
PlotSiteYrFence <- PlotSiteYrFence %>%
  bind_rows(add2024)

asmiddl$Phi$fence <-  ifelse(asmiddl$Phi$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                               asmiddl$Phi$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiddl$Phi$fence <- as.factor(asmiddl$Phi$fence)

asmiddl$p$fence <-  ifelse(asmiddl$p$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                               asmiddl$p$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiddl$p$fence <- as.factor(asmiddl$p$fence)

  
asmiddl$f$fence <-  ifelse(asmiddl$f$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                               asmiddl$f$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiddl$f$fence <- as.factor(asmiddl$f$fence)


```


#########################################################################################################################
#########################################################################################################################
#########################################################################################################################




### Let detection differ by site 
See if trend of survival or reproduction or if depends on winter ppt, summer tmax, or interaction of the two
```{r}
run.asmi.Phif <- function(){
  Phi.dot = list(formula =  ~ 1)
  Phi.age = list(formula = ~ Age)
  Phi.creek = list(formula = ~ Creek)
  Phi.winterppt = list(formula = ~ WinterPPT)
  Phi.summermax = list(formula = ~ SummerMax)
  Phi.winpptsummax = list(formula = ~ WinterPPT:SummerMax)
    
  f.dot = list(formula =  ~ 1)
  f.age = list(formula = ~ Age)
  f.creek = list(formula = ~ Creek)
  f.winterppt = list(formula = ~ WinterPPT)
  f.summermax = list(formula = ~ SummerMax)
  f.winpptsummax = list(formula = ~ WinterPPT:SummerMax)

  p.Plot = list(formula = ~ Plot,
               fixed = list(index = p.indices, value = p.values))
  
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiproc, ddl = asmiddl)
  
  return(asmi.results)
}

asmiSurvRep <- run.asmi.Phif()
save(asmiSurvRep, file = paste(savepath, currentyr, "asmiSurvRep.rdata", sep=""))
```

```{r}

load(paste(savepath, currentyr, "asmiSurvRep.rdata", sep=""))
asmiSurvRep

Phi.table <- get.real(asmiSurvRep$Phi.age.p.Plot.f.age, "Phi", se = TRUE)
Phi.table %>%
  filter(fixed != "Fixed") %>%
  group_by(Age) %>%
  summarise(estimates = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(estimates != 0) %>%
ggplot(  aes(Age, estimates)) +
  geom_point(position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.9,position = position_dodge(width = 0.9)) +
  theme_bw()+
  ylab("Survival")


f.table <- get.real(asmiSurvRep$Phi.age.p.Plot.f.age, "f", se = TRUE)
f.table %>%
  filter(fixed != "Fixed") %>%
  group_by(Age) %>%
  summarise(estimates = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(estimates != 0) %>%
ggplot(  aes(Age, estimates)) +
  geom_point(position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.9,position = position_dodge(width = 0.9)) +
  theme_bw()+
  ylab("Recruitment")


## Need ~ time
## Lambda 
Phi.table2 <- get.real(asmiSurvRep$Phi.winpptsummax.p.Plot.f.age, "Phi", se = TRUE)
f.table2 <- get.real(asmiSurvRep$Phi.winpptsummax.p.Plot.f.age, "f", se = TRUE)
Phi.table2 %>%
  select(estimate:Plot) %>%
  rename(Phi.estimate = estimate,
         Phi.se = se,
         Phi.lcl = lcl,
         Phi.ucl = ucl) %>%
  left_join(f.table2) %>%
  left_join(plotdf) %>%
  mutate(lambda = estimate + Phi.estimate) %>%
  mutate(time = as.numeric(as.character(time))) %>%
  mutate(Creek = case_when(Site %in% c(5,15,19,26) ~ "South Beaver",
                           TRUE ~ "Cebolla")) %>%
  group_by(Creek,time) %>%
  summarise(mean.lambda = mean(lambda)) %>%
  filter(Creek == "South Beaver" | (Creek == "Cebolla" & time > 2013)) %>%
ggplot(  aes(time, log(mean.lambda))) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw() +
  facet_wrap(~Creek, scales = "free")
  

### Is it that there are more individuals as precipitation increases so per capita recruitment goes down? 
f.table2 <- get.real(asmiSurvRep$Phi.age.p.Plot.f.winpptsummax, "f", se = TRUE)
f.table2 %>%
  filter(fixed != "Fixed") %>%
  left_join(WinterPPT, join_by("time" == "Prev12", "Plot" == "Plot")) %>%
  left_join(SummerMax, join_by("time" == "Prev12", "Plot" == "Plot")) %>%
  mutate(ppt_winter_bin = cut(WinterPPT, breaks = 3)) %>%
  group_by(Age,ppt_winter_bin, SummerMax) %>%
  summarise(estimates = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(estimates != 0) %>%
ggplot(  aes(SummerMax, estimates, color = Age)) +
  geom_point() +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.9) +
  theme_bw()+
  facet_wrap(~ppt_winter_bin) +
  ylab("Recruitment")

f.table2 %>%
  filter(fixed != "Fixed") %>%
  left_join(WinterPPT, join_by("time" == "Prev12", "Plot" == "Plot")) %>%
  left_join(SummerMax, join_by("time" == "Prev12", "Plot" == "Plot")) %>%
  mutate(max_summer_bin = cut(SummerMax, breaks = 3)) %>%
  group_by(max_summer_bin, WinterPPT) %>%
  summarise(estimates = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(estimates != 0) %>%
ggplot(  aes(WinterPPT, estimates)) +
  geom_point() +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.9) +
  theme_bw()+
  facet_wrap(~max_summer_bin) +
  ylab("Recruitment")

## No population size from this, need closed estimates to get population size
asmiSurvRep$Phi.age.p.Plot.f.age$results$derived$`log(Lambda) Population Change`
```




## Compare models  
```{r}


run.asmi.det2 <- function(){
  Phi.dot = list(formula =  ~ 1)
  f.dot = list(formula =  ~ 1)

  p.dot = list(formula =  ~ 1,
               fixed = list(index = p.indices, value = p.values))
  # Climate that might influence surrounding vegetation
  p.winterppt = list(formula = ~ WinterPPT,
               fixed = list(index = p.indices, value = p.values))
  p.summerMax = list(formula = ~ SummerMax,
               fixed = list(index = p.indices, value = p.values))
  p.summerppt = list(formula = ~ SummerPPT,
               fixed = list(index = p.indices, value = p.values))
  ## Did we get better over time? and as we amended our protocols - look for plants, not tags
  p.Time = list(formula = ~ Time,
               fixed = list(index = p.indices, value = p.values))
  p.Plot = list(formula = ~ -1 + Plot,
               fixed = list(index = p.indices, value = p.values))
  p.TimePlot = list(formula = ~ -1 + Time:Plot,
               fixed = list(index = p.indices, value = p.values))
  
  asmi.model.list <- create.model.list("Pradlambda")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiproc, ddl = asmiddl)
  
  return(asmi.results)
}

asmiPradeldetection <- run.asmi.det2()
save(asmiPradeldetection, file = paste(savepath, currentyr, "asmiPradeldetection.rdata", sep=""))


```

```{r}

load( paste(savepath, currentyr, "asmiPradeldetection.rdata", sep=""))

asmiPradeldetection$model.table

```


```{r}

## Need to merge climate
p.table <- get.real(asmiPradeldetection$Phi.dot.p.winterppt, "p", se = TRUE)
p.table %>%
  filter(fixed != "Fixed") %>%
  left_join(WinterPPT, join_by("time" == "Prev12", "Plot" == "Plot")) %>%
  group_by(WinterPPT) %>%
  summarise(estimates = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(estimates != 0) %>%
ggplot(  aes(WinterPPT, estimates)) +
  geom_point(position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.9,position = position_dodge(width = 0.9)) +
  theme_bw()


```


```{r}

asmiSurvRep$model.table

asmiSurvRep$Phi.Time.p.Time.f.Time$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  mutate(time = as.numeric(gsub("t", "", time))) %>%
  filter(parameter == "Phi") %>%
  filter(fixed != "Fixed") %>%
ggplot(   aes(time, estimate)) +
  geom_point()  +
  geom_errorbar(aes(ymin = lcl, ymax = ucl)) +
  xlab("Year")+
  ylab("Survival") +
  theme_bw()


asmiSurvRep$Phi.Time.p.Time.f.Time$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  mutate(time = as.numeric(gsub("t", "", time))) %>%
  filter(parameter == "f") %>%
  filter(fixed != "Fixed") %>%
ggplot(   aes(time, estimate)) +
  geom_point()  +
  geom_errorbar(aes(ymin = lcl, ymax = ucl)) +
  xlab("Year")+
  ylab("recruitment") +
  theme_bw()




```


### Let detection increase over time  !! No, don't
See if survival or reproduction change each year and if depends on winter ppt, summer tmax, or interaction of the two
```{r}
run.asmi.Phif2<- function(){
  Phi.dot = list(formula = 1)
  Phi.fence = list(formula = ~ -1 + fence)
  Phi.fencetime = list(formula = ~ -1 +time:fence)
  Phi.time = list(formula = ~ time)
  Phi.plot = list(formula = ~ Plot)
  Phi.plotfence = list(formula = ~ -1 + Plot:fence)
  Phi.winterppt = list(formula = ~ WinterPPT)
  Phi.summermax = list(formula = ~ SummerMax)
  Phi.winpptsummax = list(formula = ~ WinterPPT:SummerMax)
    
  f.dot = list(formula = 1)
  f.fence = list(formula = ~ -1 + fence)
  f.fencetime = list(formula = ~ -1 +time:fence)
  f.time = list(formula = ~ time)
  f.plot = list(formula = ~ Plot)
  f.plotfence = list(formula = ~ -1 + Plot:fence)
  f.winterppt = list(formula = ~ WinterPPT)
  f.summermax = list(formula = ~ SummerMax)
  f.winpptsummax = list(formula = ~ WinterPPT:SummerMax)

  ## Did we get better over time? and as we ammended our protocols - look for plants, not tags
  p.Time = list(formula = ~ Time,
               fixed = list(index = p.indices, value = p.values))
  
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiproc, ddl = asmiddl)
  
  return(asmi.results)
}

asmiSurvRep2 <- run.asmi.Phif2()
save(asmiSurvRep2, file = paste(savepath, currentyr, "asmiSurvRep2.rdata", sep=""))
```



```{r}

load(paste(savepath, currentyr, "asmiSurvRep2.rdata", sep=""))
asmiSurvRep2$model.table

asmiSurvRep2$Phi.fencetime.p.Time.f.time$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  mutate(time = as.numeric(gsub("t", "", time))) %>%
  filter(parameter == "f") %>%
  filter(fixed != "Fixed") %>%
ggplot(   aes(time, estimate)) +
  geom_point()  +
  geom_errorbar(aes(ymin = lcl, ymax = ucl)) +
  xlab("Year")+
  ylab("recruitment") +
  theme_bw()


asmiSurvRep2$Phi.fencetime.p.Time.f.time$results$real %>%
  mutate(prams = row.names(.)) %>%
  tidyr::separate(prams, into = c("parameter","group","age","time")) %>%
  mutate(time = as.numeric(gsub("t", "", time))) %>%
  filter(parameter == "Phi") %>%
  filter(fixed != "Fixed") %>%
  mutate(group = case_when(group == "g1" ~ "Open",
                           group == "g238" ~ "Fenced")) %>%
ggplot(   aes(time, estimate, color = group)) +
  geom_point(position = position_dodge(width = 0.9))  +
  geom_errorbar(aes(ymin = lcl, ymax = ucl),position = position_dodge(width = 0.9)) +
  xlab("Year")+
  ylab("Survival") +
  theme_bw()

## 870 rows   
## 38 distinct rows
asmiSurvRep2$Phi.fencetime.p.Time.f.time$results$derived$`log(Lambda) Population Change` %>%
  distinct()

### There are 10 years with fencing
require(plotrix)

p.table <- get.real(asmiSurvRep2$Phi.fencetime.p.Time.f.time, "p", se = TRUE)
## Missing 2024
PlotSiteYrFence <- PlotSiteYrFence %>%
  mutate(plot = as.factor(plot),
         year = as.factor(year),
         fenced = as.factor(fenced),
         site = as.factor(site))
p.table <- p.table %>%
  left_join(PlotSiteYrFence, join_by("time" == "year", "group" == "plot")) %>%
  mutate(fenced = case_when(fenced == "NA" ~ "n",
                            TRUE ~ fenced))
  # mutate(fenced = case_when(time == "2024" ~ "n",
  #                           TRUE ~ fenced))
estimates <- by(p.table$estimate, p.table$Time, mean)
lcl <- by(p.table$lcl, p.table$Time, mean)
ucl <- by(p.table$ucl, p.table$Time, mean)
plotCI(c(1:length(estimates)), estimates, ucl-estimates, estimates-lcl, 
         xlab = "Year",
         ylab = "Capture probability")

p.table %>%
  filter(fixed != "Fixed") %>%
  group_by(time,fenced) %>%
  summarise(estimates = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(estimates != 0) %>%
ggplot(  aes( time,estimates, color = fenced) ) +
  geom_point(position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.9,position = position_dodge(width = 0.9)) +
  theme_bw()

## Lambda = phi + f
phi.table <-  get.real(asmiSurvRep2$Phi.fencetime.p.Time.f.time, "Phi", se = TRUE)
phi.table <- phi.table %>%
  left_join(PlotSiteYrFence, join_by("time" == "year", "group" == "plot")) 
phi.table %>%
  filter(fixed != "Fixed") %>%
  filter(group != "NA") %>%
  filter(fenced != "NA") %>%
  group_by(time,fenced) %>%
  summarise(Survival = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(Survival != 0) %>%
  mutate(time = as.numeric(as.character(time))) %>%
ggplot(  aes( time,Survival, color = fenced) ) +
  geom_point(position = position_dodge(width = 0.7)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.7,position = position_dodge(width = 0.7)) +
  theme_bw()

f.table <-  get.real(asmiSurvRep2$f.fencetime.p.Time.f.time, "f", se = TRUE)
f.table <- f.table %>%
  left_join(PlotSiteYrFence, join_by("time" == "year", "group" == "plot")) 
f.table %>%
  filter(fixed != "Fixed") %>%
  filter(group != "NA") %>%
  filter(fenced != "NA") %>%
  group_by(time,fenced) %>%
  summarise(Recruit = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(Recruit != 0) %>%
  mutate(time = as.numeric(as.character(time))) %>%
ggplot(  aes( time,Recruit, color = fenced) ) +
  geom_point(position = position_dodge(width = 0.7)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.7,position = position_dodge(width = 0.7)) +
  theme_bw()


paramsAll <- row.names(asmiSurvRep2$Phi.fencetime.p.Time.f.time$model.parameters)
params <- paramsAll[grepl("Phi", paramsAll)]
length(paramsAll[grepl("g1", paramsAll)]) ## 1 row for fixed p at zero, 29 Phi rows, one for each year
length(paramsAll[grepl("g238", paramsAll)]) ## 10 Phi rows, one for each year of fencing


nrow(asmiSurvRep2$Phi.fencetime.p.Time.f.time$results$real)

lambdaPlot <- asmiSurvRep2$Phi.fencetime.p.Time.f.time$results$derived$`log(Lambda) Population Change` %>%
  bind_cols(Params = params) %>%
  separate(Params, into = c("parameter","plot","age","time")) %>%
  mutate(time = as.character(time)) %>%
  mutate(time = gsub("t", "",time)) %>%
  mutate(time = as.numeric(time)) %>%
  mutate(plot = gsub("g","", plot)) %>%
  mutate(Plot = as.factor(plot)) %>%
  left_join(plotsSite) %>%
  mutate(site = as.factor(Site))

```



Pradlambda
```{r}

asmiprocPradlambda <- process.data(asmi.inp, model = "Pradlambda",
                         groups = "Plot",
                         begin.time = 1995)
asmiddlPradlambda <- make.design.data(asmiprocPradlambda)


run.asmi.lambda <- function(){
  Phi.timePlot = list(formula =  ~ time:Plot)
  Lambda.timeplot = list(formula =  ~ time:Plot)

  p.plot = list(formula = ~ Plot,
               fixed = list(index = p.indices, value = p.values))
  
  asmi.model.list <- create.model.list("Pradlambda")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiprocPradlambda, ddl = asmiddlPradlambda)
  
  return(asmi.results)
}

asmi.lambdaresults <- run.asmi.lambda()
save(asmi.lambdaresults, file = "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmiLambda.Rdata")

```

```{r}

load("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmiLambda.Rdata")

lam.table <- get.real(asmi.lambdaresults$Phi.timePlot.p.plot.Lambda.timeplot, "Lambda", se = TRUE)
lam.table %>%
  left_join(plotdf) %>%
  group_by(Site, Plot) %>%
  summarise(lambda = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) 


lam.table %>%
  filter(fixed != "Fixed") %>%
  filter(fenced != "NA") %>%
  group_by(time,fenced) %>%
  summarise(f_recruit = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(f_recruit != 0) %>%
  mutate(time = as.numeric(as.character(time))) %>%
ggplot(  aes( time,f_recruit, color = fenced) ) +
  geom_point(position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.9,position = position_dodge(width = 0.9)) +
  theme_bw()





```



```{r}


run.asmi.lambda2 <- function(){
  Phi.dot = list(formula = 1)
  Phi.fence = list(formula = ~ -1 + fence)
  Phi.fencetime = list(formula = ~ -1 +time:fence)
  Phi.time = list(formula = ~ time)
  Phi.plot = list(formula = ~ Plot)
  Phi.plotfence = list(formula = ~ -1 + Plot:fence)
  Phi.winterppt = list(formula = ~ WinterPPT)
  Phi.summermax = list(formula = ~ SummerMax)
  Phi.winpptsummax = list(formula = ~ WinterPPT:SummerMax)
    
  f.dot = list(formula = 1)
  f.fence = list(formula = ~ -1 + fence)
  f.fencetime = list(formula = ~ -1 +time:fence)
  f.time = list(formula = ~ time)
  f.plot = list(formula = ~ Plot)
  f.plotfence = list(formula = ~ -1 + Plot:fence)
  f.winterppt = list(formula = ~ WinterPPT)
  f.summermax = list(formula = ~ SummerMax)
  f.winpptsummax = list(formula = ~ WinterPPT:SummerMax)

  ## Did we get better over time? and as we ammended our protocols - look for plants, not tags
  p.Time = list(formula = ~ Time,
               fixed = list(index = p.indices, value = p.values))
    
  asmi.model.list <- create.model.list("Pradlambda")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiproc, ddl = asmiddl)
  
  return(asmi.results)
}

asmi.lambdaresults2 <- run.asmi.lambda2()
save(asmi.lambdaresults2, file = "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/asmi_lambdaresults2.rdata")

```

```{r}

load( "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/asmi_lambdaresults2.rdata")
asmi.lambdaresults2$model.table


```


## Just change in lambda over years allowed to vary by year 'time', Plot, and fence while detection incraeses over time  
```{r}


run.detTimePlottimefence <- function(){
  
  Phi.plottimefence = list(formula = ~ -1 + time:Plot:fence)

  f.plottimefence = list(formula = ~ -1 + time:Plot:fence)

  ## Did we get better over time? and as we amended our protocols - look for plants, not tags
  p.Time = list(formula = ~ Time,
               fixed = list(index = p.indices, value = p.values))
    
  asmi.model.list <- create.model.list("Pradlambda")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiproc, ddl = asmiddl)
  
  return(asmi.results)
}

asmi.detTimePlottimefence2 <- run.detTimePlottimefence()
save(asmi.detTimePlottimefence2, file = "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/asmi_detTimePlottimefence2.rdata")


```



```{r}


run.Pradrec <- function(){
  
  Phi.plottimefence = list(formula = ~ -1 + time:Plot:fence)

  f.plottimefence = list(formula = ~ -1 + time:Plot:fence)

  ## Did we get better over time? and as we amended our protocols - look for plants, not tags
  p.Time = list(formula = ~ 1,
               fixed = list(index = p.indices, value = p.values))
    
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiproc, ddl = asmiddl)
  
  return(asmi.results)
}

asmi.Pradrec <- run.Pradrec()
save(asmi.Pradrec, file = "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/asmi_Pradrec.rdata")


```

```{r}
load("Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/asmi_Pradrec.rdata")

Phi.table <- get.real(asmi.Pradrec$Phi.plottimefence.p.Time.f.plottimefence, "Phi", se = TRUE)

PlotSiteYrFence <- PlotSiteYrFence %>%
  mutate(plot = as.factor(plot),
         year = as.factor(year),
         fenced = as.factor(fenced),
         site = as.factor(site))

Phi.table %>%
  left_join(PlotSiteYrFence, join_by("time" == "year", "group" == "plot")) %>%
  filter(fixed != "Fixed") %>%filter(group != "NA") %>%
  filter(fenced != "NA") %>%
  group_by(time,fenced) %>%
  summarise(Survival = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%
  filter(Survival != 0) %>%
  mutate(time = as.numeric(as.character(time))) %>%
ggplot(  aes( time,Survival, color = fenced) ) +
  geom_point(position = position_dodge(width = 0.7)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.7,position = position_dodge(width = 0.7)) +
  theme_bw()
  

### Recruitment, negative recruitment?!?! What now?!
f.table <- get.real(asmi.Pradrec$Phi.plottimefence.p.Time.f.plottimefence, "f", se = TRUE)

f.table %>%
  left_join(PlotSiteYrFence, join_by("time" == "year", "group" == "plot")) %>%
  filter(fixed != "Fixed") %>%filter(group != "NA") %>%
  filter(fenced != "NA") %>%
  group_by(time,fenced) %>%
  summarise(Recruitment = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl)) %>%Or 
  filter(Recruitment != 0) %>%
  mutate(time = as.numeric(as.character(time))) %>%
ggplot(  aes( time,log(Recruitment), color = fenced) ) +
  geom_point(position = position_dodge(width = 0.7)) +
  # geom_errorbar(aes(ymin = lcl, ymax = ucl), width = 0.7,position = position_dodge(width = 0.7)) +
  theme_bw()
  

```



## 
## Too meaningless to allow survival and reproduction to just vary year to year. Need to have it be a function of climate and fencing and Plot is ok
```{r}


run.Pradrec <- function(){
  
  Phi.winterppt = list(formula = ~ WinterPPT)
  Phi.summerppt = list(formula = ~ SummerPPT)
  Phi.summertmax = list(formula = ~ SummerMax)
  Phi.plotfence = list(formula = ~ -1 + Plot:fence)
  Phi.plotfenceWinterPPT = list(formula = ~ -1 + WinterPPT:Plot:fence)
  Phi.plotfenceSummerppt = list(formula = ~ -1 + SummerPPT:Plot:fence)
  Phi.plotfenceSummerMax = list(formula = ~ -1 + SummerMax:Plot:fence)


  f.winterppt = list(formula = ~ WinterPPT)
  f.summerppt = list(formula = ~ SummerPPT)
  f.summertmax = list(formula = ~ SummerMax)
  f.plotfence = list(formula = ~ -1 + Plot:fence)
  f.plotfenceWinterPPT = list(formula = ~ -1 + WinterPPT:Plot:fence)
  f.plotfenceSummerppt = list(formula = ~ -1 + SummerPPT:Plot:fence)
  f.plotfenceSummerMax = list(formula = ~ -1 + SummerMax:Plot:fence)

  ## Did we get better over time? and as we amended our protocols - look for plants, not tags
  p.Time = list(formula = ~ 1,
               fixed = list(index = p.indices, value = p.values))
    
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiproc, ddl = asmiddl)
  
  return(asmi.results)
}

asmi.PradrecClimate <- run.Pradrec()
save(asmi.PradrecClimate, file = "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/2024_forDBG0077/asmi_PradrecClimate.rdata")


```
