---
title: "Pradel for annual report"
author: "Michelle DePrenger-Levin"
date: "2024-10-28"
output: html_document
---


```{r}
rm(list=ls())
```

To combine recruitment with multistage transitions and survival for a MPM at high and low precipitation
```{r}


currentyr <- as.numeric(format(as.Date(Sys.Date(),format="%Y-%m-%d"), "%Y"))
userpath <- "C:/Users/DePrengm/"
savepath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_AnnualReports/", currentyr, "_Astragalus-microcymbus_AnnualReport/", collapse = '', sep = '')


library(dplyr)
library(tidyr)
library(RMark)
library(boot)
library(ggplot2)
library(lme4)

userpath <- "C:/Users/DePrengm/"
rawdatapath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/", currentyr, "_Astragalus-microcymbus_RawData.csv", collapse = '', sep = '')
asmi.raw <- read.csv(rawdatapath, na.strings = "na")
# need to remove all plot 89, site 1 in 2020 and beyond
## Don't remove, just make 0 in the detection probability, wasn't any chance of observing anything
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 89 & asmi.raw$year > 2019),]
# Need to remove all plot 52 in 2024, data from tablet was lost, didn't save 
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 52 & asmi.raw$year == 2024),]


# In a Stage-Fate format. stage is year t, fate is year t+1; asmi.all2
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/stage-fate",currentyr,".Rdata",
           sep=""))

asmi.raw %>%
  filter(is.na(year))

asmi.raw <- asmi.raw %>%
  filter(!is.na(year))

```

Climate
```{r, eval = FALSE}

load( paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.climate", currentyr,".Rdata", sep=""))
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.annual", currentyr,".Rdata", sep=""))
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.season", currentyr,".Rdata", sep=""))

```

```{r}

load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.season.wide2", currentyr,".Rdata", sep=""))

## From manual PRISM download
asmi.season.wide
```



# Cebolla Creek
```{r}
CJS_asmi_Cebolla <- asmi.raw %>%
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence)) %>%
  filter(AsMi_site_id < 3) %>%
  ungroup() %>%
  mutate(Obs = case_when(length > 0 ~ 1, 
                         length == 0 ~ 0,
                         is.na(length) ~ 0)) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id)) %>%
  ### UPDATE TO CURRENT YEAR ####
  unite(ch, Year2014:Year2025, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id)) %>%
  filter(grepl("1",ch))

## Why are there NAs for the missing 89 years?
CJS_asmi_Cebolla %>%
  mutate(n = nchar(ch)) %>%
  group_by(n) %>%
  dplyr::summarise(num = n())
length(2014:currentyr)  
## All 89 should be zeros in 2020 and after 
## Recorded 17 plants in 2020 
CJS_asmi_Cebolla %>%
  # Remove first 5
  mutate(lost_ch = substr(ch, 7, length(2014:currentyr) )) %>%
  filter(AsMi_plot_id == "89" & grepl("1", lost_ch)) %>%
  print(n = 68)


dmPlot_Cebolla <- model.matrix(~ -1 + Plot, CJS_asmi_Cebolla)  
dimnames(dmPlot_Cebolla)[[2]][length(dimnames(dmPlot_Cebolla)[[2]])]  # Plot98
```


Mark-recapture by plot    
```{r}

CJS_asmi_Cebolla %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlot_Cebolla) %>%
  mutate(Plot98 = paste(Plot98, ";", sep = "")) %>%
  write.table(file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/",
                           currentyr,
                           "CJS_asmi_markPradelCebolla.inp",sep=""),
              sep = " ", 
              col.names = FALSE, row.names = FALSE)

plotdfCebolla <- CJS_asmi_Cebolla %>%
  distinct(Site,Plot)

## Convert for MARK  
asmi.inpCebolla <- convert.inp(paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/",
                              currentyr,
                              "CJS_asmi_markPradelCebolla.inp",sep=""), 
                        group.df = plotdfCebolla,
                        covariates = NULL,
                        use.comments = FALSE)


asmiprocCebolla <- process.data(asmi.inpCebolla, model = "Pradrec",
                         groups = "Plot",
                         begin.time = 2014)

asmiddlCebolla <- make.design.data(asmiprocCebolla)

## Add site to ddl
asmiddlCebolla$Phi <- asmiddlCebolla$Phi %>%
  left_join(plotdfCebolla)
asmiddlCebolla$f <- asmiddlCebolla$f %>%
  left_join(plotdfCebolla)



## Fix detection, p, to zero when no survey
pindexCebolla <- asmiddlCebolla$p %>%
  left_join(plotdfCebolla)

## Data taken in 2019
p.Cebolla89 <- pindexCebolla %>%
  mutate(time = as.numeric(as.character(time))) %>%
  mutate(group = as.numeric(as.character(group))) %>%
  filter((group %in% c(89) & 
            time > 2019))

p.Cebolla.missing2024 <- pindexCebolla %>%
  filter((group %in% c("52") &
            time %in% c("2024")))

p.Cebollaindices <- unique(c(p.Cebolla89$par.index, 
                             p.Cebolla.missing2024$par.index)) 

p.Cebollavalues <- rep(0, length(p.Cebollaindices))

```



Add climate variables, time varying  
Add fencing as a factor to South Beaver Creek
```{r}

##Winter precipitation, summer maximum temp

# WinterPPTCebolla <-  asmi.season %>%
#   filter(Prev12 >= 2014) %>%
#   filter(site %in% c("Cebolla Mid","Cebolla North")) %>%
#   ungroup() %>%
#   filter(Variable == "ppt" & season == "winter") %>%
#   dplyr::rename(WinterPPT = Value) %>%
#   dplyr::mutate(WinterPPT_scaled = c(scale(WinterPPT))) %>%
#   dplyr::select(Plot, Prev12, WinterPPT, WinterPPT_scaled) %>%
#   # mutate(Prev12 = as.factor(Prev12)) %>%
#   mutate(Plot = as.factor(Plot))

asmi.season.wide <- asmi.season.wide %>%
  mutate(Plot = as.factor(Plot))

## Did it a year ahead because the year represents the probability of survival from time t to t+1 so the climate impact would be the climate previous to t+1
asmiddlCebolla$Phi <- asmiddlCebolla$Phi  %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 

asmiddlCebolla$p <- asmiddlCebolla$p  %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 

asmiddlCebolla$f <- asmiddlCebolla$f  %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 
  


# SummerMaxCebolla <- asmi.season %>%
#   filter(Prev12 >= 2014) %>%
#   filter(site %in% c("Cebolla Mid","Cebolla North")) %>%
#   ungroup() %>%
#   filter(Variable == "tmax" & season == "springSummer") %>%
#   dplyr::rename(SummerMax = Value) %>%
#   dplyr::mutate(SummerMax_scaled = c(scale(SummerMax))) %>%
#   dplyr::select(Plot, Prev12, SummerMax, SummerMax_scaled) %>%
#   mutate(Prev12 = as.factor(Prev12)) %>%
#   mutate(Plot = as.factor(Plot))
# 
# ## Only goes through 2023!
# asmi.season %>%
#   filter(Prev12 >= 2014) %>%
#   filter(site %in% c("Cebolla Mid","Cebolla North")) %>%
#   ungroup() %>%
#   filter(Variable == "tmax" & season == "springSummer") %>%
#   dplyr::rename(SummerMax = Value) %>%
#   arrange(desc(Prev12))
# 
# asmiddlCebolla$Phi <- asmiddlCebolla$Phi %>%
#   # mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
#   mutate(Prev12 = time) %>%
#   left_join(SummerMaxCebolla)
# 
# asmiddlCebolla$p <- asmiddlCebolla$p %>%
#   # mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
#   mutate(Prev12 = time) %>%
#   left_join(SummerMaxCebolla)
# 
# asmiddlCebolla$f <- asmiddlCebolla$f %>%
#   # mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
#   mutate(Prev12 = time) %>%
#   left_join(SummerMaxCebolla)

```


NAs need to get rid of them! What!?! WinterPPT and SummerMax are missing data in 2023 and 2024
#########################################  Cebolla ##########################################
```{r}

run.PradrecCebolla <- function(){
  Phi.site = list(formula =  ~ Site)
  Phi.siteWinter = list(formula =  ~ Site:ppt_winter) #Site:WinterPPT_scaled)
  Phi.siteWinterSummer = list(formula =  ~  Site:ppt_winter:tmax_springSummer)# Site:WinterPPT_scaled:SummerMax)
  Phi.siteSummer = list(formula =  ~ Site:tmax_springSummer) #SummerMax)
  
  f.site = list(formula =  ~ Site)
  f.siteWinter = list(formula =  ~ Site:ppt_winter) #WinterPPT_scaled)
  f.siteWinterSummer = list(formula =  ~ Site:ppt_winter:tmax_springSummer) # WinterPPT_scaled:SummerMax)
  f.siteSummer = list(formula =  ~ Site:tmax_springSummer) # SummerMax)

  p.dot = list(formula =  ~ 1,
               fixed = list(index = p.Cebollaindices, value = p.Cebollavalues))
  
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiprocCebolla, ddl = asmiddlCebolla)
  
  return(asmi.results)
}

asmiPradelmodelsCebolla <- run.PradrecCebolla()
save(asmiPradelmodelsCebolla, file = paste(savepath, currentyr, "asmiPradelmodelsCebolla.rdata", sep=""))
```
#########################################  Cebolla ##########################################

```{r}
load( paste(savepath, currentyr, "asmiPradelmodelsCebolla.rdata", sep=""))

asmiPradelmodelsCebolla

f.table <- get.real(asmiPradelmodelsCebolla$Phi.siteWinter.p.dot.f.siteSummer, "f", se = TRUE) %>%
# f.table <- get.real(asmiPradelmodelsCebolla$Phi.siteWinter.p.dot.f.siteWinter, "f", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  dplyr::rename(f.estimates = estimate,
         f.lcl = lcl,
         f.ucl = ucl) %>%
  dplyr::select(group, Plot,f.estimates, f.lcl, f.ucl, time) %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 

f.table %>%
  dplyr::mutate(tmax_springSummer_bin = cut(tmax_springSummer, breaks = 3)) %>%
ggplot(     aes(ppt_winter, f.estimates, color = Plot)) +
  geom_point(position = position_dodge(width = .1)) +
  geom_errorbar(aes(ymin = f.lcl, ymax = f.ucl),position = position_dodge(width = .1))+
  stat_smooth(method = "lm")+
  theme_bw() +
  # facet_grid(~tmax_springSummer_bin, scales = "free")+
  ylab("Per capita recruitment")+
  scale_color_manual("Site", values = c("darkblue","skyblue"),
                     labels = c("1" = "Cebolla Mid",
                                "16" = "Cebolla North"))

f.table %>%
  dplyr::mutate(ppt_winter_bin = cut(ppt_winter, breaks = 2)) %>%
ggplot(     aes(tmax_springSummer, f.estimates, color = Plot)) +
  geom_point(position = position_dodge(width = .5)) +
  geom_errorbar(aes(ymin = f.lcl, ymax = f.ucl),position = position_dodge(width = .5))+
  stat_smooth(method = "gam")+
  theme_bw() +
  # facet_grid(~ppt_winter_bin, scales = "free")+
  ylab("Per capita recruitment")+
  scale_color_manual("Site", values = c("darkblue","skyblue"),
                     labels = c("1" = "Cebolla Mid",
                                "16" = "Cebolla North"))+
  xlab("Summer maximum temperature")

Phi.table <- get.real(asmiPradelmodelsCebolla$Phi.siteWinter.p.dot.f.siteSummer, "Phi", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  dplyr::rename(Phi.estimates = estimate,
         Phi.lcl = lcl,
         Phi.ucl = ucl) %>%
  dplyr::select(group, Plot,Phi.estimates, Phi.lcl, Phi.ucl, time) %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 

## Yes impactful
ggplot(Phi.table, aes(ppt_winter, Phi.estimates, color = Plot)) +
  geom_point(position = position_dodge(width = .1)) +
  geom_errorbar(aes(ymin = Phi.lcl, ymax = Phi.ucl),position = position_dodge(width = .1))+
  stat_smooth(method = "lm")+
  theme_bw() +
  ylab("Survival")+
  scale_color_manual("Site", values = c("darkblue","skyblue"),
                     labels = c("1" = "Cebolla Mid",
                                "16" = "Cebolla North"))

# Phi is winter, not summer temp
ggplot(Phi.table, aes(tmax_springSummer, Phi.estimates, color = Plot)) +
  geom_point(position = position_dodge(width = .1)) +
  geom_errorbar(aes(ymin = Phi.lcl, ymax = Phi.ucl),position = position_dodge(width = .1))+
  stat_smooth(method = "lm")+
  theme_bw() +
  ylab("Survival")+
  scale_color_manual("Site", values = c("darkblue","skyblue"),
                     labels = c("1" = "Cebolla Mid",
                                "16" = "Cebolla North"))
lambdaCC <- 
f.table %>%
  left_join(Phi.table) %>%
  mutate(Lambda.estimate = f.estimates + Phi.estimates) %>%
ggplot(   aes(as.numeric(as.character(time)), log(Lambda.estimate), color = Plot)) +
  geom_point(position = position_dodge(width = 0.1))+
  stat_smooth(method = "lm", position = position_dodge(width = 0.1)) +
  theme_bw() +
  xlab("Year") +
  ylab(expression(log(lambda))) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  ggtitle("a)")+
  scale_color_manual("Site", values = c("darkblue","skyblue"),
                     labels = c("1" = "Cebolla Mid",
                                "16" = "Cebolla North"))
  

f.table %>%
  left_join(Phi.table) %>%
  mutate(Lambda.estimate = f.estimates + Phi.estimates) %>%
ggplot(   aes(ppt_winter, log(Lambda.estimate), color = Plot)) +
  geom_point(position = position_dodge(width = 0.1))+
  stat_smooth(method = "lm", position = position_dodge(width = 0.1)) +
  theme_bw() +
  xlab("Winter precipitation") +
  ylab(expression(log(lambda))) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  ggtitle("a)")+
  scale_color_manual("Site", values = c("darkblue","skyblue"),
                     labels = c("1" = "Cebolla Mid",
                                "16" = "Cebolla North"))


f.table %>%
  left_join(Phi.table) %>%
  mutate(Lambda.estimate = f.estimates + Phi.estimates) %>%
ggplot(   aes(tmax_springSummer, log(Lambda.estimate), color = Plot)) +
  geom_point(position = position_dodge(width = 0.1))+
  stat_smooth(method = "lm", position = position_dodge(width = 0.1)) +
  theme_bw() +
  xlab("Summer maximum temperature") +
  ylab(expression(log(lambda))) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  ggtitle("a)")+
  scale_color_manual("Site", values = c("darkblue","skyblue"),
                     labels = c("1" = "Cebolla Mid",
                                "16" = "Cebolla North"))

```


### South Beaver Creek  

```{r}

CJS_asmi_Southbeaver <- asmi.raw %>%
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence)) %>%
  filter(AsMi_site_id > 3) %>%
  ungroup() %>%
  mutate(Obs = case_when(length > 0 ~ 1, 
                         length == 0 ~ 0,
                         is.na(length) ~ 0)) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence)) %>%
  unite(ch, Year1995:Year2025, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id)) %>%
  filter(grepl("1",ch)) %>%
  mutate(Plot = factor(Plot))

dmPlot_SouthBeaver <- model.matrix(~ -1 + Plot, CJS_asmi_Southbeaver)  
dimnames(dmPlot_SouthBeaver)[[2]][length(dimnames(dmPlot_SouthBeaver)[[2]])]  # Plot799

CJS_asmi_Southbeaver %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlot_SouthBeaver) %>%
  mutate(Plot799 = paste(Plot799, ";", sep = "")) %>%
  write.table(file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/",
                           currentyr,
                           "CJS_asmi_markPradelSouthBeaver.inp",sep=""),
              sep = " ", 
              col.names = FALSE, row.names = FALSE)

plotdfSouthBeaver <- CJS_asmi_Southbeaver %>%
  distinct(Site,Plot) %>%
  arrange(Plot) 

## Convert for MARK  
asmi.inpSouthBeaver <- convert.inp(paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/",
                              currentyr,
                              "CJS_asmi_markPradelSouthBeaver.inp",sep=""), 
                        group.df = plotdfSouthBeaver,
                        covariates = NULL,
                        use.comments = FALSE)


asmiprocSouthBeaver <- process.data(asmi.inpSouthBeaver, model = "Pradrec",
                         groups = "Plot",
                         begin.time = 1995)

asmiddlSouthBeaver <- make.design.data(asmiprocSouthBeaver)

## Add site to ddl
asmiddlSouthBeaver$Phi <- asmiddlSouthBeaver$Phi %>%
  left_join(plotdfSouthBeaver)
asmiddlSouthBeaver$f <- asmiddlSouthBeaver$f %>%
  left_join(plotdfSouthBeaver)

## Fix detection, p, to zero when no survey
pindexSouthBeaver <- asmiddlSouthBeaver$p %>%
  left_join(plotdfSouthBeaver)

p.idAddedlater <- pindexSouthBeaver %>%
  mutate(time = as.numeric(as.character(time)))%>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(598) & time < 2004))
p.idAdded1996 <- pindexSouthBeaver %>%
  mutate(time = as.numeric(as.character(time)))%>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(300,238) & time < 1996))  

p.SouthBeaverindicies <- c(p.idAdded1996$par.index, 
                           p.idAddedlater$par.index)
p.SouthBeavervalues <- rep(0, length(p.SouthBeaverindicies))


############### Add climate
# WinterPPTSouthBeaver <-  asmi.season %>%
#   filter(Prev12 >= 1995) %>%
#   filter(!(site %in% c("Cebolla Mid","Cebolla North"))) %>%
#   ungroup() %>%
#   filter(Variable == "ppt" & season == "winter") %>%
#   dplyr::rename(WinterPPT = Value) %>%
#   dplyr::mutate(WinterPPT_scaled = c(scale(WinterPPT))) %>%
#   dplyr::select(Plot, Prev12, WinterPPT, WinterPPT_scaled) %>%
#   mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
#   mutate(Plot = as.factor(Plot))


asmiddlSouthBeaver$Phi <- asmiddlSouthBeaver$Phi  %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 

asmiddlSouthBeaver$p <- asmiddlSouthBeaver$p  %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 

asmiddlSouthBeaver$f <- asmiddlSouthBeaver$f  %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 


# Which sites/plots were fenced when; when Open == 1, when fenced == 0; switched back, that didn't fix it
PlotSiteYrFence <- asmi.all2 %>%
  distinct(year,plot,site,fenced) %>%
  arrange(site,plot,year)
### Use PlotSiteYrFence to specify years when a plot was fenced

plotdfSBC <- plotdfSouthBeaver %>%
  mutate(Plot = as.integer(as.character(Plot)))

### expand
PlotSiteYrFence <- PlotSiteYrFence %>%
  filter(site > 3) %>%
  dplyr::select(year, plot, fenced) %>%
  tidyr::complete(year, plot, fill = list(fenced = "n")) %>%
  arrange(plot,year) %>%
  left_join(plotdfSBC, join_by("plot" == "Plot"))


asmiddlSouthBeaver$Phi$fence <-  ifelse(asmiddlSouthBeaver$Phi$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                               asmiddlSouthBeaver$Phi$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiddlSouthBeaver$Phi$fence <- as.factor(asmiddlSouthBeaver$Phi$fence)

## looks correct
asmiddlSouthBeaver$Phi %>% filter(fence == "1")

asmiddlSouthBeaver$p$fence <-  ifelse(asmiddlSouthBeaver$p$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                               asmiddlSouthBeaver$p$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiddlSouthBeaver$p$fence <- as.factor(asmiddlSouthBeaver$p$fence)

  
asmiddlSouthBeaver$f$fence <-  ifelse(asmiddlSouthBeaver$f$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                               asmiddlSouthBeaver$f$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiddlSouthBeaver$f$fence <- as.factor(asmiddlSouthBeaver$f$fence)


```



############################### South Beaver Creek ###########################################
```{r}

run.PradrecSBC <- function(){
  Phi.site = list(formula =  ~ Site:fence)
  Phi.siteWinter = list(formula =  ~ ppt_winter:Site:fence)
  Phi.siteSummer = list(formula =  ~ tmax_springSummer:Site:fence)
  Phi.siteWinterSummer = list(formula =  ~ ppt_winter:tmax_springSummer:Site:fence)
  
  f.site = list(formula =  ~ Site:fence)
  f.siteWinter = list(formula =  ~ ppt_winter:Site:fence)
  f.siteSummer = list(formula =  ~ tmax_springSummer:Site:fence)
  f.siteWinterSummer = list(formula =  ~ ppt_winter:tmax_springSummer:Site:fence)

  p.dot = list(formula =  ~ 1,
               fixed = list(index = p.SouthBeaverindicies, value = p.SouthBeavervalues))
  
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiprocSouthBeaver, ddl = asmiddlSouthBeaver)
  
  return(asmi.results)
}

asmiPradelmodelsSBC <- run.PradrecSBC()
save(asmiPradelmodelsSBC, file = paste(savepath, currentyr, "asmiPradelmodelsSBC.rdata", sep=""))
```
############################### South Beaver Creek ###########################################



```{r}

load(paste(savepath, currentyr, "asmiPradelmodelsSBC.rdata", sep=""))
asmiPradelmodelsSBC

f.tableSBCPradrec <- get.real(asmiPradelmodelsSBC$Phi.siteSummer.p.dot.f.siteWinter, "f", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  dplyr::rename(f.estimates = estimate,
         f.lcl = lcl,
         f.ucl = ucl) %>%
  dplyr::select(group, Plot,f.estimates, f.lcl, f.ucl, time) %>%
  mutate(time = as.numeric(as.character(time)),
         group = as.integer(as.character(group))) %>%
  left_join(PlotSiteYrFence, join_by("group" == "plot", "time" == "year")) %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 

pRecruitment <- ggplot(f.tableSBCPradrec, aes(ppt_winter, f.estimates, color = as.factor(Site))) +
                    geom_point(position = position_dodge(width = 3)) +
                    geom_errorbar(aes(ymin = f.lcl, ymax = f.ucl),position = position_dodge(width = 3))+
                    theme_bw() +
                    stat_smooth(method = "gam") +
                    ylab("Per capita recruitment")+
                    scale_color_manual("Site", values = c("darkblue","skyblue","gold","brown")) +
                    xlab("Winter Precipitation (mm)") +
                    facet_wrap(~ fenced,
                               labeller = labeller(fenced = c("n" = "Open",
                                                   "y" = "Fenced")))
  
Phi.tableSBCPradrec <- get.real(asmiPradelmodelsSBC$Phi.siteSummer.p.dot.f.siteWinter, "Phi", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  dplyr::rename(Phi.estimates = estimate,
         Phi.lcl = lcl,
         Phi.ucl = ucl) %>%
  dplyr::select(group, Plot,Phi.estimates, Phi.lcl, Phi.ucl,time) %>%
  mutate(time = as.numeric(as.character(time)),
         group = as.integer(as.character(group))) %>%
  left_join(PlotSiteYrFence, join_by("group" == "plot", "time" == "year")) %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 


lambdaSBC <- 
f.tableSBCPradrec %>%
  left_join(Phi.tableSBCPradrec) %>%
  mutate(Lambda.estimate = f.estimates + Phi.estimates) %>%
ggplot(   aes(as.numeric(as.character(time)), log(Lambda.estimate), color = as.factor(Site))) +
  geom_point(position = position_dodge(width = 0.5))+
  stat_smooth(method = "lm", position = position_dodge(width = 0.5), alpha = 0.15) +
  theme_bw() +
  xlab("Year") +
  ylab(expression(log(lambda))) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  ggtitle("b)")+
  scale_color_manual("Site", values = c("darkblue","skyblue","gold","brown")) +
  facet_wrap(~fenced, labeller = labeller(fenced = c("n" = "Open Plots", "y" = "Fenced Plots")),
             nrow = 2, scales = "free_y")
  

```


## Lambda for both Drainages   
```{r}

ggsave(paste(savepath, "Fig3PradelLambda.jpg", sep = ""),
       lambdaCC + lambdaSBC +
         plot_layout(design = "
                     ABB
                     #BB
                     "),
width=275, height=200,units='mm', dpi=300)
       

## Check for site 15 line
f.tableSBCPradrec %>%
  left_join(Phi.tableSBCPradrec) %>%
  mutate(Lambda.estimate = f.estimates + Phi.estimates) %>%
  filter(Site %in% c(15,26)) %>%
ggplot(   aes(as.numeric(as.character(time)), log(Lambda.estimate), color = as.factor(Site))) +
  geom_point(position = position_dodge(width = 0.1))+
  stat_smooth(method = "lm", position = position_dodge(width = 0.1), alpha = 0.15) +
  theme_bw() +
  xlab("Year") +
  ylab(expression(log(lambda))) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  ggtitle("b)")+
  scale_color_manual("Site", values = c("darkblue","skyblue","gold","brown")) +
  facet_wrap(~fenced, labeller = labeller(fenced = c("n" = "Open Plots", "y" = "Fenced Plots")),
             nrow = 2) 
       
```

# Per captia recruitment for MPM in Multistate CJS for annual report   
```{r}
# Cebolla Creek
f.mpm <- get.real(asmiPradelmodelsCebolla$Phi.siteWinter.p.dot.f.siteSummer, "f", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  group_by(Plot) %>%
  summarise(f.estimates = median(estimate),
         f.lcl = median(lcl),
         f.ucl = median(ucl)) 

f.mpmSBCsite <- get.real(asmiPradelmodelsSBC$Phi.siteSummer.p.dot.f.siteWinter, "f", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  mutate(time = as.numeric(as.character(time)),
         group = as.integer(as.character(group))) %>%
  left_join(PlotSiteYrFence, join_by("group" == "plot", "time" == "year")) %>%
  group_by(Site, fenced) %>%
  summarise(f.estimates = median(estimate),
         f.lcl = median(lcl),
         f.ucl = median(ucl)) 

f.mpmSBC <- get.real(asmiPradelmodelsSBC$Phi.siteSummer.p.dot.f.siteWinter, "f", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  mutate(time = as.numeric(as.character(time)),
         group = as.integer(as.character(group))) %>%
  left_join(PlotSiteYrFence, join_by("group" == "plot", "time" == "year")) %>%
  ungroup() %>%
  summarise(f.estimates = median(estimate),
         f.lcl = median(lcl),
         f.ucl = median(ucl)) 

```




Pradlambda
```{r}
# Cebolla
asmiprocPradlambda.CC <- process.data(asmi.inpCebolla, model = "Pradlambda",
                         groups = "Plot",
                         begin.time = 2014)
asmiddlPradlambda.CC <- make.design.data(asmiprocPradlambda.CC)

asmiddlPradlambda.CC$Phi <- asmiddlPradlambda.CC$Phi %>%
  left_join(plotdfCebolla)
asmiddlPradlambda.CC$Lambda <- asmiddlPradlambda.CC$Lambda %>%
  left_join(plotdfCebolla)


## Fix detection, p, to zero when no survey
pindexCebolla <- asmiddlPradlambda.CC$p %>%
  left_join(plotdfCebolla)

p.Cebolla89 <- pindexCebolla %>%
  mutate(time = as.numeric(as.character(time))) %>%
  mutate(group = as.numeric(as.character(group))) %>%
  filter((group %in% c(89) & 
            time >= 2019))

p.Cebolla.missing2024 <- pindexCebolla %>%
  filter((group %in% c("52") &
            time %in% c("2024")))

p.Cebollaindices <- unique(c(p.Cebolla89$par.index, 
                             p.Cebolla.missing2024$par.index)) 

p.Cebollavalues <- rep(0, length(p.Cebollaindices))


##Winter precipitation, summer maximum temp
asmiddlPradlambda.CC$Phi <- asmiddlPradlambda.CC$Phi  %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 

asmiddlPradlambda.CC$p <- asmiddlPradlambda.CC$p  %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 

asmiddlPradlambda.CC$Lambda <- asmiddlPradlambda.CC$Lambda  %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 



```


```{r}
run.asmi.lambda.CC <- function(){
  
  Phi.Site = list(formula =  ~ Site)
  Phi.SiteWinter = list(formula =  ~ ppt_winter:Site)
  Phi.SiteSummer = list(formula =  ~ tmax_springSummer:Site)
  Phi.SiteSummerWinter = list(formula =  ~ ppt_winter:SummerMax:Site)
  
  Lambda.Site = list(formula =  ~ Site)
  Lambda.SiteWinter = list(formula =  ~ ppt_winter:Site)
  Lambda.SiteSummer = list(formula =  ~ tmax_springSummer:Site)
  Lambda.SiteSummerWinter = list(formula =  ~ ppt_winter:tmax_springSummer:Site)

  p.dot = list(formula = ~ 1,
               fixed = list(index = p.Cebollaindices, value = p.Cebollavalues))
  
  asmi.model.list <- create.model.list("Pradlambda")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiprocPradlambda.CC, ddl = asmiddlPradlambda.CC)
  
  return(asmi.results)
}

asmi.lambdaresults.CC <- run.asmi.lambda.CC()
save(asmi.lambdaresults.CC, file =  paste(savepath, currentyr, "asmiLambda_CC.Rdata", sep=""))

```


```{r}

load(paste(savepath, currentyr, "asmiLambda_CC.Rdata", sep=""))

asmi.lambdaresults.CC

Pradlambda.tableCC <- get.real(asmi.lambdaresults.CC$Phi.SiteWinter.p.dot.Lambda.SiteSummer, "Lambda", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  dplyr::rename(Lambda.estimates = estimate,
         Lambda.lcl = lcl,
         Lambda.ucl = ucl) %>%
  dplyr::select(group, Plot,Lambda.estimates, Lambda.lcl, Lambda.ucl, time) %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) %>%
  mutate(logLambda = log(Lambda.estimates),
         log.lcl = log(Lambda.lcl),
         log.ucl = log(Lambda.ucl))

ggplot(Pradlambda.tableCC, aes(tmax_springSummer, logLambda, color = Plot)) +
  geom_point(position = position_dodge(width = .1)) +
  geom_errorbar(aes(ymin = log.lcl, ymax = log.ucl), position = position_dodge(width = .1))+
  theme_bw() +
  ylab(expression(lambda))+
  scale_color_manual("Site", values = c("darkblue","skyblue"),
                     labels = c("1" = "Cebolla Mid",
                                "16" = "Cebolla North"))+
  xlab("Summer maximum temperature")+
  geom_hline(yintercept = 0, linetype = "dotted")

## By site
get.real(asmi.lambdaresults.CC$Phi.SiteWinter.p.dot.Lambda.SiteSummer, "Lambda", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE)

```






```{r}
# South Beaver Creek
asmiprocPradlambda.SBC <- process.data(asmi.inpSouthBeaver, model = "Pradlambda",
                         groups = "Plot",
                         begin.time = 1995)
asmiddlPradlambda.SBC <- make.design.data(asmiprocPradlambda.SBC)


PlotSiteYrFence <- PlotSiteYrFence %>%
  mutate(Plot = as.factor(plot),
         time = as.factor(year))

asmiddlPradlambda.SBC$Phi <- asmiddlPradlambda.SBC$Phi %>%
  left_join(PlotSiteYrFence)
asmiddlPradlambda.SBC$Lambda <- asmiddlPradlambda.SBC$Lambda %>%
  left_join(PlotSiteYrFence)

## Fix detection, p, to zero when no survey
pindexSouthBeaver <- asmiddlPradlambda.SBC$p %>%
  left_join(PlotSiteYrFence)

p.idAddedlater <- pindexSouthBeaver %>%
  mutate(time = as.numeric(as.character(time)))%>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(598) & time < 2004))
p.idAdded1996 <- pindexSouthBeaver %>%
  mutate(time = as.numeric(as.character(time)))%>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(300,238) & time < 1996))  

p.SouthBeaverindicies <- c(p.idAdded1996$par.index, 
                           p.idAddedlater$par.index)
p.SouthBeavervalues <- rep(0, length(p.SouthBeaverindicies))

##Winter precipitation, summer maximum temp
asmiddlPradlambda.SBC$Phi <- asmiddlPradlambda.SBC$Phi  %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 

asmiddlPradlambda.SBC$p <- asmiddlPradlambda.SBC$p  %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 

asmiddlPradlambda.SBC$Lambda <- asmiddlPradlambda.SBC$Lambda  %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) 

```


```{r}
run.asmi.lambda.SBC <- function(){
  
  Phi.Site = list(formula =  ~ Site:fenced)
  Phi.SiteWinter = list(formula =  ~ ppt_winter:Site:fenced)
  Phi.SiteSummer = list(formula =  ~ tmax_springSummer:Site:fenced)
  Phi.SiteSummerWinter = list(formula =  ~ ppt_winter:tmax_springSummer:Site:fenced)
  
  Lambda.Site = list(formula =  ~ Site:fenced)
  Lambda.SiteWinter = list(formula =  ~ ppt_winter:Site:fenced)
  Lambda.SiteSummer = list(formula =  ~ tmax_springSummer:Site:fenced)
  Lambda.SiteSummerWinter = list(formula =  ~ ppt_winter:tmax_springSummer:Site:fenced)

  p.dot = list(formula = ~ 1,
               fixed = list(index = p.SouthBeaverindicies, value = p.SouthBeavervalues))
  
  asmi.model.list <- create.model.list("Pradlambda")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiprocPradlambda.SBC, ddl = asmiddlPradlambda.SBC)
  
  return(asmi.results)
}

asmi.lambdaresults.SBC <- run.asmi.lambda.SBC()
save(asmi.lambdaresults.SBC, file =  paste(savepath, currentyr, "asmiLambda_SBC.Rdata", sep=""))

```


```{r}

load(paste(savepath, currentyr, "asmiLambda_SBC.Rdata", sep=""))

asmi.lambdaresults.SBC

Pradlambda.tableSBC <- get.real(asmi.lambdaresults.SBC$Phi.SiteWinter.p.dot.Lambda.SiteWinter, "Lambda", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  dplyr::rename(Lambda.estimates = estimate,
         Lambda.lcl = lcl,
         Lambda.ucl = ucl) %>%
  dplyr::select(group, Plot,Lambda.estimates, Lambda.lcl, Lambda.ucl, time) %>%
  mutate(Prev12 = as.numeric(as.character(time)) +1 ) %>%
  left_join(asmi.season.wide, by = c("Prev12" = "Prev12", "Plot" = "Plot")) %>%
  left_join(PlotSiteYrFence) %>%
  mutate(logLambda = log(Lambda.estimates),
         log.lcl = log(Lambda.lcl),
         log.ucl = log(Lambda.ucl))

Pradlambda.tableSBC %>%
  distinct(time,Plot,Site,fenced) %>%
  filter(time %in% c(1995,2010,2023))

ggplot(Pradlambda.tableSBC, aes(ppt_winter, logLambda, color = Site)) +
  geom_point(position = position_dodge(width = .1)) +
  geom_errorbar(aes(ymin = log.lcl, ymax = log.ucl), position = position_dodge(width = .1))+
  theme_bw() +
  ylab(expression(lambda))+
  scale_color_manual("Site", values = c("darkblue","skyblue","gold","brown"))+
  xlab("Winter precipitation")+
  geom_hline(yintercept = 0, linetype = "dotted") +
  facet_wrap(~fenced)


## By site
get.real(asmi.lambdaresults.SBC$Phi.SiteWinter.p.dot.Lambda.SiteWinter, "Lambda", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  mutate(fence = case_when(time == 1995 ~ "n",
                           TRUE ~ "y")) %>%
  left_join(PlotSiteYrFence) %>%
  dplyr::select(estimate, lcl,ucl,Site,fenced) %>%
  arrange(Site, desc(fenced))

```


