---
title: "Pradel for annual report"
author: "Michelle DePrenger-Levin"
date: "2024-10-28"
output: html_document
---


```{r}
rm(list=ls())
```

To combine recruitment with multistage transitions and survival for a MPM at high and low precipitation
```{r}


currentyr <- as.numeric(format(as.Date(Sys.Date(),format="%Y-%m-%d"), "%Y"))
userpath <- "C:/Users/DePrengm/"
savepath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_AnnualReports/", currentyr, "_Astragalus-microcymbus_AnnualReport/", collapse = '', sep = '')


library(dplyr)
library(tidyr)
library(RMark)
library(boot)
library(ggplot2)
library(lme4)

userpath <- "C:/Users/DePrengm/"
rawdatapath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/", currentyr, "_Astragalus-microcymbus_RawData.csv", collapse = '', sep = '')
asmi.raw <- read.csv(rawdatapath, na.strings = "na")
# need to remove all plot 89, site 1 in 2020 and beyond
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 89 & asmi.raw$year > 2019),]
# Need to remove all plot 52 in 2024, data from tablet was lost, didn't save 
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 52 & asmi.raw$year == 2024),]


# In a Stage-Fate format. stage is year t, fate is year t+1; asmi.all2
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/stage-fate",currentyr,".Rdata",
           sep=""))

asmi.raw %>%
  filter(is.na(year))

asmi.raw <- asmi.raw %>%
  filter(!is.na(year))

```

Climate
```{r}

load( paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.climate", currentyr,".Rdata", sep=""))
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.annual", currentyr,".Rdata", sep=""))
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.season", currentyr,".Rdata", sep=""))
```

# Cebolla Creek
```{r}
CJS_asmi_Cebolla <- asmi.raw %>%
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence)) %>%
  filter(AsMi_site_id < 3) %>%
  ungroup() %>%
  mutate(Obs = case_when(length > 0 ~ 1, 
                         length == 0 ~ 0,
                         is.na(length) ~ 0)) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id)) %>%
  unite(ch, Year2014:Year2024, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id)) %>%
  filter(grepl("1",ch))

dmPlot_Cebolla <- model.matrix(~ -1 + Plot, CJS_asmi_Cebolla)  
dimnames(dmPlot_Cebolla)[[2]][length(dimnames(dmPlot_Cebolla)[[2]])]  # Plot98
```


Mark-recapture by plot    
```{r}

CJS_asmi_Cebolla %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlot_Cebolla) %>%
  mutate(Plot98 = paste(Plot98, ";", sep = "")) %>%
  write.table(file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/",
                           currentyr,
                           "CJS_asmi_markPradelCebolla.inp",sep=""),
              sep = " ", 
              col.names = FALSE, row.names = FALSE)

plotdfCebolla <- CJS_asmi_Cebolla %>%
  distinct(Site,Plot)

## Convert for MARK  
asmi.inpCebolla <- convert.inp(paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/",
                              currentyr,
                              "CJS_asmi_markPradelCebolla.inp",sep=""), 
                        group.df = plotdfCebolla,
                        covariates = NULL,
                        use.comments = FALSE)


asmiprocCebolla <- process.data(asmi.inpCebolla, model = "Pradrec",
                         groups = "Plot",
                         begin.time = 2014)

asmiddlCebolla <- make.design.data(asmiprocCebolla)

## Add site to ddl
asmiddlCebolla$Phi <- asmiddlCebolla$Phi %>%
  left_join(plotdfCebolla)
asmiddlCebolla$f <- asmiddlCebolla$f %>%
  left_join(plotdfCebolla)



## Fix detection, p, to zero when no survey
pindexCebolla <- asmiddlCebolla$p %>%
  left_join(plotdfCebolla)

p.Cebolla89 <- pindexCebolla %>%
  mutate(time = as.numeric(as.character(time))) %>%
  mutate(group = as.numeric(as.character(group))) %>%
  filter((group %in% c(89) & 
            time >= 2019))

p.Cebolla.missing2024 <- pindexCebolla %>%
  filter((group %in% c("52") &
            time %in% c("2024")))

p.Cebollaindices <- unique(c(p.Cebolla89$par.index, 
                             p.Cebolla.missing2024$par.index)) 

p.Cebollavalues <- rep(0, length(p.Cebollaindices))

```



Add climate variables, time varying  
Add fencing as a factor to South Beaver Creek
```{r}

##Winter precipitation, summer maximum temp

WinterPPTCebolla <-  asmi.season %>%
  filter(Prev12 >= 2014) %>%
  filter(site %in% c("Cebolla Mid","Cebolla North")) %>%
  ungroup() %>%
  filter(Variable == "ppt" & season == "winter") %>%
  rename(WinterPPT = Value) %>%
  dplyr::select(Plot, Prev12, WinterPPT) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))



asmiddlCebolla$Phi <- asmiddlCebolla$Phi  %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPTCebolla) 

asmiddlCebolla$p <- asmiddlCebolla$p  %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPTCebolla) 

asmiddlCebolla$f <- asmiddlCebolla$f  %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPTCebolla) 
  
SummerMaxCebolla <- asmi.season %>%
  filter(Prev12 >= 2014) %>%
  filter(site %in% c("Cebolla Mid","Cebolla North")) %>%
  ungroup() %>%
  filter(Variable == "tmax" & season == "springSummer") %>%
  rename(SummerMax = Value) %>%
  dplyr::select(Plot, Prev12, SummerMax) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))

asmiddlCebolla$Phi <- asmiddlCebolla$Phi %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(SummerMaxCebolla)

asmiddlCebolla$p <- asmiddlCebolla$p %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(SummerMaxCebolla)

asmiddlCebolla$f <- asmiddlCebolla$f %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(SummerMaxCebolla)

```




```{r}

run.PradrecCebolla <- function(){
  Phi.site = list(formula =  ~ Site)
  Phi.siteWinter = list(formula =  ~ Site:WinterPPT)
  Phi.siteWinterSummer = list(formula =  ~ Site:WinterPPT:SummerMax)
  Phi.siteSummer = list(formula =  ~ Site:SummerMax)
  
  f.site = list(formula =  ~ Site)
  f.siteWinter = list(formula =  ~ Site:WinterPPT)
  f.siteWinterSummer = list(formula =  ~ Site:WinterPPT:SummerMax)
  f.siteSummer = list(formula =  ~ Site:SummerMax)

  p.dot = list(formula =  ~ 1,
               fixed = list(index = p.Cebollaindices, value = p.Cebollavalues))
  
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiprocCebolla, ddl = asmiddlCebolla)
  
  return(asmi.results)
}

asmiPradelmodelsCebolla <- run.PradrecCebolla()
save(asmiPradelmodelsCebolla, file = paste(savepath, currentyr, "asmiPradelmodelsCebolla.rdata", sep=""))
```

```{r}
load( paste(savepath, currentyr, "asmiPradelmodelsCebolla.rdata", sep=""))

asmiPradelmodelsCebolla

f.table <- get.real(asmiPradelmodelsCebolla$Phi.siteWinter.p.dot.f.siteSummer, "f", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  rename(f.estimates = estimate,
         f.lcl = lcl,
         f.ucl = ucl) %>%
  select(group, Plot,f.estimates, f.lcl, f.ucl, time) %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPTCebolla) %>%
  left_join(SummerMaxCebolla)

ggplot(f.table, aes(SummerMax, f.estimates, color = Plot)) +
  geom_point(position = position_dodge(width = .1)) +
  geom_errorbar(aes(ymin = f.lcl, ymax = f.ucl),position = position_dodge(width = .1))+
  theme_bw() +
  ylab("Per capita recruitment")+
  scale_color_manual("Site", values = c("darkblue","skyblue"),
                     labels = c("1" = "Cebolla Mid",
                                "16" = "Cebolla North"))

Phi.table <- get.real(asmiPradelmodelsCebolla$Phi.siteWinter.p.dot.f.siteSummer, "Phi", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  rename(Phi.estimates = estimate,
         Phi.lcl = lcl,
         Phi.ucl = ucl) %>%
  select(group, Plot,Phi.estimates, Phi.lcl, Phi.ucl, time) %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPTCebolla) %>%
  left_join(SummerMaxCebolla)

ggplot(Phi.table, aes(WinterPPT, Phi.estimates, color = Plot)) +
  geom_point(position = position_dodge(width = 1)) +
  geom_errorbar(aes(ymin = Phi.lcl, ymax = Phi.ucl),position = position_dodge(width = 1))+
  theme_bw() +
  ylab("Survival")+
  scale_color_manual("Site", values = c("darkblue","skyblue"),
                     labels = c("1" = "Cebolla Mid",
                                "16" = "Cebolla North"))
lambdaCC <- 
f.table %>%
  left_join(Phi.table) %>%
  mutate(Lambda.estimate = f.estimates + Phi.estimates) %>%
ggplot(   aes(as.numeric(as.character(time)), log(Lambda.estimate), color = Plot)) +
  geom_point(position = position_dodge(width = 0.1))+
  stat_smooth(method = "lm", position = position_dodge(width = 0.1)) +
  theme_bw() +
  xlab("Year") +
  ylab(expression(log(lambda))) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  ggtitle("a)")+
  scale_color_manual("Site", values = c("darkblue","skyblue"),
                     labels = c("1" = "Cebolla Mid",
                                "16" = "Cebolla North"))
  

```


### South Beaver Creek  

```{r}

CJS_asmi_Southbeaver <- asmi.raw %>%
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence)) %>%
  filter(AsMi_site_id > 3) %>%
  ungroup() %>%
  mutate(Obs = case_when(length > 0 ~ 1, 
                         length == 0 ~ 0,
                         is.na(length) ~ 0)) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence)) %>%
  unite(ch, Year1995:Year2024, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id)) %>%
  filter(grepl("1",ch)) %>%
  mutate(Plot = factor(Plot))

dmPlot_SouthBeaver <- model.matrix(~ -1 + Plot, CJS_asmi_Southbeaver)  
dimnames(dmPlot_SouthBeaver)[[2]][length(dimnames(dmPlot_SouthBeaver)[[2]])]  # Plot799

CJS_asmi_Southbeaver %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlot_SouthBeaver) %>%
  mutate(Plot799 = paste(Plot799, ";", sep = "")) %>%
  write.table(file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/",
                           currentyr,
                           "CJS_asmi_markPradelSouthBeaver.inp",sep=""),
              sep = " ", 
              col.names = FALSE, row.names = FALSE)

plotdfSouthBeaver <- CJS_asmi_Southbeaver %>%
  distinct(Site,Plot) %>%
  arrange(Plot) 

## Convert for MARK  
asmi.inpSouthBeaver <- convert.inp(paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/",
                              currentyr,
                              "CJS_asmi_markPradelSouthBeaver.inp",sep=""), 
                        group.df = plotdfSouthBeaver,
                        covariates = NULL,
                        use.comments = FALSE)


asmiprocSouthBeaver <- process.data(asmi.inpSouthBeaver, model = "Pradrec",
                         groups = "Plot",
                         begin.time = 1995)

asmiddlSouthBeaver <- make.design.data(asmiprocSouthBeaver)

## Add site to ddl
asmiddlSouthBeaver$Phi <- asmiddlSouthBeaver$Phi %>%
  left_join(plotdfSouthBeaver)
asmiddlSouthBeaver$f <- asmiddlSouthBeaver$f %>%
  left_join(plotdfSouthBeaver)

## Fix detection, p, to zero when no survey
pindexSouthBeaver <- asmiddlSouthBeaver$p %>%
  left_join(plotdfSouthBeaver)

p.idAddedlater <- pindexSouthBeaver %>%
  mutate(time = as.numeric(as.character(time)))%>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(598) & time < 2004))
p.idAdded1996 <- pindexSouthBeaver %>%
  mutate(time = as.numeric(as.character(time)))%>%
  ## Added in 2004 %in% c(598), added in 1996 %in% c(300,238)
  filter((group %in% c(300,238) & time < 1996))  

p.SouthBeaverindicies <- c(p.idAdded1996$par.index, 
                           p.idAddedlater$par.index)
p.SouthBeavervalues <- rep(0, length(p.SouthBeaverindicies))


############### Add climate
WinterPPTSouthBeaver <-  asmi.season %>%
  filter(Prev12 >= 1995) %>%
  filter(!(site %in% c("Cebolla Mid","Cebolla North"))) %>%
  ungroup() %>%
  filter(Variable == "ppt" & season == "winter") %>%
  rename(WinterPPT = Value) %>%
  dplyr::select(Plot, Prev12, WinterPPT) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))


asmiddlSouthBeaver$Phi <- asmiddlSouthBeaver$Phi  %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPTSouthBeaver) 

asmiddlSouthBeaver$p <- asmiddlSouthBeaver$p  %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPTSouthBeaver) 

asmiddlSouthBeaver$f <- asmiddlSouthBeaver$f  %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(WinterPPTSouthBeaver) 
  
SummerMaxSouthBeaver <- asmi.season %>%
  filter(Prev12 >= 1995) %>%
  filter(!(site %in% c("Cebolla Mid","Cebolla North"))) %>%
  ungroup() %>%
  filter(Variable == "tmax" & season == "springSummer") %>%
  rename(SummerMax = Value) %>%
  dplyr::select(Plot, Prev12, SummerMax) %>%
  mutate(Prev12 = as.factor(Prev12)) %>%
  mutate(Plot = as.factor(Plot))

asmiddlSouthBeaver$Phi <- asmiddlSouthBeaver$Phi %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(SummerMaxSouthBeaver)

asmiddlSouthBeaver$p <- asmiddlSouthBeaver$p %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(SummerMaxSouthBeaver)

asmiddlSouthBeaver$f <- asmiddlSouthBeaver$f %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) +1)) %>%
  left_join(SummerMaxSouthBeaver)




# Which sites/plots were fenced when; when Open == 1, when fenced == 0; switched back, that didn't fix it
PlotSiteYrFence <- asmi.all2 %>%
  distinct(year,plot,site,fenced) %>%
  arrange(site,plot,year)
### Use PlotSiteYrFence to specify years when a plot was fenced
PlotSiteYrFence %>%
  filter(fenced == "y")

PlotSiteYrFence %>%
  filter(plot == 238) ## needs 1995
## Huh! 
PlotSiteYrFence %>%
  filter(plot == 606) ## needs 2022, 2023
PlotSiteYrFence %>%
  filter(plot == 300) ## needs 1995


## Add  
plot2381995 <- PlotSiteYrFence %>%
  filter(plot %in% c(300,238) &
           year == 1996) %>%
  mutate(year = 1995)
plot606 <- PlotSiteYrFence %>%
  filter(plot %in% c(606) &
           year == 2021) 
plot606 <- plot606 %>%
  bind_rows(plot606) %>%
  mutate(year = c(2022, 2023))

add2024 <- PlotSiteYrFence %>%
  filter(year == 2023) %>%
  mutate(year = 2024)
PlotSiteYrFence <- PlotSiteYrFence %>%
  bind_rows(add2024) %>%
  bind_rows(plot2381995) %>%
  bind_rows(plot606)

asmiddlSouthBeaver$Phi$fence <-  ifelse(asmiddlSouthBeaver$Phi$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                               asmiddlSouthBeaver$Phi$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiddlSouthBeaver$Phi$fence <- as.factor(asmiddlSouthBeaver$Phi$fence)

## looks correct
asmiddlSouthBeaver$Phi %>% filter(fence == "1")

asmiddlSouthBeaver$p$fence <-  ifelse(asmiddlSouthBeaver$p$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                               asmiddlSouthBeaver$p$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiddlSouthBeaver$p$fence <- as.factor(asmiddlSouthBeaver$p$fence)

  
asmiddlSouthBeaver$f$fence <-  ifelse(asmiddlSouthBeaver$f$time %in% PlotSiteYrFence$year[PlotSiteYrFence$fenced == "y"] &
                               asmiddlSouthBeaver$f$Plot %in% PlotSiteYrFence$plot[PlotSiteYrFence$fenced == "y"],1,0)
asmiddlSouthBeaver$f$fence <- as.factor(asmiddlSouthBeaver$f$fence)


```

```{r}

run.PradrecSBC <- function(){
  Phi.site = list(formula =  ~ Site:fence)
  Phi.siteWinter = list(formula =  ~ WinterPPT:Site:fence)
  Phi.siteSummer = list(formula =  ~ SummerMax:Site:fence)
  Phi.siteWinterSummer = list(formula =  ~ WinterPPT:SummerMax:Site:fence)
  
  f.site = list(formula =  ~ Site:fence)
  f.siteWinter = list(formula =  ~ WinterPPT:Site:fence)
  f.siteSummer = list(formula =  ~ SummerMax:Site:fence)
  f.siteWinterSummer = list(formula =  ~ WinterPPT:SummerMax:Site:fence)

  p.dot = list(formula =  ~ 1,
               fixed = list(index = p.SouthBeaverindicies, value = p.SouthBeavervalues))
  
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiprocSouthBeaver, ddl = asmiddlSouthBeaver)
  
  return(asmi.results)
}

asmiPradelmodelsSBC <- run.PradrecSBC()
save(asmiPradelmodelsSBC, file = paste(savepath, currentyr, "asmiPradelmodelsSBC.rdata", sep=""))
```

```{r}

load(paste(savepath, currentyr, "asmiPradelmodelsSBC.rdata", sep=""))
asmiPradelmodelsSBC

f.tableSBCPradrec <- get.real(asmiPradelmodelsSBC$Phi.siteSummer.p.dot.f.siteWinter, "f", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  rename(f.estimates = estimate,
         f.lcl = lcl,
         f.ucl = ucl) %>%
  select(group, Plot,f.estimates, f.lcl, f.ucl, time) %>%
  mutate(time = as.numeric(as.character(time)),
         group = as.integer(as.character(group))) %>%
  left_join(PlotSiteYrFence, join_by("group" == "plot", "time" == "year")) %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) + 1)) %>%
  left_join(WinterPPTSouthBeaver) %>%
  left_join(SummerMaxSouthBeaver)

pRecruitment <- ggplot(f.tableSBCPradrec, aes(WinterPPT, f.estimates, color = as.factor(site))) +
                    geom_point(position = position_dodge(width = 3)) +
                    geom_errorbar(aes(ymin = f.lcl, ymax = f.ucl),position = position_dodge(width = 3))+
                    theme_bw() +
                    ylab("Per capita recruitment")+
                    scale_color_manual("Site", values = c("darkblue","skyblue","gold","brown")) +
                    xlab("Winter Precipitation (mm)")
  
Phi.tableSBCPradrec <- get.real(asmiPradelmodelsSBC$Phi.siteSummer.p.dot.f.siteWinter, "Phi", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  rename(Phi.estimates = estimate,
         Phi.lcl = lcl,
         Phi.ucl = ucl) %>%
  select(group, Plot,Phi.estimates, Phi.lcl, Phi.ucl,time) %>%
  mutate(time = as.numeric(as.character(time)),
         group = as.integer(as.character(group))) %>%
  left_join(PlotSiteYrFence, join_by("group" == "plot", "time" == "year")) %>%
  mutate(Prev12 = as.factor(as.numeric(as.character(time)) + 1)) %>%
  left_join(WinterPPTSouthBeaver) %>%
  left_join(SummerMaxSouthBeaver)


lambdaSBC <- 
f.tableSBCPradrec %>%
  left_join(Phi.tableSBCPradrec) %>%
  mutate(Lambda.estimate = f.estimates + Phi.estimates) %>%
ggplot(   aes(as.numeric(as.character(time)), log(Lambda.estimate), color = as.factor(site))) +
  geom_point(position = position_dodge(width = 0.1))+
  stat_smooth(method = "lm", position = position_dodge(width = 0.1), alpha = 0.15) +
  theme_bw() +
  xlab("Year") +
  ylab(expression(log(lambda))) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  ggtitle("b)")+
  scale_color_manual("Site", values = c("darkblue","skyblue","gold","brown")) +
  facet_wrap(~fenced, labeller = labeller(fenced = c("n" = "Open Plots", "y" = "Fenced Plots")),
             nrow = 2)
  

```


## Lambda for both Drainages   
```{r}

ggsave(paste(savepath, "Fig3PradelLambda.jpg", sep = ""),
       lambdaCC + lambdaSBC +
         plot_layout(design = "
                     ABB
                     #BB
                     "),
width=275, height=200,units='mm', dpi=300)
       

## Check for site 15 line
f.tableSBCPradrec %>%
  left_join(Phi.tableSBCPradrec) %>%
  mutate(Lambda.estimate = f.estimates + Phi.estimates) %>%
  filter(site %in% c(15,26)) %>%
ggplot(   aes(as.numeric(as.character(time)), log(Lambda.estimate), color = as.factor(site))) +
  geom_point(position = position_dodge(width = 0.1))+
  stat_smooth(method = "lm", position = position_dodge(width = 0.1), alpha = 0.15) +
  theme_bw() +
  xlab("Year") +
  ylab(expression(log(lambda))) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  ggtitle("b)")+
  scale_color_manual("Site", values = c("darkblue","skyblue","gold","brown")) +
  facet_wrap(~fenced, labeller = labeller(fenced = c("n" = "Open Plots", "y" = "Fenced Plots")),
             nrow = 2) 
       
```

# Per captia recruitment for MPM in Multistate CJS for annual report   
```{r}
# Cebolla Creek
f.mpm <- get.real(asmiPradelmodelsCebolla$Phi.siteWinter.p.dot.f.siteSummer, "f", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  group_by(Plot) %>%
  summarise(f.estimates = median(estimate),
         f.lcl = median(lcl),
         f.ucl = median(ucl)) 

f.mpmSBCsite <- get.real(asmiPradelmodelsSBC$Phi.siteSummer.p.dot.f.siteWinter, "f", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  mutate(time = as.numeric(as.character(time)),
         group = as.integer(as.character(group))) %>%
  left_join(PlotSiteYrFence, join_by("group" == "plot", "time" == "year")) %>%
  group_by(site) %>%
  summarise(f.estimates = median(estimate),
         f.lcl = median(lcl),
         f.ucl = median(ucl)) 

f.mpmSBC <- get.real(asmiPradelmodelsSBC$Phi.siteSummer.p.dot.f.siteWinter, "f", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  mutate(time = as.numeric(as.character(time)),
         group = as.integer(as.character(group))) %>%
  left_join(PlotSiteYrFence, join_by("group" == "plot", "time" == "year")) %>%
  ungroup() %>%
  summarise(f.estimates = median(estimate),
         f.lcl = median(lcl),
         f.ucl = median(ucl)) 

```






####################################################################################################################
####################################################################################################################
####################################################################################################################
####################################################################################################################
## Get site based confidence intervals on lambda
```{r}
asmiprocSouthBeaverpradlambda <- process.data(asmi.inpSouthBeaver, model = "Pradlambda",
                         groups = "Plot",
                         begin.time = 1995)

asmiddlSouthBeaverPradlambda <- make.design.data(asmiprocSouthBeaverpradlambda)

## Add site to ddl
asmiddlSouthBeaverPradlambda$Phi <- asmiddlSouthBeaverPradlambda$Phi %>%
  left_join(plotdfSouthBeaver)
asmiddlSouthBeaverPradlambda$Lambda <- asmiddlSouthBeaverPradlambda$Lambda %>%
  left_join(plotdfSouthBeaver)

### Number and factor, fix
asmiddlSouthBeaverPradlambda$Phi <-  asmiddlSouthBeaverPradlambda$Phi %>%
  left_join(PSYF, join_by("group" == "plot", "time"== "year")) %>%
  mutate(fence = case_when(is.na(fenced) ~ "n",
                           TRUE ~ fenced))

asmiddlSouthBeaverPradlambda$p <- asmiddlSouthBeaverPradlambda$p %>%
  left_join(PSYF, join_by("group" == "plot", "time"== "year")) %>%
  mutate(fence = case_when(is.na(fenced) ~ "n",
                           TRUE ~ fenced))

asmiddlSouthBeaverPradlambda$Lambda <- asmiddlSouthBeaverPradlambda$Lambda %>%
  left_join(PSYF, join_by("group" == "plot", "time"== "year")) %>%
  mutate(fence = case_when(is.na(fenced) ~ "n",
                           TRUE ~ fenced))

## Fix detection, p, to zero when no survey
pindexSouthBeaverPradlambda <- asmiddlSouthBeaverPradlambda$p %>%
  left_join(plotdfSouthBeaver)

identical(pindexSouthBeaver, pindexSouthBeaverPradlambda)


run.PradlambdaSBC <- function(){
  Phi.dot = list(formula =  ~ Site:fence)
  Lambda.dot = list(formula =  ~ Site:fence)

  p.dot = list(formula =  ~ 1,
               fixed = list(index = p.SouthBeaverindicies, value = p.SouthBeavervalues))
  
  asmi.model.list <- create.model.list("Pradlambda")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiprocSouthBeaverpradlambda, ddl = asmiddlSouthBeaverPradlambda)
  
  return(asmi.results)
}

asmiPradellambdaSouthBeaver <- run.PradlambdaSBC()
save(asmiPradellambdaSouthBeaver, file = paste(savepath, currentyr, "asmiPradellambdaSouthBeaver.rdata", sep=""))

lambda.tableSBC <- get.real(asmiPradellambdaSouthBeaver$Phi.dot.p.dot.Lambda.dot, "Lambda", se = TRUE)

## Higher lambda while fenced
lambda.tableSBC  %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  rename(Lambda.estimates = estimate,
         Lambda.lcl = lcl,
         Lambda.ucl = ucl) %>%
  select(group, Plot,time, Time, Lambda.estimates, Lambda.lcl, Lambda.ucl) %>%
  mutate(fenced = case_when(Time == 0 ~ "0",
                            Time == 11 ~ "1"))


## Annual
run.PradlambdaannSBC <- function(){
  Phi.timesite = list(formula =  ~ time + Site:fence)
  Lambda.timesite = list(formula =  ~ time + Site:fence)

  p.dot = list(formula =  ~ 1,
               fixed = list(index = p.SouthBeaverindicies, value = p.SouthBeavervalues))
  
  asmi.model.list <- create.model.list("Pradlambda")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiprocSouthBeaverpradlambda, ddl = asmiddlSouthBeaverPradlambda)
  
  return(asmi.results)
}

asmiPradellambdaannSouthBeaver <- run.PradlambdaannSBC()
save(asmiPradellambdaannSouthBeaver, file = paste(savepath, currentyr, "asmiPradelLambdaSBC.rdata", sep=""))
```

```{r}

load(paste(savepath, currentyr, "asmiPradelLambdaSBC.rdata", sep=""))
lambda.tableSBGLambdaann <- get.real(asmiPradellambdaannSouthBeaver$Phi.timesite.p.dot.Lambda.timesite, 
                                     "Lambda", se = TRUE)

SouthBeaverLambdaPlot <- lambda.tableSBGLambdaann %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  left_join(plotdfSouthBeaver) %>%
  mutate(time = as.numeric(as.character(time))) %>%
ggplot(  aes(time, log(estimate), color = Site)) +
  geom_point(position = position_dodge(width = 0.2)) +
  geom_errorbar(position = position_dodge(width = 0.2), aes(ymin = log(lcl), ymax = log(ucl)), width = 0.1) +
  stat_smooth(method = "lm") +
  theme_bw() +
  scale_color_manual("South Beaver Creek",
                     values = hcl.colors(4,palette = "Hawaii"),
                     labels = c("5" = "05", "15" = "15", "19" = "19", "26" = "26")) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  ylab(expression(log(frac(N[t+1],N[t])))) +
  xlab("Year t")

lambda.tableSBGLambdaann %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  left_join(plotdfSouthBeaver) %>%
  mutate(time = as.numeric(as.character(time))) %>%
  mutate(log.estimate = log(estimate)) %>%
  filter(log.estimate > -10 & log.estimate < 4) %>%
ggplot(  aes(time, log(estimate), color = Site)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(position = position_dodge(width = 0.5), aes(ymin = log(lcl), ymax = log(ucl)), width = 0.75) +
  stat_smooth(method = "lm") +
  theme_bw() +
  scale_color_manual("South Beaver Creek",
                     values = hcl.colors(4,palette = "Hawaii"),
                     labels = c("5" = "05", "15" = "15", "19" = "19", "26" = "26")) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  ylab(expression(log(frac(N[t+1],N[t])))) +
  xlab("Year t") 
  
```

#########################################################################################################################
#########################################################################################################################
#########################################################################################################################



Pradlambda
```{r}

asmiprocPradlambda <- process.data(asmi.inp, model = "Pradlambda",
                         groups = "Plot",
                         begin.time = 1995)
asmiddlPradlambda <- make.design.data(asmiprocPradlambda)


run.asmi.lambda <- function(){
  Phi.timePlot = list(formula =  ~ time:Plot)
  Lambda.timeplot = list(formula =  ~ time:Plot)

  p.plot = list(formula = ~ Plot,
               fixed = list(index = p.indices, value = p.values))
  
  asmi.model.list <- create.model.list("Pradlambda")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiprocPradlambda, ddl = asmiddlPradlambda)
  
  return(asmi.results)
}

asmi.lambdaresults <- run.asmi.lambda()
save(asmi.lambdaresults, file = "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmiLambda.Rdata")

```
