---
title: "RMark_Astragalus-microcymbus"
author: "Michelle DePrenger-Levin"
date: "2023-10-24"
output: html_document
---
RMark
```{r}
library(dplyr)
library(tidyr)
library(RMark)
library(boot)

## capture histories 2013-2015, secondary capture occasions within in closed periods
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Observations3years.Rdata")

currentyr <- 2022

userpath <- "C:/Users/DePrengm/"
# currentyr <- 2022

rawdatapath <- paste(userpath, "Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/", currentyr, "_Astragalus-microcymbus_RawData.csv", collapse = '', sep = '')

asmi.raw <- read.csv(rawdatapath, na.strings = "na")

# need to remove all plot 89, site 1 in 2020 and beyond
asmi.raw <- asmi.raw[!(asmi.raw$AsMi_plot_id == 89 & asmi.raw$year > 2019),]


## Raw data but kept dormancy to 3 or fewer years, then given a new tag number
asmi.raw3 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/Chapter AsMi/Dormant3yrmax_rawdata2022.csv")

## Load climate data up to 2022
load( paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.climate", currentyr,".Rdata", sep=""))
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.annual", currentyr,".Rdata", sep=""))
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.season", currentyr,".Rdata", sep=""))
load(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Astragalus-microcymbus/Astragalus-microcymbus_Data/asmi.season.wide", currentyr,".Rdata", sep=""))


```

Format data of all years
```{r}
asmi.raw <- asmi.raw3

# assign group at first encounter, age 0 if 'seedling' or vegetative, older if reproductive. 
CJS_asmi <- asmi.raw %>%
  filter(!(AsMi_tag_id %in% c(1177,1171,2746,3903))) %>%
  filter(!is.na(year)) %>%  
  filter(AsMi_site_id %in% c(5,15,19,26)) %>% ## exclude Cebolla Creek that was added 2014
  filter(AsMi_plot_id != 598) %>% ## exclude the one plot added in 2004
  filter(!(AsMi_plot_id %in% c(300,238))) %>% ## Since two plots, 300, 238 were added in 1996
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence)) %>%
  group_by(AsMi_tag_id) %>%
  filter(any(length > 0)) %>%
  mutate(first.cap.stage = status[which.min(year[length>0])]) %>%
  ungroup() %>%
  mutate(Obs = case_when(length > 0 ~ 1, 
                         length == 0 ~ 0,
                         is.na(length) ~ 0)) %>%
  arrange(year) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = 0, names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence,first.cap.stage)) %>%
  unite(ch, Year1995:Year2022, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id))

CJS_asmi %>%
  select(first.cap.stage:Plot) %>%
  # process.data() %>%
  # export.chdata("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_asmi_mark",
  #               covariates = c("first.cap.stage","Site","Plot"), replace = TRUE)
  save(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_asmi_mark.inp")



CJS_asmi %>%
  filter(first.cap.stage == "dormant")

table(CJS_asmi$first.cap.stage)

fenced <- asmi.raw %>%
  dplyr::select(c(year,AsMi_site_id, AsMi_plot_id,fence)) %>%
  group_by(year,AsMi_site_id, AsMi_plot_id,fence) %>%
  distinct(.keep_all = TRUE) %>%
  arrange(year, AsMi_site_id, AsMi_plot_id) %>%
  mutate(fence = if_else(year < 2006, "n", fence)) %>%
  mutate(fence = if_else(year == 2006 & AsMi_plot_id %in% c(300, 512), "n", fence)) %>%
  mutate(fence = if_else(year > 2015, "n", fence)) %>%
  mutate(fence = (as.factor(fence))) %>%
  mutate(fence = as.numeric(fence)-1) %>%
  pivot_wider(names_from = year, values_from = fence, values_fill = 0, 
              names_prefix = "Year", id_cols = c(AsMi_site_id,AsMi_plot_id))

fenced %>%
  select(Year2005:Year2017) %>%
  filter(AsMi_site_id == 19) 

asmi.data <- CJS_asmi %>%
  mutate(Seedlings = if_else(first.cap.stage == "seedling", 1, 0),
         Veg = if_else(first.cap.stage == "vegetative", 1, 0),
         Rep = if_else(first.cap.stage == "reproductive", 1, 0)) %>%
  mutate(fence = as.factor(fence),
         Site = as.factor(AsMi_site_id)) %>%
  dplyr::select(c(first.cap.stage:Rep)) %>%
  mutate(first.cap.stage = factor(first.cap.stage, levels = c("seedling","vegetative","reproductive"))) %>%
  as.data.frame() 

write.table(asmi.data[,c(2,4:8)], file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_asmi_mark.inp",
            sep = " ", col.names = FALSE, row.names = FALSE)


write.table(asmi.data[,c(2,4:8)], file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_asmi_mark.txt",
            sep = " ", col.names = TRUE, row.names = FALSE)




## simple CJS model, intercept only for survival and detection
asmi.model <- mark(asmi.data)
asmi.model$results$real

inv.logit(asmi.model$results$beta$estimate)
asmi.model$results$real
```


Create data frame of time varying covariates    
Merge time varying covariates into the deign matrix    
```{r}

asmi.proc.covar <- process.data(asmi.data, model = ) 

```


```{r}
## Not really a MARK dbf, doesn't have ";" at the end of each row
# asmi.mark <- convert.inp("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_asmi_mark.inp",
#                          group.df = data.frame(age = c("Seedlings","Veg","Rep")),
#                          covariates = c("Site","Plot"))
# asmi.mark <- import.chdata("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_asmi_mark.txt",
#                            field.types = c("f","f","f","f","f"))

## make.design.data should do the group data.frame


asmi.procmark <- process.data(asmi.data, model = "CJS",
                          age.var = 1,
                          initial.ages = c(0,1,2),
                          groups = c("first.cap.stage","Site","Plot"))

asmi.ddl <- make.design.data(asmi.procmark, 
                             parameters = list(Phi = list(age.bins = c(0,1,2,38)),
                             p = list(age.bins = c(1,2,3,39))),
                             right = FALSE)

asmi.proc <- process.data(asmi.data, model = "CJS",
                          groups = c("Site","Plot"))
asmi.ddl <- make.design.data(asmi.proc)

```

```{r}
# Create default design data  
# For Phi, values are 0 or older; for p the values are 1 or older
# This provides an individual covariate for each parameter type
# that changes as the animal ages
asmi.ddl = make.design.data(asmi.proc,
                          parameters = list(Phi = list(age.bins = c(0, 1, 28)),
                                            p = list(age.bins = c(1, 2, 29))),
                          right = FALSE)

## Could add a precipitation or temperature variable https://www.montana.edu/rotella/documents/502/lab05Rmark.html


```


Run competing models
```{r}
run.asmi = function() {
  #  Define range of models for Phi
  #
  Phi.dot = list(formula =  ~ 1)
  # Phi.age = list(formula =  ~ age)
  # Phi.age.site = list(formula =  ~ age + Site)
  # Phi.age.plot = list(formula =  ~ age + Plot)
  # Phi.age.Plot.Site = list(formula =  ~ age + Plot + Site)
  Phi.time = list(formula = ~ time)
  
  #  Define range of models for p
  p.dot = list(formula =  ~ 1)
  p.time = list(formula = ~ time)
  # p.time = list(formula =  ~ -1 + time)
  # p.time.age = list(formula = ~ age + time)
  # p.precip = list(formula = ~ precip)
  
  # Create models for all combinations of phi & p
  asmi.model.list = create.model.list("CJS")
  
  # NOTE: to avoid having all the output for each model appear when you
  # call the function, add ', output=FALSE' after 'ddl=sq.ddl' below.
  # Here, I don't do that so you can see the output for each model,
  # but this might not be desired if you have many models.
  asmi.results = mark.wrapper(asmi.model.list,
                            data = asmi.proc, ddl = asmi.ddl)
  #
  # Return model table and list of models
  #
  return(asmi.results)
}

# asmi.proc <- process.data(asmi.data, model = "CJS", groups = c("Site","Plot"))
# asmi.ddl <- make.design.data(asmi.proc)


asmi.res <- run.asmi()
## survival differs by age, detection over time and by age,
asmi.res
names(asmi.res)

asmi.res[[18]]$results$real
asmi.res$Phi.time.p.time.age$results$beta ## This has intercept and values on logit scale

phitable <- get.real(asmi.res$Phi.dot.p.dot, "Phi", se = TRUE)
ptable <- get.real(asmi.res$Phi.dot.p.dot, "p", se = TRUE)

phitable <- get.real(asmi.res$Phi.time.p.time, "Phi", se = TRUE)
ptable <- get.real(asmi.res$Phi.time.p.time, "p", se = TRUE)

asmi.res$Phi.time.p.time$results$beta
asmi.res$Phi.time.p.time$results$real
```

https://rdrr.io/cran/RMark/man/make.design.data.html  

https://rstudio-pubs-static.s3.amazonaws.com/201392_14ed145562b34fc785f5ff2ca00ba011.html 
# Abundance   
Bootstrap for uncertainty
```{r}
## Calc the number of recaptures/occasion
obs <- gregexec("1", asmi.data$ch)
n_obs <- summary(as.factor(unlist(obs)))
estim_abundance <- n_obs[-1]/ptable$estimate[1:27]

estim_abundance <- lapply(1:27)

## Cohort 1
plot(1996:2022, estim_abundance, type = "l")
points(1996:2022, estim_abundance)


nb_bootstrap <- 10
nb_years <- 28
target <- data.frame(asmi.data, stringsAsFactors = FALSE)
popsize <- matrix(NA, nb_bootstrap, nb_years-1)
set.seed(5)
pseudo <- target ## To initialize


asmi.procboot <- process.data(pseudo, model = "CJS")
asmi.ddlboot <- make.design.data(asmi.procboot)
phi.cat <- list(formula = ~ 1)
p.time <- list(formula = ~ 1)

# Run the bootstrap
for (k in 1:nb_bootstrap){
  # resample in the original dataset with replacement
  pseudo$ch = sample(target$ch, replace=T)
  # fit model with Mark
  res = mark(asmi.procboot,asmi.ddlboot,model.parameters=list(Phi=phi.cat,p=p.time),delete=TRUE,output=FALSE)
  # get recapture prob estimates
  ptable = get.real(res,"p", se= TRUE)
  # calculate the nb of recaptured individiduals / occasion
  allobs = gregexpr("1", pseudo$ch)
  n = summary(as.factor(unlist(allobs)))
  popsize[k,] <- n[-1]/ptable$estimate[1:(nb_years-1)]
}

## Get confidence intervals
ci_asmi <- apply(popsize,2,quantile,probs=c(2.5/100,97.5/100),na.rm=T)
ci_asmi

## 
plot(1996:2022, estim_abundance, type = "l", ylim = c(0,1700))
points(1996:2022, estim_abundance)
lines(1996:2022, ci_asmi[1,], col="red")
lines(1996:2022, ci_asmi[2,], col="red")
points(rawasmi$year, rawasmi$n, col = "blue")

rawasmi <- asmi.raw %>%
  group_by(year) %>%
  filter(!(AsMi_site_id %in% c(1,2))) %>%
  dplyr::summarise(n = n()) %>%
  print(n=28)
```


Make predictions for survival from top model
```{r}
min.tail <- min(sq$tail)
max.tail <- max(sq$tail)
tail.values <- seq(from = min.tail, to = max.tail, by = 0.25)
birth.values <- c(quantile(sq$birth, 0.05), 
                  mean(sq$birth),
                  quantile(sq$birth, 0.95))
pred.dat <- expand.grid(birth = birth.values,
                       tail = tail.values)
pred.dat$tail.sq <- pred.dat$tail^2

# make predictions for rows of 'sq.ddl' associated with juveniles, 
# i.e., (par.index=1, 2) = a juvenile and an adult 
pred.top <- covariate.predictions(top, data = pred.dat, indices = c(1, 2))

# view head of the prediction data.frame
head(pred.top$estimates)


```

Clean up files created
```{r}
rm(list=ls(all=TRUE))

cleanup(ask = FALSE)
```



## Just the three years with multiple
```{r}
CJS_asmi_3 <- asmi.raw %>%
  filter(!(AsMi_tag_id %in% c(1177,1171,2746,3903))) %>%
  filter(!is.na(year)) %>%  
  filter(year %in% c(2013,2014,2015)) %>%
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence)) %>%
  group_by(AsMi_tag_id) %>%
  filter(any(length > 0)) %>%
  mutate(first.cap.stage = status[which.min(year[length>0])]) %>%
  ungroup() %>%
  mutate(Obs = case_when(length > 0 ~ 1, 
                         length == 0 ~ 0,
                         is.na(length) ~ 0)) %>%
  arrange(year) %>%
  pivot_wider(names_from = year, values_from = Obs, values_fill = 0, names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id, fence,first.cap.stage)) %>%
  unite(ch, Year2013:Year2015, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id))


CJS_asmi_3 %>%
  dplyr::select(fence:Plot) %>%
  save(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_asmi_mark3.inp")


asmi.data3 <- CJS_asmi_3 %>%
  mutate(Seedlings = if_else(first.cap.stage == "seedling", 1, 0),
         Veg = if_else(first.cap.stage == "vegetative", 1, 0),
         Rep = if_else(first.cap.stage == "reproductive", 1, 0)) %>%
  mutate(fence = as.factor(fence),
         Site = as.factor(AsMi_site_id)) %>%
  dplyr::select(c(first.cap.stage:Rep)) %>%
  as.data.frame() 

asmi.model <- mark(asmi.data3)
asmi.model$results$real

```



https://www.montana.edu/rotella/documents/502/lab07RMark.html
closed population abundance   
Huggins model closed capture - overall abundance of sampled area 
```{r}

library(mra)

## obsmodel
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Observations3years.Rdata")

asmiclosed13 <- obsmodel %>%  
  group_by(Year, Tag, Site, Plot) %>%
  dplyr::mutate(state = ifelse(any(c_across('1':'9') == 3), "Reproductive", "Vegetative")) %>%
  dplyr::mutate(begin.time = min(which(c_across('1':'9')>0))) %>%
  filter(!is.infinite(begin.time)) %>%  ## There were a few cases with no observations in 2014
  dplyr::mutate(across('1':'9', ~ ifelse(.x > 0, 1, 0))) %>%
  unite(ch, '1':'4', sep = "") %>%
  dplyr::mutate(state = as.factor(state)) %>%
  filter(Year == 2013)
  
asmi.proc13 <- process.data(asmiclosed13, begin.time = 1, model = "Closed")
asmi.dll13 <- make.design.data(asmi.proc13)

asmi.dll13$p$time

```




https://rstudio-pubs-static.s3.amazonaws.com/201392_14ed145562b34fc785f5ff2ca00ba011.html    
Abundance using open capture-recapture    
```{r}


```



openCR     
```{r}

```



Check the impact of fencing
```{r}

asmi.proc <- process.data(asmi.data, age.var = 1, initial.ages = c(0,1,2), groups = c("age","Site"))
asmi.ddl <- make.design.data(asmi.proc)

## need to make those that are reproductive 2 or more years old and those that are small start at age 1 or something
# asmi.ddl$Phi
```

Define several competing models for survival 
```{r}
Phidot=list(formula=~1)
Phitime=list(formula=~time)
Phisite=list(formula=~AsMi_site_id)
Phisitetime=list(formula=~AsMi_site_id+time)
Phisite.time=list(formula=~AsMi_site_id*time)

pdot=list(formula=~1)
ptime=list(formula=~time)
psite=list(formula=~AsMi_site_id)
psitetime=list(formula=~AsMi_site_id+time)
psite.time=list(formula=~AsMi_site_id*time)

asmi.phitime.ptime <- mark(asmi.proc, asmi.ddl,
                           model.parameters = list(Phi=Phitime, p=ptime))
asmi.phidot.pdot <- mark(asmi.proc, asmi.ddl,
                           model.parameters = list(Phi=Phidot, p=pdot))
asmi.phidot.psite <- mark(asmi.proc, asmi.ddl,
                           model.parameters = list(Phi=Phidot, p=psite))
```

Make an AIC table
```{r}
asmi.table <- collect.models(type = "CJS")

```



```{r}
output.asmiphitime.ptime <- asmi.phitime.ptime$results$real %>%
  mutate(value = row.names(.)) %>%
  mutate(Year = c(1995:2021,1996:2022)) %>%
  separate(value, c("parameter", "group", "cohort", "age", "time"))
  
ggplot(output.asmiphitime.ptime, aes(Year, estimate, color = parameter)) +
  geom_line() +
  geom_point()+
  theme_bw()

plot(1995:2022, asmi.phitime.ptime$results$real$estimate[1:28], type = "l", ylim = c(0,1),
     xlab = "Year", ylab = "Phi")
points(1995:2022, asmi.phitime.ptime$results$real$estimate[1:28], pch = 16)
lines(1995:2022, asmi.phitime.ptime$results$real$lcl[1:28],  col = "red")
points(1995:2022, asmi.phitime.ptime$results$real$lcl[1:28], col = "red")
lines(1995:2022, asmi.phitime.ptime$results$real$ucl[1:28], col = "purple")
points(1995:2022, asmi.phitime.ptime$results$real$ucl[1:28],  col = "purple")

plot(1995:2022, asmi.phitime.ptime$results$real$estimate[29:54], type = "l", ylim = c(0,1),
     xlab = "Year", ylab = "p")
points(1995:2022, asmi.phitime.ptime$results$real$estimate[29:58], pch=16)
lines(1995:2022, asmi.phitime.ptime$results$real$lcl[29:56], col = "red")
lines(1995:2022, asmi.phitime.ptime$results$real$ucl[29:56], type = "l", col = "purple")




```



```{r}
obsmodel


CJS_AsMi2013 <- obsmodel %>%
  filter(Year == 2013) %>%
  group_by(Year, Tag, Site, Plot) %>%
  mutate(state = ifelse(any(c_across('1':'9') == 3), "Reproductive", "Vegetative")) %>%
  mutate(begin.time = min(which(c_across('1':'9')>0))) %>%
  filter(!is.infinite(begin.time)) %>%  ## There were a few cases with no observations in 2014
  mutate(across('1':'9', ~ ifelse(.x > 0, 1, 0))) %>%
  unite(col = ch, '1':'4', sep = "") %>%
  mutate(state = as.factor(state)) %>%
  dplyr::select(!('5':'9'))

CJS_AsMi2014_15 <- obsmodel %>%
  filter(Year != 2013) %>%
  group_by(Year, Tag, Site, Plot) %>%
  mutate(state = ifelse(any(c_across('1':'9') == 3), "Reproductive", "Vegetative")) %>%
  mutate(begin.time = min(which(c_across('1':'9')>0))) %>%
  filter(!is.infinite(begin.time)) %>%  ## There were a few cases with no observations in 2014
  mutate(across('1':'9', ~ ifelse(.x > 0, 1, 0))) %>%
  unite(col = ch, '1':'9', sep = "") %>%
  mutate(state = as.factor(state)) 


```

1. process data to identify model and data attributes   
Use state to define groups
```{r}

# CJS_ASMI <- CJS_AsMi2013 %>%
#   bind_rows(CJS_AsMi2014_15)
# 
# asmi.proc <- process.data(as.data.frame(CJS_ASMI), model="CJS", groups = "state", begin.time = 2013)

asmi.proc <- process.data(as.data.frame(CJS_AsMi2013), model="CJS", groups = "state")
asmi.proc$nocc 

asmi.proc1415 <- process.data(as.data.frame(CJS_AsMi2014_15), model = "CJS", groups = "state", begin.time = 2014)
asmi.proc1415$nocc
asmi.proc1415$nocc.secondary ## null?
asmi.proc1415$time.intervals

asmi.proc14 <- process.data(as.data.frame(CJS_AsMi2014_15[CJS_AsMi2014_15$Year == 2014,]),
                            model = "CJS")
asmi.proc14$nocc
asmi.proc14$time.intervals


```


2. make and modify design data    
Should add fence variable for individuals that were fenced
```{r}

asmi.ddl <- make.design.data(asmi.proc)
asmi.ddl$Phi

dm <- model.matrix(~state, asmi.ddl$Phi)
```


3. Write function for set of models to be fitted
```{r}
asmi.analysis <- function()
{
  # Create specifications for Phi and p
  Phi.1 = list(formula = ~1)
  Phi.2 = list(formula = ~ state)
  p.1 = list(formula = ~1)
  p.2 = list(formula = ~ state)
  
  # Create a list of combinations of parameters
  cml <- create.model.list("CJS")
  
  mark.wrapper(cml, data = asmi.proc, ddl = asmi.ddl, output = FALSE)
}


```


4. Run functions to fit models
```{r}
asmi.results <- asmi.analysis()
```





## marked Laake et al 2013   
```{r}
library(marked)
library(boot)

# CJS_AsMi_fence13
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_2013.Rdata")

# CJS_AsMi_fence1415
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/CJS_201415.Rdata")

CJS_AsMi_fence13 <- CJS_AsMi_fence13 %>%
  mutate(fence = as.factor(fence)) %>%
  as.data.frame()

## Select among CJS, JS, or probit CJS  
asmi13.proc <- marked::process.data(CJS_AsMi_fence13, groups = c("state","fence"), model = "CJS")
                            # model = "probitCJS") ## "Huggins ??, probit MCMC

# Make data desgin matrix for each parameter of the model
design.Phi <- list(static = c("state","fence"))
design.p <- list(static = c("state","fence"))
design.params <- list(Phi=design.Phi, p=design.p)
asmi13.ddm <- marked::make.design.data(asmi13.proc, parameters = design.params)  ## ageing based on time interval, starts at 0 
# model.parameters list of formulats for the parameters 
p.state <- list(formula = ~state) 
p.fence <- list(formula = ~ fence)
p.statefence <- list(formula = ~ state + fence)
p.statefence2 <- list(formula = ~ state * fence)
model13 <- crm(asmi13.proc, asmi13.ddm, model.parameters = list(p = p.state), accumulate = FALSE)
model13fence <- crm(asmi13.proc, asmi13.ddm, model.parameters = list(p = p.fence), accumulate = FALSE)
model13statefence <- crm(asmi13.proc, asmi13.ddm, model.parameters = list(p = p.statefence), accumulate = FALSE)
model13statefence2 <- crm(asmi13.proc, asmi13.ddm, model.parameters = list(p = p.statefence2), accumulate = FALSE)


inv.logit(model13$results$beta$p$mode) ## state, veg, reproductive
inv.logit(model13fence$results$beta$p$mode) ## open, fenced
inv.logit(model13statefence$results$beta$p$mode) ## open and veg, open and reproductive, fenced and veg
inv.logit(model13statefence2$results$beta$p$mode) ## open and veg, open and reproductive, fenced and veg, fenced and repro


paste(round(inv.logit(model13$results$beta$Phi$mode),2), " (",
      round(inv.logit(model13$results$beta$Phi$CI.lower),2), ", ",
      round(inv.logit(model13$results$beta$Phi$CI.upper),2), ")", sep="")

paste(round(inv.logit(model13$results$beta$p$mode[1]),2), " (",
      round(inv.logit(model13$results$beta$p$CI.lower[1]),2), ", ",
      round(inv.logit(model13$results$beta$p$CI.upper[1]),2), ")", sep="")

paste(round(inv.logit(sum(model13$results$beta$p$mode)),2), " (",
      round(inv.logit(sum(model13$results$beta$p$CI.lower)),2), ", ",
      round(inv.logit(sum(model13$results$beta$p$CI.upper)),2), ")", sep="")


# inv.logit(model13fence$results$beta$p$mode) ## open, fenced
# inv.logit(sum(model13fence$results$beta$p$mode)) # fenced
paste(round(inv.logit(model13fence$results$beta$Phi$mode),2), " (",
      round(inv.logit(model13fence$results$beta$Phi$CI.lower),2), ", ",
      round(inv.logit(model13fence$results$beta$Phi$CI.upper),2), ")", sep="")

paste(round(inv.logit(model13fence$results$beta$p$mode[1]),2), " (",
      round(inv.logit(model13fence$results$beta$p$CI.lower[1]),2), ", ",
      round(inv.logit(model13fence$results$beta$p$CI.upper[1]),2), ")", sep="")

paste(round(inv.logit(sum(model13fence$results$beta$p$mode)),2), " (",
      round(inv.logit(sum(model13fence$results$beta$p$CI.lower)),2), ", ",
      round(inv.logit(sum(model13fence$results$beta$p$CI.upper)),2), ")", sep="")


inv.logit(model13statefence$results$beta$p$mode) ## open and veg, open and reproductive, fenced and veg
paste(round(inv.logit(model13statefence$results$beta$Phi$mode),2), " (",
      round(inv.logit(model13statefence$results$beta$Phi$CI.lower),2), ", ",
      round(inv.logit(model13statefence$results$beta$Phi$CI.upper),2), ")", sep="")
# Open and Veg
paste(round(inv.logit(model13statefence$results$beta$p$mode[1]),2), " (",
      round(inv.logit(model13statefence$results$beta$p$CI.lower[1]),2), ", ",
      round(inv.logit(model13statefence$results$beta$p$CI.upper[1]),2), ")", sep="")
# open and reproductive
paste(round(inv.logit(sum(model13statefence$results$beta$p$mode[1:2])),2), " (",
      round(inv.logit(sum(model13statefence$results$beta$p$CI.lower[1:2])),2), ", ",
      round(inv.logit(sum(model13statefence$results$beta$p$CI.upper[1:2])),2), ")", sep="")
# fenced and vegetative
paste(round(inv.logit(sum(model13statefence$results$beta$p$mode[c(1,3)])),2), " (",
      round(inv.logit(sum(model13statefence$results$beta$p$CI.lower[c(1,3)])),2), ", ",
      round(inv.logit(sum(model13statefence$results$beta$p$CI.upper[c(1,3)])),2), ")", sep="")

paste(round(inv.logit(sum(model13statefence$results$beta$p$mode)),2), " (",
      round(inv.logit(sum(model13statefence$results$beta$p$CI.lower)),2), ", ",
      round(inv.logit(sum(model13statefence$results$beta$p$CI.upper)),2), ")", sep="")



## Compare several models 


fit.models = function(proc, ddl)
{
  Phi.dot = list(formula = ~ 1)
  p.state = list(formula = ~ state)
  p.time = list(formula = ~ time)
  p.statetime = list(formula = ~ time + state)
  p.fence = list(formula = ~ fence)
  p.fencestate = list(formula = ~ fence + state)
   cml = marked::create.model.list(c("Phi","p"))
   results= marked::crm.wrapper(cml,data = proc, ddl = ddl,
                       external=FALSE,accumulate=FALSE)
   return(results)
}

models.2013 <- fit.models(asmi13.proc,asmi13.ddm)



```

2014 and 2015 
```{r}
CJS_AsMi_fence14 <- CJS_AsMi_fence1415 %>%
  mutate(fence = as.factor(fence)) %>%
  filter(Year == 2014) %>%
  as.data.frame()

## Select among CJS, JS, or probit CJS  
asmi14.proc <- process.data(CJS_AsMi_fence14, groups = c("state","fence"), model = "probitCJS") ## "Huggins ??, probit MCMC
# Make data desgin matrix for each parameter of the model
asmi14.ddm <- make.design.data(asmi14.proc)  ## ageing based on time interval, starts at 0 
# model.parameters list of formulats for the parameters 
model14 <- crm(asmi14.proc, asmi14.ddm, model.parameters = list(p = p.state), accumulate = FALSE)
model14fence <- crm(asmi14.proc, asmi14.ddm, model.parameters = list(p = p.fence), accumulate = FALSE)
model14statefence <- crm(asmi14.proc, asmi14.ddm, model.parameters = list(p = p.statefence), accumulate = FALSE)
model14statefence2 <- crm(asmi14.proc, asmi14.ddm, model.parameters = list(p = p.statefence2), accumulate = FALSE)


inv.logit(model14$results$beta$p$mode) ## state, veg, reproductive
inv.logit(model14fence$results$beta$p$mode) ## open, fenced
inv.logit(model14statefence$results$beta$p$mode) ## open and veg, open and reproductive, fenced and veg
inv.logit(model14statefence2$results$beta$p$mode) ## open and veg, open and reproductive, fenced and veg, fenced and repro

paste(round(inv.logit(model14$results$beta$Phi$mode),2), " (",
      round(inv.logit(model14$results$beta$Phi$CI.lower),2), ", ",
      round(inv.logit(model14$results$beta$Phi$CI.upper),2), ")", sep="")

paste(round(inv.logit(model14$results$beta$p$mode[1]),2), " (",
      round(inv.logit(model14$results$beta$p$CI.lower[1]),2), ", ",
      round(inv.logit(model14$results$beta$p$CI.upper[1]),2), ")", sep="")

paste(round(inv.logit(sum(model14$results$beta$p$mode)),2), " (",
      round(inv.logit(sum(model14$results$beta$p$CI.lower)),2), ", ",
      round(inv.logit(sum(model14$results$beta$p$CI.upper)),2), ")", sep="")


# inv.logit(model14fence$results$beta$p$mode) ## open, fenced
# inv.logit(sum(model14fence$results$beta$p$mode)) # fenced
# Open vs. fenced
paste(round(inv.logit(model14fence$results$beta$Phi$mode[1]),2), " (",
      round(inv.logit(model14fence$results$beta$Phi$CI.lower[1]),2), ", ",
      round(inv.logit(model14fence$results$beta$Phi$CI.upper[1]),2), ")", sep="")

paste(round(inv.logit(model14fence$results$beta$p$mode[1]),2), " (",
      round(inv.logit(model14fence$results$beta$p$CI.lower[1]),2), ", ",
      round(inv.logit(model14fence$results$beta$p$CI.upper[1]),2), ")", sep="")

paste(round(inv.logit(sum(model14fence$results$beta$p$mode)),2), " (",
      round(inv.logit(sum(model14fence$results$beta$p$CI.lower)),2), ", ",
      round(inv.logit(sum(model14fence$results$beta$p$CI.upper)),2), ")", sep="")


inv.logit(model14statefence$results$beta$p$mode) ## open and veg, open and reproductive, fenced and veg
paste(round(inv.logit(model14statefence$results$beta$Phi$mode[1]),2), " (",
      round(inv.logit(model14statefence$results$beta$Phi$CI.lower[1]),2), ", ",
      round(inv.logit(model14statefence$results$beta$Phi$CI.upper[1]),2), ")", sep="")
# Open and Veg
paste(round(inv.logit(model14statefence$results$beta$p$mode[1]),2), " (",
      round(inv.logit(model14statefence$results$beta$p$CI.lower[1]),2), ", ",
      round(inv.logit(model14statefence$results$beta$p$CI.upper[1]),2), ")", sep="")
# open and reproductive
paste(round(inv.logit(sum(model14statefence$results$beta$p$mode[1:2])),2), " (",
      round(inv.logit(sum(model14statefence$results$beta$p$CI.lower[1:2])),2), ", ",
      round(inv.logit(sum(model14statefence$results$beta$p$CI.upper[1:2])),2), ")", sep="")
# fenced and vegetative
paste(round(inv.logit(sum(model14statefence$results$beta$p$mode[c(1,3)])),2), " (",
      round(inv.logit(sum(model14statefence$results$beta$p$CI.lower[c(1,3)])),2), ", ",
      round(inv.logit(sum(model14statefence$results$beta$p$CI.upper[c(1,3)])),2), ")", sep="")

paste(round(inv.logit(sum(model14statefence$results$beta$p$mode)),2), " (",
      round(inv.logit(sum(model14statefence$results$beta$p$CI.lower)),2), ", ",
      round(inv.logit(sum(model14statefence$results$beta$p$CI.upper)),2), ")", sep="")



##########################################################################################################
CJS_AsMi_fence15 <- CJS_AsMi_fence1415 %>%
  mutate(fence = as.factor(fence)) %>%
  filter(Year == 2015) %>%
  as.data.frame()

## Select among CJS, JS, or probit CJS  
asmi15.proc <- process.data(CJS_AsMi_fence15, groups = c("state","fence"), model = "probitCJS") ## "Huggins ??, probit MCMC
# Make data desgin matrix for each parameter of the model
asmi15.ddm <- make.design.data(asmi15.proc)  ## ageing based on time interval, starts at 0 
# model.parameters list of formulats for the parameters 
model15 <- crm(asmi15.proc, asmi15.ddm, model.parameters = list(p = p.state), accumulate = FALSE)
model15fence <- crm(asmi15.proc, asmi15.ddm, model.parameters = list(p = p.fence), accumulate = FALSE)
model15statefence <- crm(asmi15.proc, asmi15.ddm, model.parameters = list(p = p.statefence), accumulate = FALSE)
model15statefence2 <- crm(asmi15.proc, asmi15.ddm, model.parameters = list(p = p.statefence2), accumulate = FALSE)


# inv.logit(model15$results$beta$p$mode) ## state, veg, repro offset
# inv.logit(sum(model15$results$beta$p$mode)) # reproductive
# inv.logit(model15$results$beta$p$CI.upper) ## state, veg, repro offset
# inv.logit(sum(model15$results$beta$p$CI.upper)) # reproductive
# inv.logit(model15$results$beta$p$CI.lower) ## state, veg, repro offset
# inv.logit(sum(model15$results$beta$p$CI.lower)) # reproductive

paste(round(inv.logit(model15$results$beta$Phi$mode),2), " (",
      round(inv.logit(model15$results$beta$Phi$CI.lower),2), ", ",
      round(inv.logit(model15$results$beta$Phi$CI.upper),2), ")", sep="")

paste(round(inv.logit(model15$results$beta$p$mode[1]),2), " (",
      round(inv.logit(model15$results$beta$p$CI.lower[1]),2), ", ",
      round(inv.logit(model15$results$beta$p$CI.upper[1]),2), ")", sep="")

paste(round(inv.logit(sum(model15$results$beta$p$mode)),2), " (",
      round(inv.logit(sum(model15$results$beta$p$CI.lower)),2), ", ",
      round(inv.logit(sum(model15$results$beta$p$CI.upper)),2), ")", sep="")


# inv.logit(model15fence$results$beta$p$mode) ## open, fenced
# inv.logit(sum(model15fence$results$beta$p$mode)) # fenced
paste(round(inv.logit(model15fence$results$beta$Phi$mode),2), " (",
      round(inv.logit(model15fence$results$beta$Phi$CI.lower),2), ", ",
      round(inv.logit(model15fence$results$beta$Phi$CI.upper),2), ")", sep="")

paste(round(inv.logit(model15fence$results$beta$p$mode[1]),2), " (",
      round(inv.logit(model15fence$results$beta$p$CI.lower[1]),2), ", ",
      round(inv.logit(model15fence$results$beta$p$CI.upper[1]),2), ")", sep="")

paste(round(inv.logit(sum(model15fence$results$beta$p$mode)),2), " (",
      round(inv.logit(sum(model15fence$results$beta$p$CI.lower)),2), ", ",
      round(inv.logit(sum(model15fence$results$beta$p$CI.upper)),2), ")", sep="")


inv.logit(model15statefence$results$beta$p$mode) ## open and veg, open and reproductive, fenced and veg
paste(round(inv.logit(model15statefence$results$beta$Phi$mode),2), " (",
      round(inv.logit(model15statefence$results$beta$Phi$CI.lower),2), ", ",
      round(inv.logit(model15statefence$results$beta$Phi$CI.upper),2), ")", sep="")
# Open and Veg
paste(round(inv.logit(model15statefence$results$beta$p$mode[1]),2), " (",
      round(inv.logit(model15statefence$results$beta$p$CI.lower[1]),2), ", ",
      round(inv.logit(model15statefence$results$beta$p$CI.upper[1]),2), ")", sep="")
# open and reproductive
paste(round(inv.logit(sum(model15statefence$results$beta$p$mode[1:2])),2), " (",
      round(inv.logit(sum(model15statefence$results$beta$p$CI.lower[1:2])),2), ", ",
      round(inv.logit(sum(model15statefence$results$beta$p$CI.upper[1:2])),2), ")", sep="")
# fenced and vegetative
paste(round(inv.logit(sum(model15statefence$results$beta$p$mode[c(1,3)])),2), " (",
      round(inv.logit(sum(model15statefence$results$beta$p$CI.lower[c(1,3)])),2), ", ",
      round(inv.logit(sum(model15statefence$results$beta$p$CI.upper[c(1,3)])),2), ")", sep="")

paste(round(inv.logit(sum(model15statefence$results$beta$p$mode)),2), " (",
      round(inv.logit(sum(model15statefence$results$beta$p$CI.lower)),2), ", ",
      round(inv.logit(sum(model15statefence$results$beta$p$CI.upper)),2), ")", sep="")


inv.logit(model15statefence2$results$beta$p$mode) ## open and veg, open and reproductive, fenced and veg, fenced and repro



```


mra   
```{r}
library(mra)

```


